### 2015-11-09 paper.tex
import

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{agda} % this should come after ams*
\usepackage{catchfilebetweentags}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{ifdraft}
\usepackage{ifthen}
\usepackage{lualatex-math}
%\usepackage{mathtools}
\usepackage{proof}
\usepackage{scalerel}
\usepackage{setspace}
\usepackage{stmaryrd}
\usepackage{supertabular}
\usepackage{tikz-cd}
\usepackage{unicode-math}
\usepackage{url}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{xunicode}

\definecolor{dark-gray}{gray}{0.20}

% fonts
\usepackage[
  activate={true,nocompatibility},
  tracking=true
]{microtype}
\setmainfont{Linux Libertine O}
\setmathfont{Asana Math}

% make lualatex play nice with agda
\makeatletter
\ifluatex%
   \newwrite\CATCHPATCHOUT%
\fi
\def\CatchFBT@Fin@l#1[#2]{%
   \begingroup%
      \endlinechar\m@ne\makeatletter #2%
      \ifluatex%
         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch%
         \immediate\write\CATCHPATCHOUT{\CatchFBT@tok{\the\CatchFBT@tok}}%
         \immediate\closeout\CATCHPATCHOUT%
         \@@input\jobname.catchpatch%
      \else%
      \scantokens\expandafter{%
         \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
      \fi%
      \CatchFBT@IsAToken{#1}%
         {\global#1\expandafter{\the\CatchFBT@tok}}%
         {\xdef#1{\the\CatchFBT@tok}}%
      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi%
   \endgroup%
}% \CatchFBT@Final
\makeatother

\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[thm]{Definition}
\newtheorem{example}[thm]{Example}
\theoremstyle{remark}
\newtheorem{remark}[thm]{Remark}
\numberwithin{equation}{section}

\def\IModeColorName{MidnightBlue}
\def\OModeColorName{Maroon}

\newcommand\IMode[1]{{\color{\IModeColorName}{#1}}}
\newcommand\OMode[1]{{\color{\OModeColorName}{#1}}}

\newcommand\HypJ[2]{#1\ \ (#2)}
\newcommand\GenJ[2]{\vert_{\IMode{#1}}\; #2}

\newcommand\IsTac[1]{\IMode{#1}\ \textit{tactic}}

\newcommand\MkSet[1]{\left\{#1\right\}}
\newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
\newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
\newcommand\Of[2]{\IMode{#1}: \IMode{#2}}

\newcommand\MkValence[3]{%
  \ifthenelse{\equal{#1}{}}{}{%
    \{#1\}%
  }%
  \ifthenelse{\equal{#2}{}}{}{%
    [#2]%
  }.\, #3%
}
\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
\newcommand\MkArity[2]{(#1)\, #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

\newcommand\ArrId[1]{1_{#1}}
\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}

\newcommand\Symbols{\mathbf{I}}
\newcommand\SymSets{\mathbb{F}_\Symbols}
\newcommand\FinSets{\mathbb{F}}
\newcommand\Sorts{\mathcal{S}}
\newcommand\Valences{\mathcal{V}}
\newcommand\Arities{\mathcal{A}}
\newcommand\Operators{\mathcal{O}}

\newcommand\Cats{\mathbf{Cat}}
\newcommand\Discrete[1]{#1_{\equiv}}
\newcommand\Arr[1]{\mathbf{{#1}^{\rightarrow}}}
\newcommand\Sets{\mathbf{Set}}
\newcommand\SCtx{\mathbf{SCtx}}
\newcommand\Ctx{\mathbf{Ctx}}
\newcommand\Seq{\mathbf{Seq}}
\newcommand\HypCat{\mathbf{H}}

\newcommand\OpCat[1]{#1^{\mathsf{op}}}

\newcommand\IsOperator[3]{\IMode{#1}\Vdash\IMode{#2}: \OMode{#3}}

\newcommand\IsMetaCtx[1]{\IMode{#1}\ \textit{mctx}}
\newcommand\IsVarCtx[1]{\IMode{#1}\ \textit{vctx}}
\newcommand\IsSymCtx[1]{\IMode{#1}\ \textit{sctx}}

\newcommand\Lookup[3]{\IMode{#1}\ni\IMode{#2}:\OMode{#3}}

\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
\newcommand\Pair[2]{\left\langle#1, #2 \right\rangle}
\newcommand\Tuple[1]{\left\langle#1 \right\rangle}
\newcommand\FLift[2]{
  \Operators\langle#1%
  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle%
}
\newcommand\PairTimes[2]{\langle#1\, {\scriptstyle{{\times}}}\, #2 \rangle}
\newcommand\SSingleton{\bullet}
\newcommand\Forget[1]{|\, #1\, |}

\newcommand\Comma[2]{#1 \downarrow#2}
\newcommand\Coslice[2]{#1\, /\, #2}

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsWf[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
}

\newcommand\IsBTm[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\IMode{#5}%
}

\newcommand\MV[1]{\mathfrak{#1}}
\newcommand\MApp[3]{#1\{#2\}(#3)}
\newcommand\App[2]{#1(#2)}
\newcommand\Dom[1]{\left\vert#1\right\vert}
\newcommand\FV[1]{\mathbf{FV}\left(#1\right)}
\newcommand\FS[1]{\mathbf{FS}\left(#1\right)}
\newcommand\NotIn[2]{\IMode{#1}\notin\IMode{#2}}
\newcommand\SortExp{\mathsf{exp}}

\newcommand\Hom[3]{#1\left[#2,#3\right]}

\newcommand\Rename[2]{%
  #2\upharpoonright#1%
}

\newcommand\IsRenaming[3]{%
  \Rename{\IMode{#1}}{\IMode{#2}}%
  \leadsto\OMode{#3}%
}

\newcommand\Subst[3]{%
  \left[#1\, /\, #2\right]#3%
}

\newcommand\IsSubst[4]{%
  \Subst{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}%
  \leadsto\OMode{#4}%
}

\newcommand\IsEq[2]{\IMode{#1} = \IMode{#2}}
\newcommand\IsApart[2]{\IMode{#1} \mathrel{\#} \IMode{#2}}

\newcommand\Define[2]{\IMode{#1}\triangleq\OMode{#2}}
\newcommand\ADefine[2]{\IMode{#1}&\triangleq\OMode{#2}}

\newcommand\SigFun[1]{\mathcal{F}_{#1}}

\newcommand\SemBrackets[1]{\left\llbracket#1\right\rrbracket}
\newcommand\Yoneda[1]{\boldsymbol{y} (#1)}

\begin{document}

\title{Type Refinements for the Working Class}
\date{}
\author{Jon Sterling and Darin Morrison}
\maketitle

There are two conflicting views which bedevil any discussion of the nature of
type theory. First, there is the notion of type theory as an extension or
generalization of universal algebra to support interdependency of sorts and
operations, possibly subject to an arbitrary equational theory
\cite{cartmell:1986, dybjer:1996}; we will call this \emph{formal type theory}.
Typing, in such a setting, is a mere matter of grammar and is nearly always
decidable. In hindsight, we may observe that this is the sort of type theory
which Martin-L\"of first proposed in 1972 \cite{martin-lof:1972}, even if we
will admit that this was not the intention at the time. A model for such a type
theory is usually given by interpreting the types or sorts as presheaves or
sheaves over contexts of hypotheses, and as such, a proof theoretic
interpretation of the hypothetical judgment is inevitable.

Secondly, there is the view of type theory as semi-formal theory of
constructions for the Brouwer-Heyting-Kolmogorov interpretation of
intuitionistic mathematical language, which we will call \emph{behavioral} or
\emph{semantic type theory}. The most widely known development of this program
is Martin-L\"of's 1979 ``extensional'' type theory \cite{martin-lof:1979,
martin-lof:1984}, but we must give priority to Dana Scott for inventing this
line of research in 1970 with his prophetic report, \emph{Constructive
Validity} \cite{scott:1970}. More recently, behavioral type theory has been
developed much further in the Nuprl family \cite{constable:1986} of proof
assistants, including MetaPRL \cite{hickey:2003} and JonPRL \cite{jonprl:2015}.

Martin-L\"of's key innovation was the commitment to pervasive functionality
(extensionality) as part of the \emph{definitions} of the judgments and the
types, in contrast to the state of affairs in formal type theory where
functionality is a metatheorem which may or may not hold depending on the
particular equational theory which has been imposed. Furthermore, models for
behavioral type theory interpret the types as partial equivalence relations on
only closed terms, and the meaning of the hypothetical judgment is defined
separately and uniformly in the logical relations style.

Our position is that these views of type theory are not in conflict, but rather
merely describe two distinct layers in a single, harmonious system. From this
perspective, formal type theories can do little more than negotiate matters of
grammar, and therefore they serve as a syntactic (linguistic) framework
for mathematical language, being responsible for the management of variable
binding and substitution. On the other hand, the meaning of mathematical
statements shall be characterized \emph{behaviorally} in the semantic type
theory.

The types of the semantic theory can then be said to \emph{refine} the types of
the syntactic theory, both by placing restrictions on membership and by
coarsening equivalence. Thus far, all developments of behavioral type theory
have been built on a \emph{unityped} syntactic framework, and so the relation
to type refinements has been difficult to see. In this paper, we contribute a
full theory of behavioral refinements over multi-sorted abstract binding trees,
a simple formal type theory \cite{harper:2016, sterling-morrison:2015}; this
hybrid system allows the deployment of a Nuprl-style type theory over any
signature of sorts and operators.

\ifdraft{}{
  \nocite{*}
  \bibliographystyle{abbrv}
  \bibliography{refs}
}

\end{document}

### 2015-11-09 paper.tex
wordsmithing

@@
 perspective, formal type theories can do little more than negotiate matters of
-grammar, and therefore they serve as a syntactic (linguistic) framework
+grammar, and therefore may serve as a syntactic (linguistic) framework
 for mathematical language, being responsible for the management of variable
 binding and substitution. On the other hand, the meaning of mathematical
-statements shall be characterized \emph{behaviorally} in the semantic type
+statements shall be specified \emph{behaviorally} in the semantic type
 theory.

### 2015-11-09 paper.tex
turn off no-cite
Add bob citations

@@
 The types of the semantic theory can then be said to \emph{refine} the types of
-the syntactic theory, both by placing restrictions on membership and by
-coarsening equivalence. Thus far, all developments of behavioral type theory
-have been built on a \emph{unityped} syntactic framework, and so the relation
-to type refinements has been difficult to see. In this paper, we contribute a
-full theory of behavioral refinements over multi-sorted abstract binding trees,
-a simple formal type theory \cite{harper:2016, sterling-morrison:2015}; this
-hybrid system allows the deployment of a Nuprl-style type theory over any
-signature of sorts and operators.
+the syntactic theory \cite{harper-davies:2014, harper-duff:2015, harper:2016},
+both by placing restrictions on membership and by coarsening equivalence. Thus
+far, all developments of behavioral type theory have been built on a
+\emph{unityped} syntactic framework, and so the relation to type refinements
+has been difficult to see. In this paper, we contribute a full theory of
+behavioral refinements over multi-sorted abstract binding trees, a simple
+formal type theory \cite{harper:2016, sterling-morrison:2015}; this hybrid
+system allows the deployment of a Nuprl-style type theory over any signature of
+sorts and operators.
 
 \ifdraft{}{
-  \nocite{*}
   \bibliographystyle{abbrv}

### 2015-11-09 paper.tex
begin developing meaning explanation for refinement

@@
 }
-\newcommand\IsWf[5]{%
-  \IMode{#1}\triangleright%
-  \IMode{#2}\parallel%
-  \IMode{#3}\vdash%
-  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
-}
-
 \newcommand\IsBTm[5]{%
@@
 \newcommand\Yoneda[1]{\boldsymbol{y} (#1)}
+
+\newcommand\Refines[3]{\IMode{#1}\parallel \IMode{#2} \sqsubset \OMode{#3}}
+\newcommand\GRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}
+\newcommand\RefinesCtx[2]{\IMode{#1}\sqsubset^\star\OMode{#2}}
 
@@
 
+\section{Abstract Binding Trees and Symbols}
+
+See \cite{sterling-morrison:2015} for the development of abstract binding trees
+with symbols.
+
+TODO: give a brief description of the framework, and present its rules.
+
+\section{Behavioral Refinements}
+
+Fixing a signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ in the abt
+framework, we will define the notion of behavioral refinement by propounding
+several judgments and their semantical explanations. We will require two
+judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$, which means that
+$\Phi$ is a context which refines the symbol context $\Upsilon$ (presupposing
+$\IsSymCtx{\Upsilon}$, and $\Refines{\Phi}{\phi}{\tau}$, which shall presuppose
+$\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
+
+\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathsf{nil}}}
+\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathsf{snoc}}}
+
+The first judgment we define inductively:
+\begin{gather*}
+  \infer[\RefinesCtxNil]{
+    \RefinesCtx{\cdot}{\cdot}
+  }{
+  }\qquad
+  \infer[\RefinesCtxSnoc]{
+    \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
+  }{
+    \RefinesCtx{\Phi}{\Upsilon} &
+    \Refines{\Phi}{\phi}{\tau}
+  }
+\end{gather*}
+
+The second judgment is defined coinductively via the following meaning
+explanation:
+\begin{quote}
+  To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
+  $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
+  $\Of{\varrho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, for any $M, N$,
+  what it means for $M$ and $N$ to be identified under $\phi$, supposing
+  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
+  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
+  $\IMode{M}\sim_\phi\IMode{N}$ then $\IMode{N}\sim_\phi\IMode{M}$, and if
+  $\IMode{M}\sim_\phi\IMode{N}$ and $\IMode{N}\sim_\phi\IMode{O}$, then
+  $\IMode{M}\sim_\phi\IMode{O}$.
+\end{quote}
+
+In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
+$\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
+$\Upsilon'$, which partial equivalence relation on the closed $\Upsilon'$-terms
+of $\tau$ that $\phi$ denotes. At this point, it should be remarked that this
+is very similar to the semantical explanation of typehood given in
+\cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
+setting, and that we have fibred the entire apparatus over collections
+$\Upsilon$ of symbols.
+
+\begin{thm}[Weakening]
+  If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
+  $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
+\end{thm}
+\begin{proof}
+  This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
+  $\Upsilon\hookrightarrow\Upsilon,u:\sigma$.
+\end{proof}
+
+When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ is an
+\emph{global} refinement and may simply write $\GRefines{\phi}{\tau}$. By
+weakening, if $\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
+$\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.
+
 \ifdraft{}{

### 2015-11-09 paper.tex
formatting

@@
 explanation:
-\begin{quote}
+\begin{definition}[Refinement]
   To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
@@
   $\IMode{M}\sim_\phi\IMode{O}$.
-\end{quote}
+\end{definition}
 

### 2015-11-09 paper.tex
typo

@@
 $\Phi$ is a context which refines the symbol context $\Upsilon$ (presupposing
-$\IsSymCtx{\Upsilon}$, and $\Refines{\Phi}{\phi}{\tau}$, which shall presuppose
+$\IsSymCtx{\Upsilon}$), and $\Refines{\Phi}{\phi}{\tau}$, which shall presuppose
 $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
@@
   $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
-  $\Of{\varrho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, for any $M, N$,
+  $\Of{\varrho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, and for any $M, N$,
   what it means for $M$ and $N$ to be identified under $\phi$, supposing

### 2015-11-09 paper.tex
wording

@@
 $\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
-$\Upsilon'$, which partial equivalence relation on the closed $\Upsilon'$-terms
-of $\tau$ that $\phi$ denotes. At this point, it should be remarked that this
-is very similar to the semantical explanation of typehood given in
+$\Upsilon'$, which partial equivalence relation on the closed $\tau$-sorted
+$\Upsilon'$-terms $\phi$ denotes. At this point, it should be remarked that
+this is very similar to the semantical explanation of typehood given in
 \cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
@@
 
-When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ is an
-\emph{global} refinement and may simply write $\GRefines{\phi}{\tau}$. By
-weakening, if $\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
+When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ \emph{globally}
+refines $\tau$ and may simply write $\GRefines{\phi}{\tau}$. By weakening, if
+$\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
 $\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.

### 2015-11-11 paper.tex
Switch to pdflatex; change fonts
Make colors conditional; add (conditional) remark explaining mode colors

@@
 
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
-\usepackage{agda} % this should come after ams*
 \usepackage{catchfilebetweentags}
-\usepackage{fontspec}
 \usepackage{geometry}
@@
 \usepackage{ifthen}
-\usepackage{lualatex-math}
-%\usepackage{mathtools}
+\usepackage{mathtools}
 \usepackage{proof}
@@
 \usepackage{tikz-cd}
-\usepackage{unicode-math}
 \usepackage{url}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage{xunicode}
 
-\definecolor{dark-gray}{gray}{0.20}
+\usepackage{pxfonts}
+\usepackage{eulervm}
+
+\newif\ifcolored
+\coloredtrue
 
@@
 ]{microtype}
-\setmainfont{Linux Libertine O}
-\setmathfont{Asana Math}
-
-% make lualatex play nice with agda
-\makeatletter
-\ifluatex%
-   \newwrite\CATCHPATCHOUT%
-\fi
-\def\CatchFBT@Fin@l#1[#2]{%
-   \begingroup%
-      \endlinechar\m@ne\makeatletter #2%
-      \ifluatex%
-         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch%
-         \immediate\write\CATCHPATCHOUT{\CatchFBT@tok{\the\CatchFBT@tok}}%
-         \immediate\closeout\CATCHPATCHOUT%
-         \@@input\jobname.catchpatch%
-      \else%
-      \scantokens\expandafter{%
-         \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
-      \fi%
-      \CatchFBT@IsAToken{#1}%
-         {\global#1\expandafter{\the\CatchFBT@tok}}%
-         {\xdef#1{\the\CatchFBT@tok}}%
-      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi%
-   \endgroup%
-}% \CatchFBT@Final
-\makeatother
 
@@
 \newtheorem{remark}[thm]{Remark}
+\newtheorem*{remark*}{Remark}
 \numberwithin{equation}{section}
 
+\ifcolored
   \def\IModeColorName{MidnightBlue}
   \def\OModeColorName{Maroon}
+\else
+  \def\IModeColorName{Black}
+  \def\OModeColorName{Black}
+\fi
 
@@
 
+\ifcolored
+  \bigskip
+  \begin{remark*}[Typographical Conventions]
+    In this paper, we hint at the \emph{modes} of judgments and assertions
+    \cite{harper:2016} using colors, marking inputs with
+    $\IMode{\texttt{blue}}$ and outputs with $\OMode{\texttt{red}}$. As a rule
+    of thumb, inputs are things which are supplied when checking the
+    correctness of a judgment, and outputs are things which are synthesized in
+    the process.
+  \end{remark*}
+\fi
+
 \section{Abstract Binding Trees and Symbols}
@@
 
-\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathsf{nil}}}
-\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathsf{snoc}}}
+\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
+\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
 
@@
 
+% TODO: define maps of refined contexts, then change weakening lemma to
+% monotonicity, which is stronger and more useful.
 \begin{thm}[Weakening]

### 2015-11-11 paper.tex
More progress

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage{leftidx}
 \usepackage{amsmath}
@@
 \newcommand\RefinesCtx[2]{\IMode{#1}\sqsubset^\star\OMode{#2}}
+\newcommand\EqRefines[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\sqsubset\OMode{#4}}
+
+\newcommand\HRefines[4]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}\sqsubset\OMode{#4}}
+\newcommand\HEqRefines[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\sqsubset\OMode{#5}}
+\newcommand\PRefinesCtx[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset^\star\OMode{#3}}
+
+\newcommand\RMemEq[4]{\IMode{#1}\parallel \IMode{#2}=\IMode{#3}\in \IMode{#4}}
+\newcommand\HRMemEq[5]{\IMode{#1}\parallel \IMode{#2} \vDash \IMode{#3}=\IMode{#4}\in \IMode{#5}}
 
@@
 framework, we will define the notion of behavioral refinement by propounding
-several judgments and their semantical explanations. We will require two
-judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$, which means that
-$\Phi$ is a context which refines the symbol context $\Upsilon$ (presupposing
-$\IsSymCtx{\Upsilon}$), and $\Refines{\Phi}{\phi}{\tau}$, which shall presuppose
+several judgments and their semantical explanations.
+
+\subsection{Parametric Refinement}
+
+We will require two judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$,
+which means that $\Phi$ is a symbol context which refines the symbol context
+$\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
+$\Refines{\Phi}{\phi}{\tau}$, which shall presuppose
 $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
@@
 explanation:
-\begin{definition}[Refinement]
+\begin{definition}[Parametric Refinement]
   To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
@@
   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
-  $\IMode{M}\sim_\phi\IMode{N}$ then $\IMode{N}\sim_\phi\IMode{M}$, and if
-  $\IMode{M}\sim_\phi\IMode{N}$ and $\IMode{N}\sim_\phi\IMode{O}$, then
-  $\IMode{M}\sim_\phi\IMode{O}$.
+  $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
+  $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
+  $\IMode{M}\sim^\rho_\phi\IMode{O}$.
 \end{definition}
@@
 
+\subsubsection{Equality of parametric refinement}
+
+Two refinements are equal when they denote the same PER. That is, will define
+the refinement equality judgment $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ as
+follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
+$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$:
+\begin{definition}
+  To know $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
+  $\Of{\rho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, and for any $M, N$
+  such that $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
+  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$,
+  $\IMode{M}\sim^\rho_\phi\IMode{N}$ if and only if
+  $\IMode{M}\sim^\rho_\psi\IMode{N}$.
+\end{definition}
+
+\subsection{Parametric Equality}
+
+The primary judgment concerning refinements is the parametric equality,
+$\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
+$\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
+$\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
+$M$ and $N$ are identified by $\phi$ at the identity renaming
+$\Of{1}{\IMode{\Upsilon}\hookrightarrow\OMode{\Upsilon}}$:
+\[
+  \infer{
+    \RMemEq{\Phi}{M}{N}{\phi}
+  }{
+    \IMode{M}\sim^1_\phi\IMode{N}
+  }
+\]
+
+\subsection{Functional Refinement and Equality}
+
+\newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
+\newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}
+
+Next, we define hypothetico-general (or \emph{functional}) refinement
+$\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
+simultaneously with parametric context refinement
+$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and functional equality
+$\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$. We will write
+$\HRefines{\Phi}{\Psi}{\phi}{\tau}$ as a shorthand for
+$\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.
+
+\newcommand\DefPRefinesCtxNil[1]{
+  \infer[\PRefinesCtxNil]{
+    \PRefinesCtx{#1}{\cdot}{\cdot}
+  }{
+  }
+}
+\newcommand\DefPRefinesCtxSnoc[6]{
+  \infer[\PRefinesCtxSnoc]{
+    \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
+  }{
+    \PRefinesCtx{#1}{#2}{#5} &
+    \HRefines{#1}{#2}{#4}{#6}
+  }
+}
+
+Parametric context refinement $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ shall be
+defined as follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$ and
+$\IsVarCtx{\Gamma}$:
+\begin{gather*}
+  \DefPRefinesCtxNil{\Phi}\qquad
+  \DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\phi}{\Gamma}{\tau}
+\end{gather*}
+
+Functional refinement $\HRefines{\Phi}{\Psi}{\phi}{\tau}$ presupposes
+$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
+induction on the evidence for its first presupposition.
+
+\paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
+\[
+  \infer{
+    \HRefines{\Phi}{\cdot}{\phi}{\tau}
+  }{
+    \Refines{\Phi}{\phi}{\tau}
+  }
+\]
+
+\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$.
+\begin{gather*}
+  \infer{
+    \HRefines{\Phi}{\Psi,x:\psi}{\phi}{\tau}
+  }{
+    \GenJ{y,z:\sigma}{
+      \HypJ{
+        \HEqRefines{\Phi}{\Psi}{\Subst{y}{x}{\phi}}{\Subst{z}{x}{\phi}}{\tau}
+      }{
+        \HRMemEq{\Phi}{\Psi}{y}{z}{\phi}
+      }
+    }
+  }
+\end{gather*}
+
+%% Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ is defined by
+%% induction on the evidence for its presupposition
+%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
+%% \begin{gather*}
+%%   \infer{
+%%     \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
+%%   }{
+%%     \RMemEq{\Phi}{M}{N}{\phi}
+%%   }\qquad
+%%   \infer{
+%%     \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
+%%   }{
+%%  }
+%%\end{gather*}
+
+%%
+%% \subsection{Parametric functional equality}
+%%
+%% Next, we will define the parametric hypothetico-general (or \emph{functional})
+%% equality
+%% $\IMode{\Phi}\parallel\IMode{\Psi}\vDash\IMode{M}=\IMode{N}\in\IMode{\phi}$
+%% simultaneously with variable context well-formedness,
+%% $\IMode{\Phi}\parallel\RefinesCtx{\Psi}{\Gamma}$.
+
 \ifdraft{}{

### 2015-11-11 paper.tex
Add ambient LF section; use cool todonotes package

@@
 \usepackage{url}
+\usepackage{todonotes}
 
@@
 
+\section{Abstract Binding Trees and Symbols}
+
+See \cite{sterling-morrison:2015} for the development of abstract binding trees
+with symbols. \todo[inline]{give a brief description of the framework, and present its
+rules.}
+
+\section{The Ambient Logical Framework}
+
 \ifcolored
-  \bigskip
-  \begin{remark*}[Typographical Conventions]
   In this paper, we hint at the \emph{modes} of judgments and assertions
@@
   the process.
-  \end{remark*}
 \fi
 
-\section{Abstract Binding Trees and Symbols}
-
-See \cite{sterling-morrison:2015} for the development of abstract binding trees
-with symbols.
-
-TODO: give a brief description of the framework, and present its rules.
+\todo[inline]{explain general and hypothetical judgment with respect to ABTs (probably
+following Bob's book).}
 
@@
 
-% TODO: define maps of refined contexts, then change weakening lemma to
-% monotonicity, which is stronger and more useful.
+\todo[inline]{%
+  define maps of refined contexts, then change weakening lemma to
+  monotonicity, which is stronger and more useful.
+}
 \begin{thm}[Weakening]

### 2015-11-11 paper.tex
corrections

@@
 
-Next, we define hypothetico-general (or \emph{functional}) refinement
+Next, we define functional refinement
 $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
@@
 $\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.
+
+\todo[inline]{%
+  Properly express the \emph{syntactic} presuppositions of the sequent
+  judgments so that the use of hypothetical and general judgment is
+  well-formed.
+}
 
@@
 
-Functional refinement $\HRefines{\Phi}{\Psi}{\phi}{\tau}$ presupposes
+Functional refinement $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ presupposes
 $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
@@
   \infer{
-    \HRefines{\Phi}{\cdot}{\phi}{\tau}
+    \HEqRefines{\Phi}{\cdot}{\phi}{\psi}{\tau}
   }{
-    \Refines{\Phi}{\phi}{\tau}
+    \EqRefines{\Phi}{\phi}{\psi}{\tau}
   }
@@
 
-\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$.
+\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
 \begin{gather*}
   \infer{
-    \HRefines{\Phi}{\Psi,x:\psi}{\phi}{\tau}
+    \HEqRefines{\Phi}{\Psi,x:\chi}{\phi}{\psi}{\tau}
   }{
@@
       \HypJ{
-        \HEqRefines{\Phi}{\Psi}{\Subst{y}{x}{\phi}}{\Subst{z}{x}{\phi}}{\tau}
+        \HEqRefines{\Phi}{\Psi}{\Subst{y}{x}{\phi}}{\Subst{z}{x}{\psi}}{\tau}
       }{
-        \HRMemEq{\Phi}{\Psi}{y}{z}{\phi}
+        \HRMemEq{\Phi}{\Psi}{y}{z}{\chi}
       }

### 2015-11-11 paper.tex
Explain the hypothetical and general judgments / ambient LF

@@
 
-\newcommand\IsAbt[5]{%
-  \IMode{#1}\triangleright%
-  \IMode{#2}\parallel%
-  \IMode{#3}\vdash%
-  \IMode{#4}:\OMode{#5}%
+\newcommand\IsAbtUnmoded[5]{
+  #1\triangleright%
+  #2\parallel%
+  #3\vdash%
+  #4:\OMode{#5}%
 }
-\newcommand\IsBTm[5]{%
-  \IMode{#1}\triangleright%
-  \IMode{#2}\parallel%
-  \IMode{#3}\vdash%
-  \IMode{#4}:\IMode{#5}%
-}
+\newcommand\IsAbt[5]{\IsAbtUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}{\IMode{#4}}{\OMode{#5}}}
+\newcommand\IsBTm[5]{\IsAbtUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}{\IMode{#4}}{\IMode{#5}}}
 
@@
 \newcommand\HRMemEq[5]{\IMode{#1}\parallel \IMode{#2} \vDash \IMode{#3}=\IMode{#4}\in \IMode{#5}}
+
+\newcommand\Pow[1]{\wp\left(#1\right)}
 
@@
 
-\todo[inline]{explain general and hypothetical judgment with respect to ABTs (probably
-following Bob's book).}
+For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we may
+deploy a generic logical framework, where we interpret a judgment as the set of
+its derivations.
+
+\begin{definition}[General Judgment]
+  For a metavariable context $\IsMetaCtx{\Omega}$, let
+  \[\Define{[\Omega]}{\coprod_{\MV{m}:v}
+  \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}}\] Then for any
+  $\Of{\mathcal{J}}{[\Omega]\to\Sets}$ we define the \emph{general judgment}
+  $\GenJ{\Omega}{\mathcal{J}}$ as the section
+  $\prod_{\omega\in[\Omega]}{\mathcal{J}(\omega)}$. Recall that metavariables
+  have \emph{valences} rather than sorts; as such, the general judgment allows
+  us to quantify over terms of higher type. However, for clarity we will often
+  write $\GenJ{\MV{m}:\sigma}{\mathcal{J}}$ rather than
+  $\GenJ{\MV{m}:\MkValence{\cdot}{\cdot}{\sigma}}{\mathcal{J}}$; likewise, in
+  the body of $\mathcal{J}$ we will simply write $\MV{m}$ rather than the more
+  proper $\MApp{\MV{m}}{\cdot}{\cdot}$.
+\end{definition}
+
+\begin{definition}[Hypothetical Judgment]
+  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we have the
+  hypothetical judgment
+  $\Define{\HypJ{\mathcal{J}_2}{\mathcal{J}_1}}{\mathcal{J}_1\to\mathcal{J}_2}$.
+  In other words, evidence for a hypothetical judgment is an effective
+  transformation of evidence for the antecedent into evidence for the
+  consequent. Note that this is the \emph{semantic} hypothetical judgment,
+  which expresses admissibilities, rather than the \emph{syntactic}
+  hypothetical judgment, which expresses derivabilities. We will not have any
+  need for the latter.
+\end{definition}
 

### 2015-11-11 paper.tex
fix refs

@@
   \[\Define{[\Omega]}{\coprod_{\MV{m}:v}
-  \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}}\] Then for any
+  \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}}\] Then for any family of judgments
   $\Of{\mathcal{J}}{[\Omega]\to\Sets}$ we define the \emph{general judgment}
@@
   which expresses admissibilities, rather than the \emph{syntactic}
-  hypothetical judgment, which expresses derivabilities. We will not have any
-  need for the latter.
+  hypothetical judgment, which expresses derivabilities \cite{harper:2016}. We
+  will not have any need for the latter.
 \end{definition}

### 2015-11-12 paper.tex
Revamp the ambient logical framework in terms of presheaves

--- 2015-11-11 paper.tex
@@
 
-\newcommand\Hom[3]{#1\left[#2,#3\right]}
+\newcommand\Hom[3]{#1\!\left[#2,#3\right]}
 
@@
 
-\todo[inline]{explain general and hypothetical judgment with respect to ABTs (probably
-following Bob's book).}
+For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ and a worlds
+category $\mathcal{W}$, we may deploy a general logical framework in which the
+meaning of a judgment is an object of the presheaf category
+$\Define{\mathbb{J}}{\Sets^\mathcal{W}}$ (that is, a judgment is identified
+with the intensional set of its derivations).
+
+\begin{definition}[General Judgment]
+  For a metavariable context $\IsMetaCtx{\Omega}$, we can define the set
+  of its substitutions as the following product:
+  \[
+    \Define{[\Omega]}{
+      \prod_{\MV{m}:v} \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
+    }
+  \]
+  Let $\Discrete{[\Omega]}$ be the discrete category on $[\Omega]$. Then, for
+  any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\Discrete{[\Omega]}}}$, we can
+  define the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows:
+  \[
+    \textstyle
+    \Define{
+      \left(\GenJ{\Omega}{\mathcal{J}}\right)(w)
+    }{
+      \Hom{\mathbb{J}}{
+        \Yoneda{w}
+      }{
+        \varinjlim_{\Discrete{[\Omega]}}\mathcal{J}
+      }
+    }\qquad
+    (\Member{w}{\mathcal{W}})
+  \]
+
+  Intuitively, the general judgment $\GenJ{\Omega}{\mathcal{J}}$ is evident
+  when the family of judgments $\mathcal{J}$ shall forevermore have a section
+  (i.e. for any substitution of $\Member{\omega}{[\Omega]}$,
+  $\mathcal{J}(\omega)$ shall be evident). Recall that metavariables have
+  \emph{valences} rather than sorts; as such, the general judgment allows us to
+  quantify over terms of higher type.  However, for clarity we will often write
+  $\GenJ{\MV{m}:\sigma}{\mathcal{J}}$ rather than
+  $\GenJ{\MV{m}:\MkValence{\cdot}{\cdot}{\sigma}}{\mathcal{J}}$; likewise, in
+  the body of $\mathcal{J}$ we will simply write $\MV{m}$ rather than the more
+  proper $\MApp{\MV{m}}{\cdot}{\cdot}$.
+
+\end{definition}
+
+\begin{definition}[Hypothetical Judgment]
+
+  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
+  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$, which expresses
+  the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
+  \begin{align*}
+    \ADefine{
+      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(w)
+    }{
+      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(w)
+    }\\
+    &\cong
+    \OMode{
+      \Hom{\mathbb{J}}{
+        \Yoneda{w}\times\mathcal{J}_1
+      }{
+        \mathcal{J}_2
+      }
+    }
+  \end{align*}
+
+  Intuitively, the hypothetical judgment
+  $\Define{\HypJ{\mathcal{J}_2}{\mathcal{J}_1}}{\mathcal{J}_1\to\mathcal{J}_2}$
+  is evident when there shall forevermore be an effective transformation of the
+  evidence into the evidence of the consequent. It is important to keep in mind
+  that this is a \emph{semantic} consequence and therefore expresses
+  admissibilities, in contrast to the the \emph{logical} consequence which
+  expresses derivabilities \cite{harper:2016}. We will not have any need for
+  the latter.
+\end{definition}
 

### 2015-11-12 paper.tex
typos

@@
   when the family of judgments $\mathcal{J}$ shall forevermore have a section
-  (i.e. for any substitution of $\Member{\omega}{[\Omega]}$,
+  (i.e. for any substitution $\Member{\omega}{[\Omega]}$,
   $\mathcal{J}(\omega)$ shall be evident). Recall that metavariables have
@@
 
-  Intuitively, the hypothetical judgment
-  $\Define{\HypJ{\mathcal{J}_2}{\mathcal{J}_1}}{\mathcal{J}_1\to\mathcal{J}_2}$
-  is evident when there shall forevermore be an effective transformation of the
-  evidence into the evidence of the consequent. It is important to keep in mind
-  that this is a \emph{semantic} consequence and therefore expresses
-  admissibilities, in contrast to the the \emph{logical} consequence which
-  expresses derivabilities \cite{harper:2016}. We will not have any need for
-  the latter.
+  In other words, the hypothetical judgment
+  $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
+  be an effective transformation of evidence for the antecedent into evidence
+  for the consequent. It is important to keep in mind that this is a
+  \emph{semantic} consequence and therefore expresses admissibilities, in
+  contrast to the the \emph{logical} consequence which expresses derivabilities
+  \cite{harper:2016}.  We will not have any need for the latter.
 \end{definition}

### 2015-11-12 paper.tex
remark on implicit definitions of judgments
remove (w \in W)
typos

@@
 
-For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ and a worlds
-category $\mathcal{W}$, we may deploy a general logical framework in which the
+For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ and small worlds
+category $\mathcal{W}$, we may deploy a generic logical framework in which the
 meaning of a judgment is an object of the presheaf category
@@
       }
-    }\qquad
-    (\Member{w}{\mathcal{W}})
+    }
   \]
@@
 \end{definition}
+
+After this section, we will define a judgment either by rules, or by an
+informal semantical explanation, in both cases leaving its intensional
+character implicit. A judgment will said to be \emph{correct} or \emph{valid}
+in case it is globally inhabited.
 

### 2015-11-12 paper.tex
notation

@@
   \[
-    \Define{[\Omega]}{
+    \Define{\square\Omega}{
       \prod_{\MV{m}:v} \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
@@
   \]
-  Let $\Discrete{[\Omega]}$ be the discrete category on $[\Omega]$. Then, for
-  any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\Discrete{[\Omega]}}}$, we can
+  Let $\Discrete{\square\Omega}$ be the discrete category on $\square\Omega$. Then, for
+  any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\Discrete{\square\Omega}}}$, we can
   define the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows:
@@
       }{
-        \varinjlim_{\Discrete{[\Omega]}}\mathcal{J}
+        \varinjlim_{\Discrete{\square\Omega}}\mathcal{J}
       }
@@
   when the family of judgments $\mathcal{J}$ shall forevermore have a section
-  (i.e. for any substitution $\Member{\omega}{[\Omega]}$,
+  (i.e. for any substitution $\Member{\omega}{\square\Omega}$,
   $\mathcal{J}(\omega)$ shall be evident). Recall that metavariables have

### 2015-11-12 paper.tex
"the the"
missing word

@@
   \emph{semantic} consequence and therefore expresses admissibilities, in
-  contrast to the the \emph{logical} consequence which expresses derivabilities
+  contrast to the \emph{logical} consequence which expresses derivabilities
   \cite{harper:2016}.  We will not have any need for the latter.
@@
 informal semantical explanation, in both cases leaving its intensional
-character implicit. A judgment will said to be \emph{correct} or \emph{valid}
-in case it is globally inhabited.
+character implicit. A judgment is said to be \emph{correct} or \emph{valid}
+just in case it is globally inhabited.
 

### 2015-11-12 paper.tex
Fix \square\Omega definition

@@
 category $\mathcal{W}$, we may deploy a generic logical framework in which the
-meaning of a judgment is an object of the presheaf category
+meaning of a judgment is an object in the presheaf category
 $\Define{\mathbb{J}}{\Sets^\mathcal{W}}$ (that is, a judgment is identified
@@
     \Define{\square\Omega}{
-      \prod_{\MV{m}:v} \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
+      \prod_{(\MV{m}:v)\in\Omega} \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
     }

### 2015-11-12 paper.tex
Fix all chktex warnings

@@
 
-\newif\ifcolored
-\coloredtrue
+\newif\ifcolored%
+\coloredtrue%
 
@@
 
-\ifcolored
+\ifcolored%
   \def\IModeColorName{MidnightBlue}
@@
 
-\newcommand\Refines[3]{\IMode{#1}\parallel \IMode{#2} \sqsubset \OMode{#3}}
+\newcommand\Refines[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset\OMode{#3}}
 \newcommand\GRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}
@@
 
-\newcommand\RMemEq[4]{\IMode{#1}\parallel \IMode{#2}=\IMode{#3}\in \IMode{#4}}
-\newcommand\HRMemEq[5]{\IMode{#1}\parallel \IMode{#2} \vDash \IMode{#3}=\IMode{#4}\in \IMode{#5}}
+\newcommand\RMemEq[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\in\IMode{#4}}
+\newcommand\HRMemEq[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\in\IMode{#5}}
 
@@
 decidable. In hindsight, we may observe that this is the sort of type theory
-which Martin-L\"of first proposed in 1972 \cite{martin-lof:1972}, even if we
+which Martin-L\"of first proposed in 1972~\cite{martin-lof:1972}, even if we
 will admit that this was not the intention at the time. A model for such a type
@@
 \emph{semantic type theory}. The most widely known development of this program
-is Martin-L\"of's 1979 ``extensional'' type theory \cite{martin-lof:1979,
+is Martin-L\"of's 1979 ``extensional'' type theory~\cite{martin-lof:1979,
 martin-lof:1984}, but we must give priority to Dana Scott for inventing this
 line of research in 1970 with his prophetic report, \emph{Constructive
-Validity} \cite{scott:1970}. More recently, behavioral type theory has been
-developed much further in the Nuprl family \cite{constable:1986} of proof
-assistants, including MetaPRL \cite{hickey:2003} and JonPRL \cite{jonprl:2015}.
+Validity}~\cite{scott:1970}. More recently, behavioral type theory has been
+developed much further in the Nuprl family~\cite{constable:1986} of proof
+assistants, including MetaPRL~\cite{hickey:2003} and JonPRL~\cite{jonprl:2015}.
 
@@
 The types of the semantic theory can then be said to \emph{refine} the types of
-the syntactic theory \cite{harper-davies:2014, harper-duff:2015, harper:2016},
+the syntactic theory~\cite{harper-davies:2014, harper-duff:2015, harper:2016},
 both by placing restrictions on membership and by coarsening equivalence. Thus
@@
 behavioral refinements over multi-sorted abstract binding trees, a simple
-formal type theory \cite{harper:2016, sterling-morrison:2015}; this hybrid
+formal type theory~\cite{harper:2016, sterling-morrison:2015}; this hybrid
 system allows the deployment of a Nuprl-style type theory over any signature of
@@
 
-See \cite{sterling-morrison:2015} for the development of abstract binding trees
+See~\cite{sterling-morrison:2015} for the development of abstract binding trees
 with symbols. \todo[inline]{give a brief description of the framework, and present its
@@
 
-\ifcolored
-  In this paper, we hint at the \emph{modes} of judgments and assertions
-  \cite{harper:2016} using colors, marking inputs with
-  $\IMode{\texttt{blue}}$ and outputs with $\OMode{\texttt{red}}$. As a rule
-  of thumb, inputs are things which are supplied when checking the
-  correctness of a judgment, and outputs are things which are synthesized in
-  the process.
-\fi
+\ifcolored%
+In this paper, we hint at the \emph{modes} of judgments and
+assertions~\cite{harper:2016} using colors, marking inputs with
+$\IMode{\texttt{blue}}$ and outputs with $\OMode{\texttt{red}}$. As a rule of
+thumb, inputs are things which are supplied when checking the correctness of a
+judgment, and outputs are things which are synthesized in the process. \fi
 
@@
   when the family of judgments $\mathcal{J}$ shall forevermore have a section
-  (i.e. for any substitution $\Member{\omega}{\square\Omega}$,
+  (i.e.\ for any substitution $\Member{\omega}{\square\Omega}$,
   $\mathcal{J}(\omega)$ shall be evident). Recall that metavariables have
@@
   \emph{semantic} consequence and therefore expresses admissibilities, in
-  contrast to the \emph{logical} consequence which expresses derivabilities
-  \cite{harper:2016}.  We will not have any need for the latter.
+  contrast to the \emph{logical} consequence which expresses
+  derivabilities~\cite{harper:2016}. We will not have any need for the latter.
 \end{definition}
@@
 $\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
-$\Refines{\Phi}{\phi}{\tau}$, which shall presuppose
+$\Refines{\Phi}{\phi}{\tau}$ (which shall presuppose
 $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
@@
 
-Two refinements are equal when they denote the same PER. That is, will define
+Two refinements are equal when they denote the same PER.\@That is, will define
 the refinement equality judgment $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ as

### 2015-11-12 paper.tex
Activate kerning and spacing for microtype
Set up blackboard bold for digits
Express colimit as Lan
Rearrange the LF to support symbol renaming using fibration
@freebroccolo; does it look like this will work? If so, do you think
there is any nicer way to do it?

@@
 
-\usepackage{pxfonts}
-\usepackage{eulervm}
-
 \newif\ifcolored%
@@
 
-% fonts
+%% fonts
+\usepackage{pxfonts}
+\usepackage{eulervm}
 \usepackage[
   activate={true,nocompatibility},
+  kerning=true,
+  spacing=true,
   tracking=true
 ]{microtype}
+\DeclareSymbolFont{bbold}{U}{bbold}{m}{n}
+\DeclareSymbolFontAlphabet{\mathbbold}{bbold}
 
@@
 
-For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ and small worlds
-category $\mathcal{W}$, we may deploy a generic logical framework in which the
-meaning of a judgment is an object in the presheaf category
-$\Define{\mathbb{J}}{\Sets^\mathcal{W}}$ (that is, a judgment is identified
-with the intensional set of its derivations).
+Then, for any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
+deploy a generic logical framework. Recall from \cite{sterling-morrison:2015}
+that $\SCtx$ is the category of symbol-contexts and their injective,
+sort-preserving renamings. Then, fix a small category $\mathcal{W}$ of worlds
+along with a map $\Of{\pi_\mathcal{W}}{\mathcal{W}\to\SCtx}$ such that
+$\pi_\mathcal{W}$ is a Grothendieck opfibration. Then, a \emph{judgment} shall
+be interpreted as an object in the presheaf category
+$\Define{\mathbb{J}}{\Sets^\mathcal{W}}$; in other words, a judgment is
+identified with the intensional set of its renaming-sensitive derivations.
 
 \begin{definition}[General Judgment]
-  For a metavariable context $\IsMetaCtx{\Omega}$, we can define the set
-  of its substitutions as the following product:
+  For a metavariable context $\IsMetaCtx{\Omega}$, we can define the presheaf
+  of its substitutions $\Of{\square\Omega}{\Sets^\SCtx}$ as follows:
   \[
-    \Define{\square\Omega}{
-      \prod_{(\MV{m}:v)\in\Omega} \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
+    \Define{{\square\Omega}\,(\Upsilon)}{
+      \prod_{(\MV{m}:v)\in\Omega}
+        \MkSet{E\mid\IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{E}{v}}
     }
   \]
-  Let $\Discrete{\square\Omega}$ be the discrete category on $\square\Omega$. Then, for
-  any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\Discrete{\square\Omega}}}$, we can
-  define the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows:
+
+  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Omega(\Upsilon)}}$, we can define
+  the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows, where
+  $\Of{\Yoneda{-}}{\mathcal{W}\to\mathbb{J}}$ is the Yoneda Embedding:
    \[
@@
        }{
-        \varinjlim_{\Discrete{\square\Omega}}\mathcal{J}
+         Lan_{\mathbbold{1}}{\mathcal{J}}
        }
@@
    when the family of judgments $\mathcal{J}$ shall forevermore have a section
-  (i.e.\ for any substitution $\Member{\omega}{\square\Omega}$,
+   (i.e. for any substitution $\Member{\omega}{\square\Omega(\Upsilon)}$,
    $\mathcal{J}(\omega)$ shall be evident). Recall that metavariables have

### 2015-11-12 paper.tex
Clean up previous work

@@
 \newcommand\GenJ[2]{\vert_{\IMode{#1}}\; #2}
+\newcommand\HPGenJ[4]{\vert^{\IMode{#1}\parallel\IMode{#2}}_{\IMode{#3}}\; #4}
 
@@
 \newcommand\MV[1]{\mathfrak{#1}}
-\newcommand\MApp[3]{#1\{#2\}(#3)}
+\newcommand\MApp[3]{%
+  #1
+  \ifthenelse{\equal{#2}{}}{}{%
+    \{#2\}%
+  }%
+  \ifthenelse{\equal{#3}{}}{}{%
+    [#3]%
+  }
+}
 \newcommand\App[2]{#1(#2)}
@@
 \begin{definition}[General Judgment]
-  For a metavariable context $\IsMetaCtx{\Omega}$, we can define the presheaf
-  of its substitutions $\Of{\square\Omega}{\Sets^\SCtx}$ as follows:
+  For a metavariable context $\IsMetaCtx{\Omega}$, we can define the set
+  of its substitutions as follows:
   \[
-    \Define{{\square\Omega}\,(\Upsilon)}{
+    \Define{{\square\Omega}}{
       \prod_{(\MV{m}:v)\in\Omega}
@@
 
-  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Omega(\Upsilon)}}$, we can define
+  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Omega}}$, we can define
   the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows, where
@@
    when the family of judgments $\mathcal{J}$ shall forevermore have a section
-   (i.e. for any substitution $\Member{\omega}{\square\Omega(\Upsilon)}$,
+   (i.e. for any substitution $\Member{\omega}{\square\Omega}$,
    $\mathcal{J}(\omega)$ shall be evident). Recall that metavariables have
    \emph{valences} rather than sorts; as such, the general judgment allows us to
-   quantify over terms of higher type.  However, for clarity we will often write
+   quantify over terms of higher type.
+
+   \paragraph{Notation}
+
+   We will write $\HPGenJ{\Upsilon}{\Gamma}{\Omega}{\mathcal{J}}$ for
+   $\GenJ{\Omega'}{\mathcal{J}}$, where
+   $\Lookup{\Omega'}{\MV{m}}{\MkValence{\Upsilon,\vec{\sigma}}{\Gamma,\vec{\tau}}{\tau}}$
+   if $\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$.
+   Additionally, for clarity we will often write
    $\GenJ{\MV{m}:\sigma}{\mathcal{J}}$ rather than
    $\GenJ{\MV{m}:\MkValence{\cdot}{\cdot}{\sigma}}{\mathcal{J}}$; likewise, in
-   the body of $\mathcal{J}$ we will simply write $\MV{m}$ rather than the more
-   proper $\MApp{\MV{m}}{\cdot}{\cdot}$.
+   the body of $\mathcal{J}$, when there is a contextually salient
+   $\vec{u}$, we will simply write $\MV{m}$ rather than the more proper
+   $\MApp{\MV{m}}{\vec{u}}{\cdot}$.
 
@@
 
-\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
+\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$,
+where $\RefinesCtx{\Phi}{\Upsilon}$.
+
 \begin{gather*}
@@
   }{
-    \GenJ{y,z:\sigma}{
+    \HPGenJ{\Upsilon}{\Gamma}{y,z:\sigma}{
       \HypJ{
-        \HEqRefines{\Phi}{\Psi}{\Subst{y}{x}{\phi}}{\Subst{z}{x}{\psi}}{\tau}
+        \HEqRefines{\Phi}{\Psi}{
+          \Subst{y}{x}{\phi}
+        }{
+          \Subst{z}{x}{\psi}
+        }{\tau}
       }{

### 2015-11-12 paper.tex
fix up

@@
    $\GenJ{\MV{m}:\MkValence{\cdot}{\cdot}{\sigma}}{\mathcal{J}}$; likewise, in
-   the body of $\mathcal{J}$, when there is a contextually salient
-   $\vec{u}$, we will simply write $\MV{m}$ rather than the more proper
-   $\MApp{\MV{m}}{\vec{u}}{\cdot}$.
+   the body of $\mathcal{J}$, when there are contextually salient
+   $\vec{u}$, $\vec{x}$, we will simply write $\MV{m}$ rather than the more proper
+   $\MApp{\MV{m}}{\vec{u}}{\vec{x}}$.
 

### 2015-11-12 paper.tex
Formalize equality of parametric refinements

@@
   To know $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
-  $\Of{\rho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, and for any $M, N$
-  such that $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
-  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$,
-  $\IMode{M}\sim^\rho_\phi\IMode{N}$ if and only if
-  $\IMode{M}\sim^\rho_\psi\IMode{N}$.
+  $\Of{\rho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, the following judgments:
+  \begin{gather*}
+    \GenJ{
+      \MV{m},\MV{n}:\tau
+    }{
+      \HypJ{
+        \IMode{M}\sim^\rho_\phi\IMode{N}
+      }{
+        \IMode{M}\sim^\rho_\psi\IMode{N}
+      }
+    }\\
+    \GenJ{
+      \MV{m},\MV{n}:\tau
+    }{
+      \HypJ{
+        \IMode{M}\sim^\rho_\psi\IMode{N}
+      }{
+        \IMode{M}\sim^\rho_\phi\IMode{N}
+      }
+    }
+  \end{gather*}
+
+ %% and for any $M, N$
+ %% such that $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
+ %% $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$,
+ %% $\IMode{M}\sim^\rho_\phi\IMode{N}$ if and only if
+ %% $\IMode{M}\sim^\rho_\psi\IMode{N}$.
 \end{definition}
@@
   }{
-    \HPGenJ{\Upsilon}{\Gamma}{y,z:\sigma}{
+    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
       \HypJ{
         \HEqRefines{\Phi}{\Psi}{
-          \Subst{y}{x}{\phi}
+          \Subst{\MV{m}}{x}{\phi}
         }{
-          \Subst{z}{x}{\psi}
+          \Subst{\MV{n}}{x}{\psi}
         }{\tau}
       }{
-        \HRMemEq{\Phi}{\Psi}{y}{z}{\chi}
+        \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\chi}
       }

### 2015-11-12 paper.tex
Define functional equality

@@
 
-%% Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ is defined by
-%% induction on the evidence for its presupposition
-%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
-%% \begin{gather*}
-%%   \infer{
-%%     \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
-%%   }{
-%%     \RMemEq{\Phi}{M}{N}{\phi}
-%%   }\qquad
-%%   \infer{
-%%     \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
-%%   }{
-%%  }
-%%\end{gather*}
+Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
+$\RefinesCtx{\Phi}{\Upsilon}$, $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$,
+$\HRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
+and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
+induction on the evidence for the second presupposition
+$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
 
-%%
-%% \subsection{Parametric functional equality}
-%%
-%% Next, we will define the parametric hypothetico-general (or \emph{functional})
-%% equality
-%% $\IMode{\Phi}\parallel\IMode{\Psi}\vDash\IMode{M}=\IMode{N}\in\IMode{\phi}$
-%% simultaneously with variable context well-formedness,
-%% $\IMode{\Phi}\parallel\RefinesCtx{\Psi}{\Gamma}$.
+\paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
+\[
+  \infer{
+    \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
+  }{
+    \RMemEq{\Phi}{M}{N}{\phi}
+  }
+\]
+
+\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
+where $\RefinesCtx{\Phi}{\Upsilon}$.
+\[
+  \infer{
+    \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
+  }{
+    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
+      \HypJ{
+        \HRMemEq{\Phi}{\Psi}{
+          \Subst{\MV{m}}{x}{\phi}
+        }{
+          \Subst{\MV{n}}{x}{\psi}
+        }{\phi}
+      }{
+        \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\psi}
+      }
+    }
+  }
+\]
 

### 2015-11-12 paper.tex
Typo
Remove resolved TODO
Add a header
I think we need the right kan extension here (?)
@freebroccolo Is this right? The intention was to take the limit of the
diagram (that is, the product of the family).
fix some typos

@@
 along with a map $\Of{\pi_\mathcal{W}}{\mathcal{W}\to\SCtx}$ such that
-$\pi_\mathcal{W}$ is a Grothendieck opfibration. Then, a \emph{judgment} shall
+$\pi_\mathcal{W}$ is a Grothendieck fibration. Then, a \emph{judgment} shall
 be interpreted as an object in the presheaf category
-$\Define{\mathbb{J}}{\Sets^\mathcal{W}}$; in other words, a judgment is
-identified with the intensional set of its renaming-sensitive derivations.
+$\Define{\mathbb{J}}{\Sets^{\OpCat{\mathcal{W}}}}$; in other words, a judgment
+is identified with the intensional set of its renaming-sensitive derivations.
 
@@
       \prod_{(\MV{m}:v)\in\Omega}
-        \MkSet{E\mid\IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{E}{v}}
+        \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
     }
@@
        }{
-         Lan_{\mathbbold{1}}{\mathcal{J}}
+         Ran_{\mathbbold{1}}{\mathcal{J}}
        }
@@
 
-\subsection{Functional Refinement and Equality}
+\subsection{Functional Refinement}
 
@@
 $\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.
-
-\todo[inline]{%
-  Properly express the \emph{syntactic} presuppositions of the sequent
-  judgments so that the use of hypothetical and general judgment is
-  well-formed.
-}
 
@@
 
+\subsection{Functional Equality}
+
 Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
@@
       }{
-        \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\psi}
+        \HRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
       }

### 2015-11-12 paper.tex
Add some spacing for Hom
Include pointful definitions
Fix up trailing space for Hom
Simplify definitions
Remove annotation
Fix inverted relation
Clearer name for the substitution of a mctx
"more recently" ⇒ "since the 1980s"

@@
 
-\newcommand\Hom[3]{#1\!\left[#2,#3\right]}
+\newcommand\Hom[3]{#1\!\left[#2,\,#3\right]\!}
 
@@
 line of research in 1970 with his prophetic report, \emph{Constructive
-Validity}~\cite{scott:1970}. More recently, behavioral type theory has been
+Validity}~\cite{scott:1970}. Since the 1980s, behavioral type theory has been
 developed much further in the Nuprl family~\cite{constable:1986} of proof
@@
   $\Of{\Yoneda{-}}{\mathcal{W}\to\mathbb{J}}$ is the Yoneda Embedding:
-   \[
-     \textstyle
-     \Define{
+  \begin{align*}
+    \ADefine{
       \left(\GenJ{\Omega}{\mathcal{J}}\right)(w)
@@
       }
+    }\\
+    &\cong
+    \OMode{
+      \int_{(w',\,\vec{E})}{{w'}\succeq{w}\Rightarrow{\mathcal{J}(\vec{E})(w')}}
     }
-   \]
+  \end{align*}
 
@@
    when the family of judgments $\mathcal{J}$ shall forevermore have a section
-   (i.e. for any substitution $\Member{\omega}{\square\Omega}$,
-   $\mathcal{J}(\omega)$ shall be evident). Recall that metavariables have
+   (i.e. for any substitution $\Member{\vec{E}}{\square\Omega}$,
+   $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have
    \emph{valences} rather than sorts; as such, the general judgment allows us to
@@
       }
+    }\\
+    &\cong
+    \OMode{
+      \int_{w'}{{{w'}\succeq{w}}\Rightarrow{{{\mathcal{J}}_{2}(w')}^{{\mathcal{J}}_{1}(w')}}}
     }

### 2015-11-12 paper.tex
reword sentence about functionality

@@
 types, in contrast to the state of affairs in formal type theory where
-functionality is a metatheorem which may or may not hold depending on the
-particular equational theory which has been imposed. Furthermore, models for
-behavioral type theory interpret the types as partial equivalence relations on
-only closed terms, and the meaning of the hypothetical judgment is defined
-separately and uniformly in the logical relations style.
+functionality is a metatheorem which must be shown to obtain, based on the
+(somewhat arbitrary) equational theory which has been imposed. Furthermore,
+models for behavioral type theory interpret the types as partial equivalence
+relations on only closed terms, and the meaning of the hypothetical judgment is
+defined separately and uniformly in the logical relations style.
 

### 2015-11-12 paper.tex
Remove superfluous "Then"
Add some citations; reorder definitions

@@
 
-Then, for any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
+For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
 deploy a generic logical framework. Recall from \cite{sterling-morrison:2015}
@@
 along with a map $\Of{\pi_\mathcal{W}}{\mathcal{W}\to\SCtx}$ such that
-$\pi_\mathcal{W}$ is a Grothendieck fibration. Then, a \emph{judgment} shall
-be interpreted as an object in the presheaf category
+$\pi_\mathcal{W}$ is a Grothendieck fibration~\cite{jacobs:2001}. Then, a
+\emph{judgment} shall be interpreted as an object in the presheaf category
 $\Define{\mathbb{J}}{\Sets^{\OpCat{\mathcal{W}}}}$; in other words, a judgment
 is identified with the intensional set of its renaming-sensitive derivations.
+
+\begin{definition}[Hypothetical Judgment]
+  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
+  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ using the
+  exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses the semantic consequence
+  of $\mathcal{J}_2$ from $\mathcal{J}_1$:
+  \begin{align*}
+    \ADefine{
+      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(w)
+    }{
+      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(w)
+    }\\
+    &\cong
+    \OMode{
+      \Hom{\mathbb{J}}{
+        \Yoneda{w}\times\mathcal{J}_1
+      }{
+        \mathcal{J}_2
+      }
+    }\\
+    &\cong
+    \OMode{
+      \int_{w'}{{{w'}\succeq{w}}\Rightarrow{{{\mathcal{J}}_{2}(w')}^{{\mathcal{J}}_{1}(w')}}}
+    }
+  \end{align*}
+
+  In other words, the hypothetical judgment
+  $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
+  be an effective transformation of evidence for the antecedent into evidence
+  for the consequent. It is important to keep in mind that this is a
+  \emph{semantic} consequence and therefore expresses admissibilities, in
+  contrast to the \emph{logical} consequence which expresses
+  derivabilities~\cite{harper:2016}. We will not have any need for the latter.
+\end{definition}
 
@@
 
-\end{definition}
-
-\begin{definition}[Hypothetical Judgment]
-
-  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
-  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$, which expresses
-  the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
-  \begin{align*}
-    \ADefine{
-      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(w)
-    }{
-      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(w)
-    }\\
-    &\cong
-    \OMode{
-      \Hom{\mathbb{J}}{
-        \Yoneda{w}\times\mathcal{J}_1
-      }{
-        \mathcal{J}_2
-      }
-    }\\
-    &\cong
-    \OMode{
-      \int_{w'}{{{w'}\succeq{w}}\Rightarrow{{{\mathcal{J}}_{2}(w')}^{{\mathcal{J}}_{1}(w')}}}
-    }
-  \end{align*}
-
-  In other words, the hypothetical judgment
-  $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
-  be an effective transformation of evidence for the antecedent into evidence
-  for the consequent. It is important to keep in mind that this is a
-  \emph{semantic} consequence and therefore expresses admissibilities, in
-  contrast to the \emph{logical} consequence which expresses
-  derivabilities~\cite{harper:2016}. We will not have any need for the latter.
 \end{definition}

### 2015-11-13 paper.tex
Also cite Categories for the Working Mathematician
since I used this to figure out wtf kan extensions are
Add diagram for Grothendieck fibration

@@
 
-For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
-deploy a generic logical framework. Recall from \cite{sterling-morrison:2015}
-that $\SCtx$ is the category of symbol-contexts and their injective,
-sort-preserving renamings. Then, fix a small category $\mathcal{W}$ of worlds
-along with a map $\Of{\pi_\mathcal{W}}{\mathcal{W}\to\SCtx}$ such that
-$\pi_\mathcal{W}$ is a Grothendieck fibration~\cite{jacobs:2001}. Then, a
-\emph{judgment} shall be interpreted as an object in the presheaf category
-$\Define{\mathbb{J}}{\Sets^{\OpCat{\mathcal{W}}}}$; in other words, a judgment
-is identified with the intensional set of its renaming-sensitive derivations.
+For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
+a generic logical framework. Recall from \cite{sterling-morrison:2015} that
+$\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
+renamings. Then, fix a small category $\mathcal{W}$ of worlds along with a map
+$\Of{\pi_\mathcal{W}}{\mathcal{W}\to\SCtx}$ (sometimes with $\mathcal{W}$
+omitted) such that $\pi_\mathcal{W}$ is a Grothendieck
+fibration~\cite{jacobs:2001}:
+\[
+  \begin{tikzcd}[cramped, ampersand replacement = \&, row sep = large]
+    {\IMode{\mathcal{W}}}
+      \arrow[dash, d]
+    \&
+    {\IMode{{w}_{i}}}
+      \arrow[drrr]
+      \arrow[dr, dotted]
+      \arrow[dd]
+    \&
+    {}\&
+    {}\&
+    {}\\
+    {\IMode{{\pi}}}
+      \arrow[d]
+    \&
+    {}\&
+    {\OMode{{w}_{j}}}
+      \arrow[rr, swap, "\OMode{{\iota}_{\varrho}}"]
+      \arrow[dd, bend right = 50]
+      \arrow[d]
+    \&
+    {}\&
+    |[alias = Wk]|{\IMode{{w}_{k}}}
+      \arrow[d]
+    \\
+    {\IMode{\SCtx}}\&
+    |[alias = Ui]|{\IMode{{\Upsilon}_{i}}}
+      \arrow[dr]
+    \&
+    {\OMode{\scriptstyle{{\pi}({w}_{j})}}}
+      \arrow[rr, "\OMode{{\pi}({\iota}_{\varrho})}", ""{name = Msrc, swap}]
+      \arrow[equals, d]
+    \&
+    {}\&
+    {\IMode{\scriptstyle{{\pi}({w}_{k})}}}
+      \arrow[equals, d]
+    \\
+    {}\&
+    {}\&
+    {\IMode{{\Upsilon}_{j}}}
+      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]
+    \&
+    {}\&
+    |[alias = Uk]|{\IMode{{\Upsilon}_{k}}}
+    \arrow[from = Wk, to = Uk, bend right = 50, crossing over]
+    \arrow[equals, from = Msrc, to = Mtgt]
+    \arrow[from = Ui, to = Uk, crossing over]
+  \end{tikzcd}
+\]
+Then, a \emph{judgment} shall be interpreted as an object in the presheaf
+category $\Define{\mathbb{J}}{\Sets^{\OpCat{\mathcal{W}}}}$; in other words, a
+judgment is identified with the intensional set of its renaming-sensitive
+derivations.
 
@@
 \ifdraft{}{
+  \nocite{maclane:1971}
   \bibliographystyle{abbrv}

### 2015-11-13 paper.tex
say "global section" following Mac Lane / Moerdijk p.47

@@
    Intuitively, the general judgment $\GenJ{\Omega}{\mathcal{J}}$ is evident
-   when the family of judgments $\mathcal{J}$ shall forevermore have a section
-   (i.e. for any substitution $\Member{\vec{E}}{\square\Omega}$,
+   when the family of judgments $\mathcal{J}$ shall forevermore have a global
+   section (i.e. for any substitution $\Member{\vec{E}}{\square\Omega}$,
    $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have
-   \emph{valences} rather than sorts; as such, the general judgment allows us to
-   quantify over terms of higher type.
+   \emph{valences} rather than sorts; as such, the general judgment allows us
+   to quantify over terms of higher type.
 

### 2015-11-13 paper.tex
Change arrow styles
Fix up another arrow
Cite Taylor; rename π_W to π

@@
 
+\tikzset{
+   commutative diagrams/.cd,
+   arrow style = tikz,
+   diagrams={>=latex}
+}
+
 \begin{document}
@@
 renamings. Then, fix a small category $\mathcal{W}$ of worlds along with a map
-$\Of{\pi_\mathcal{W}}{\mathcal{W}\to\SCtx}$ (sometimes with $\mathcal{W}$
-omitted) such that $\pi_\mathcal{W}$ is a Grothendieck
-fibration~\cite{jacobs:2001}:
+$\Of{\pi}{\mathcal{W}\to\SCtx}$ such that $\pi$ is a Grothendieck
+fibration~\cite{jacobs:2001, taylor:1999}:
 \[
@@
       \arrow[drrr]
-      \arrow[dr, dotted]
-      \arrow[dd]
+      \arrow[dotted, dr]
+      \arrow[mapsto, dd]
     \&
@@
       \arrow[rr, swap, "\OMode{{\iota}_{\varrho}}"]
-      \arrow[dd, bend right = 50]
-      \arrow[d]
+      \arrow[mapsto, dd, bend right = 50]
+      \arrow[mapsto, d]
     \&
@@
     |[alias = Wk]|{\IMode{{w}_{k}}}
-      \arrow[d]
+      \arrow[mapsto, d]
     \\
@@
     |[alias = Uk]|{\IMode{{\Upsilon}_{k}}}
-    \arrow[from = Wk, to = Uk, bend right = 50, crossing over]
+    \arrow[mapsto, from = Wk, to = Uk, bend right = 50, crossing over]
     \arrow[equals, from = Msrc, to = Mtgt]

### 2015-11-13 paper.tex
Add little annotations to make diagrams clearer
Clean up diagram code; use \null
More cleanup; make row/col seps more visible
Fix chktex warnings

@@
 For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
-a generic logical framework. Recall from \cite{sterling-morrison:2015} that
+a generic logical framework. Recall from~\cite{sterling-morrison:2015} that
 $\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
@@
 \&
-    {}\&
-    {}\&
-    {}\\
+    \null
+\&
+    \null
+\&
+    \null
+\\
     {\IMode{{\pi}}}
@@
 \&
-    {}\&
+    \null
+\&
     {\OMode{{w}_{j}}}
@@
 \&
-    {}\&
+    \null
+\&
     |[alias = Wk]|{\IMode{{w}_{k}}}
@@
 \\
-    {\IMode{\SCtx}}\&
+    {\IMode{\SCtx}}
+\&
     |[alias = Ui]|{\IMode{{\Upsilon}_{i}}}
@@
 \&
-    {}\&
+    \null
+\&
     {\IMode{\scriptstyle{{\pi}({w}_{k})}}}
@@
 \\
-    {}\&
-    {}\&
+    \null
+\&
+    \null
+\&
     {\IMode{{\Upsilon}_{j}}}
-      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]
+      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]\&
+    \null
 \&
-    {}\&
     |[alias = Uk]|{\IMode{{\Upsilon}_{k}}}
@@
    when the family of judgments $\mathcal{J}$ shall forevermore have a global
-   section (i.e. for any substitution $\Member{\vec{E}}{\square\Omega}$,
+   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Omega}$,
    $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have

### 2015-11-13 paper.tex
Formatting
Use tikz for inline arrows

@@
 \usepackage{tikz-cd}
+\usepackage{tikzpfeile}
 \usepackage{url}
@@
 \newcommand\Discrete[1]{#1_{\equiv}}
-\newcommand\Arr[1]{\mathbf{{#1}^{\rightarrow}}}
+\newcommand\Arr[1]{\mathbf{{#1}^{\ra}}}
 \newcommand\Sets{\mathbf{Set}}
@@
 renamings. Then, fix a small category $\mathcal{W}$ of worlds along with a map
-$\Of{\pi}{\mathcal{W}\to\SCtx}$ such that $\pi$ is a Grothendieck
+$\Of{\pi}{\mathcal{W}\ra\SCtx}$ such that $\pi$ is a Grothendieck
 fibration~\cite{jacobs:2001, taylor:1999}:
@@
     {\IMode{{\Upsilon}_{j}}}
-      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]\&
+      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]
+\&
     \null
@@
   the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows, where
-  $\Of{\Yoneda{-}}{\mathcal{W}\to\mathbb{J}}$ is the Yoneda Embedding:
+  $\Of{\Yoneda{-}}{\mathcal{W}\ra\mathbb{J}}$ is the Yoneda Embedding:
   \begin{align*}
@@
   $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
-  $\Of{\varrho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, and for any $M, N$,
+  $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
   what it means for $M$ and $N$ to be identified under $\phi$, supposing
@@
   This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
-  $\Upsilon\hookrightarrow\Upsilon,u:\sigma$.
+  $\Upsilon\inj\Upsilon,u:\sigma$.
 \end{proof}
@@
   To know $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
-  $\Of{\rho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, the following judgments:
+  $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
   \begin{gather*}
@@
 $M$ and $N$ are identified by $\phi$ at the identity renaming
-$\Of{1}{\IMode{\Upsilon}\hookrightarrow\OMode{\Upsilon}}$:
+$\Of{1}{\IMode{\Upsilon}\inj\OMode{\Upsilon}}$:
 \[

### 2015-11-14 paper.tex
Begin explaining lazy computation systems

@@
 \newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
+\newcommand\MapsTo[2]{\IMode{#1}\mapsto\OMode{#2}}
 
@@
 \newcommand\Operators{\mathcal{O}}
+\newcommand\CanOperators{\mathcal{K}}
 
@@
 \newcommand\IsApart[2]{\IMode{#1} \mathrel{\#} \IMode{#2}}
+\newcommand\IsSubsetEq[2]{\IMode{#1}\subseteq\IMode{#2}}
 
@@
 \newcommand\Pow[1]{\wp\left(#1\right)}
+
+\newcommand\TyBase[1]{\mathtt{Base}_{#1}}
+\newcommand\TyTop[1]{\top_{#1}}
+
+\newcommand\EvalN[5]{\IMode{#1}\parallel\IMode{#4}\Downarrow^{\OMode{#3}}_{\IMode{#2}}\OMode{#5}}
+\newcommand\Exprs{\mathcal{E}}
+\newcommand\Values{\mathcal{V}}
+\newcommand\Naturals{\mathbb{N}}
+\newcommand\NaturalsPsh{\mathcal{N}}
 
@@
 
+\section{Multi-Sorted Computational Type Theory}
+
+Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
+you specify how to form its equal canonical inhabitants, or values. Then, the
+inhabitants of the types are explained uniformly by extending this relation
+over an operational semantics such that two terms are equal members of a
+type just when they compute to equal values of that type.
+
+More generally, one may wish to consider types which include computations in
+addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
+is a type, which would include all the members of $\psi$ as well as a divergent
+computation. Another example is the type $\TyTop{\tau}$ which identifies all
+closed terms of sort $\tau$.
+
+In order to ensure that such types are sensible, we must formulate a notion of
+computational equivalence, and then say that a refinement is a \emph{type} just
+when it respects computational equivalence; then, we are able to define types
+as PERs on \emph{programs} in addition to values.
+
+\subsection{Lazy Computation Systems}
+
+In this section we generalize Howe's notion of lazy computation
+system~\cite{howe:1989} to the multi-sorted, symbol-parameterized setting. An
+\emph{lazy computation language} is an abt signature
+$\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished presheaf
+$\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
+\emph{canonical} operators such that $\IsSubsetEq{\CanOperators}{\Operators}$.
+For a lazy computation language
+$\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
+copresheaves of closed expressions and closed values over symbol contexts as
+follows:
+\begin{align*}
+  \ADefine{\Exprs_\tau(\Upsilon)}{
+    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{M}{\tau}}
+  }\\
+  \ADefine{\Values_\tau(\Upsilon)}{
+    \MkSet{
+      M\equiv\App{\vartheta}{\vec{E}}
+      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
+    }
+  }
+\end{align*}
+
+Then, a \emph{lazy computation system} (lcs) is a lazy computation language
+$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an indexed
+$\alpha$-functional relation
+$\Of{\Downarrow}{\Pow{\Exprs\times\Values\times\NaturalsPsh}}$ where
+$\NaturalsPsh$ is the constant presheaf $\MapsTo{(\Upsilon, \tau)}{\Naturals}$.
+We will write $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ when
+$\Member{\Tuple{M,N,n}}{\Downarrow_\tau(\Upsilon)}$, to mean that $M$ evaluates
+to $N$ in $n$ steps.
+
 \ifdraft{}{

### 2015-11-14 paper.tex
[paper] Use \varprojlim instead of Ran
@freebroccolo I'm just doing this for now since it's at least clear that
this works, and I don't quite know how to use Ran well enough to make it
work here (since it seems we accidentally introduced a requirement for
terminal objects in W). Please feel free to adjust this as you like
though.
lattice on refinements

@@
 
+\newcommand\SubRefines[4]{\IMode{#1}\parallel\IMode{#2}\subseteq\IMode{#3}\sqsubset\OMode{#4}}
 \newcommand\TyBase[1]{\mathtt{Base}_{#1}}
 \newcommand\TyTop[1]{\top_{#1}}
+\newcommand\TyBot[1]{\bot_{#1}}
+\newcommand\TyMeet[1]{\cap_{#1}}
+\newcommand\TyJoin[1]{\cup_{#1}}
 
@@
       }{
-        Ran_{\mathbbold{1}}{\mathcal{J}}
+        \varprojlim\mathcal{J}
       }
@@
 
-\subsubsection{Equality of parametric refinement}
+\subsubsection{Ordering and Equality of Parametric Refinement}
 
-Two refinements are equal when they denote the same PER.\@That is, will define
-the refinement equality judgment $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ as
-follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
-$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$:
+We will write $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ to mean that $\phi$ is a
+\emph{subrefinement} of $\psi$, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
+$\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$:
 \begin{definition}
-  To know $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
+  To know $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
   $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
-  \begin{gather*}
-    \GenJ{
-      \MV{m},\MV{n}:\tau
-    }{
-      \HypJ{
-        \IMode{M}\sim^\rho_\phi\IMode{N}
-      }{
-        \IMode{M}\sim^\rho_\psi\IMode{N}
-      }
-    }\\
+  \[
     \GenJ{
@@
     }
-  \end{gather*}
+  \]
+\end{definition}
 
- %% and for any $M, N$
- %% such that $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
- %% $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$,
- %% $\IMode{M}\sim^\rho_\phi\IMode{N}$ if and only if
- %% $\IMode{M}\sim^\rho_\psi\IMode{N}$.
-\end{definition}
+Two refinements are equal when they denote the same PER. That is, we have
+$\EqRefines{\Phi}{\phi}{\psi}{\tau}$ just when both
+$\SubRefines{\Phi}{\phi}{\psi}{\tau}$ and $\SubRefines{\Phi}{\psi}{\phi}{\tau}$.
 
@@
 
+\section{Constructions on Refinements}
+
+Refinements can be arranged in a partial order via the
+$\SubRefines{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
+at any sort $\tau$.
+
+We have the bottom refinement $\GRefines{\TyBot{\tau}}{\tau}$, which simply
+contains no elements. Next, the top refinement $\GRefines{\TyTop{\tau}}{\tau}$
+identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
+and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
+$\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
+$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
+\begin{gather*}
+  \infer{
+    \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
+  }{
+    \RMemEq{\Phi}{M}{N}{\phi}
+  }\qquad
+  \infer{
+    \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
+  }{
+    \RMemEq{\Phi}{M}{N}{\psi}
+  }
+\end{gather*}
+
+Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
+$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
+dually:
+\[
+  \infer{
+    \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
+  }{
+    \RMemEq{\Phi}{M}{N}{\phi} &
+    \RMemEq{\Phi}{M}{N}{\psi}
+  }
+\]
+
+It is trivial to verify that these constructions on refinements give rise to a
+lattice.
+
 \section{Multi-Sorted Computational Type Theory}

### 2015-11-14 paper.tex
notation

@@
 $\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an indexed
-$\alpha$-functional relation
+$\equiv_\alpha$-functional relation
 $\Of{\Downarrow}{\Pow{\Exprs\times\Values\times\NaturalsPsh}}$ where

### 2015-11-14 paper.tex
clarify

@@
 \emph{lazy computation language} is an abt signature
-$\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished presheaf
+$\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished copresheaf
 $\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
@@
 $\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-copresheaves of closed expressions and closed values over symbol contexts as
+copresheaves of closed expressions and closed values over symbol contexts and sorts as
 follows:

### 2015-11-14 paper.tex
[paper] special arrow for display
explain "internal judgments" in the topos logic
It's impredicative, but this is a mere formalization of what goes on in
the meaning explanations all the way back to at least Heyting. For
instance, to know "P prop" is to furnish another judgment "P verified"
and its meaning, which is in general an impredicative construction.

I think that what Mark van Atten has said about impredicativity is
relevant, which is that predicativity is not an intrinsic constraint on
constructions by the creating subject, but is rather something that may
be observed *about* construc

@@
 \newcommand\EvalN[5]{\IMode{#1}\parallel\IMode{#4}\Downarrow^{\OMode{#3}}_{\IMode{#2}}\OMode{#5}}
+\newcommand\Eval[4]{\EvalN{#1}{#2}{}{#3}{#4}}
 \newcommand\Exprs{\mathcal{E}}
@@
   arrow style = tikz,
-   diagrams={>=latex}
+  diagrams = {>=latex},
 }
+\tikzstyle{displays} = [
+  ->,
+  >=open triangle 45,
+]
 
@@
     {\IMode{{\pi}}}
-      \arrow[d]
+      \arrow[displays, d]
 \&
@@
 judgment is identified with the intensional set of its renaming-sensitive
-derivations.
+derivations. $\mathbb{J}$, like all presheaf categories, is a topos.
 
@@
   derivabilities~\cite{harper:2016}. We will not have any need for the latter.
+
+  \paragraph{Notation}
+  For a hypothetical judgment with multiple hypotheses, we will write
+  $\HypJ{\mathcal{J}_3}{\mathcal{J}_1, \mathcal{J}_2}$ rather than
+  $\HypJ{\mathcal{J}_3}{\mathcal{J}_1\times\mathcal{J}_2}$ or
+  $\HypJ{(\HypJ{\mathcal{J}_3}{\mathcal{J}_2})}{\mathcal{J}_1}$.
 \end{definition}
@@
 
+\end{definition}
+
+\begin{definition}[Internal Judgments]
+  It often happens that the evidence for a judgment is the knowledge of the
+  meaning of a new form of judgment. This can be neatly expressed in our topos
+  using the power object $\Pow{-}$: for any judgment
+  $\mathcal{J}\in\mathbb{J}$, the derivations of $\Pow{\mathcal{J}}$ shall
+  consist in subsets of the derivations of $\mathcal{J}$. The evidence of
+  $\Pow{\mathcal{J}}$ shall be called an \emph{internal judgment} $J$. In most
+  cases, we will say something like ``To know $\mathcal{J}$ is to have a
+  judgment $\mathcal{J'}$'' rather than appealing directly to the power object
+  or the subobject classifier.
 \end{definition}
@@
 $\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-copresheaves of closed expressions and closed values over symbol contexts and sorts as
+presheaves of closed expressions and closed values over symbol contexts and sorts as
 follows:
 \begin{align*}
-  \ADefine{\Exprs_\tau(\Upsilon)}{
-    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{M}{\tau}}
+  \ADefine{\Exprs_\tau(w)}{
+    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\pi(w)}{\cdot}{M}{\tau}}
   }\\
-  \ADefine{\Values_\tau(\Upsilon)}{
+  \ADefine{\Values_\tau(w)}{
     \MkSet{
       M\equiv\App{\vartheta}{\vec{E}}
-      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
+      \mid M\in\Exprs_\tau(w)\land \exists a. \CanOperators\Pair{\pi(w)}{a}\ni\vartheta
     }
@@
 
+\begin{remark}
+  If read as a judgment, $\Exprs_\tau$ (resp.\ $\Values_\tau$) means ``there is
+  a closed expression (resp.\ value) of sort $\tau$'', and has as its evidence any closed
+  expression (resp.\ value) of sort $\tau$.
+\end{remark}
+
 Then, a \emph{lazy computation system} (lcs) is a lazy computation language
-$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an indexed
-$\equiv_\alpha$-functional relation
-$\Of{\Downarrow}{\Pow{\Exprs\times\Values\times\NaturalsPsh}}$ where
-$\NaturalsPsh$ is the constant presheaf $\MapsTo{(\Upsilon, \tau)}{\Naturals}$.
-We will write $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ when
-$\Member{\Tuple{M,N,n}}{\Downarrow_\tau(\Upsilon)}$, to mean that $M$ evaluates
-to $N$ in $n$ steps.
+$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with a
+$\equiv_\alpha$-functional evaluation relation. Formally, the evaluation
+relation is given in our logical framework by a sort-indexed family of
+intensional sets of evaluation rules
+$\Of{\mathfrak{R}^\tau_\Downarrow}{\mathbb{J}}$, such that the judgment
+$\HypJ{\Pow{\mathfrak{R}^\tau_\Downarrow}}{\Exprs_\tau, \Values_\tau,
+\NaturalsPsh}$ shall be evident by means of some internal family of judgments
+$[\Downarrow_\tau]$, where $\NaturalsPsh$ is the constant presheaf
+$\MapsTo{w}{\Naturals}$. In other words, we shall forevermore know what it
+means to know that some expression evaluates to some value in $n$ steps.  By
+this token, we write $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ to mean the judgment
+whose derivations at $w$ are the set $[\Downarrow_\tau](w,M,N,n)$.
 

### 2015-11-14 paper.tex
Correct mistake in previous commit

@@
 $\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-presheaves of closed expressions and closed values over symbol contexts and sorts as
-follows:
+presheaves of closed expressions and closed values as follows:
 \begin{align*}

### 2015-11-15 paper.tex
Simplify semantic universe (close #4); show how to externalize internal judgments
- change semantic universe to Set^Sctx for simplicity (#4)

- changed \Omega to \Theta for metavariable contexts to avoid confusion
  w/ subobject classifier

- improve the definition of internal judgments, and show how to
  externalize them

- comment out some of the computation system stuff for now until I get
  things in order

--- 2015-11-14 paper.tex
@@
 
-\newcommand\Hom[3]{#1\!\left[#2,\,#3\right]\!}
+\newcommand\Hom[3]{#1\!\left[#2,\,#3\right]}
 
@@
 \newcommand\SemBrackets[1]{\left\llbracket#1\right\rrbracket}
+\newcommand\Embed[1]{\left\lfloor#1\right\rfloor}
 \newcommand\Yoneda[1]{\boldsymbol{y} (#1)}
@@
 $\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
-renamings. Then, fix a small category $\mathcal{W}$ of worlds along with a map
-$\Of{\pi}{\mathcal{W}\ra\SCtx}$ such that $\pi$ is a Grothendieck
-fibration~\cite{jacobs:2001, taylor:1999}:
-\[
-  \begin{tikzcd}[cramped, ampersand replacement = \&, row sep = large]
-    {\IMode{\mathcal{W}}}
-      \arrow[dash, d]
-\&
-    {\IMode{{w}_{i}}}
-      \arrow[drrr]
-      \arrow[dotted, dr]
-      \arrow[mapsto, dd]
-\&
-    \null
-\&
-    \null
-\&
-    \null
-\\
-    {\IMode{{\pi}}}
-      \arrow[d]
-\&
-    \null
-\&
-    {\OMode{{w}_{j}}}
-      \arrow[rr, swap, "\OMode{{\iota}_{\varrho}}"]
-      \arrow[mapsto, dd, bend right = 50]
-      \arrow[mapsto, d]
-\&
-    \null
-\&
-    |[alias = Wk]|{\IMode{{w}_{k}}}
-      \arrow[mapsto, d]
-\\
-    {\IMode{\SCtx}}
-\&
-    |[alias = Ui]|{\IMode{{\Upsilon}_{i}}}
-      \arrow[dr]
-\&
-    {\OMode{\scriptstyle{{\pi}({w}_{j})}}}
-      \arrow[rr, "\OMode{{\pi}({\iota}_{\varrho})}", ""{name = Msrc, swap}]
-      \arrow[equals, d]
-\&
-    \null
-\&
-    {\IMode{\scriptstyle{{\pi}({w}_{k})}}}
-      \arrow[equals, d]
-\\
-    \null
-\&
-    \null
-\&
-    {\IMode{{\Upsilon}_{j}}}
-      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]
-\&
-    \null
-\&
-    |[alias = Uk]|{\IMode{{\Upsilon}_{k}}}
-      \arrow[mapsto, from = Wk, to = Uk, bend right = 50, crossing over]
-      \arrow[equals, from = Msrc, to = Mtgt]
-      \arrow[from = Ui, to = Uk, crossing over]
-  \end{tikzcd}
-\]
-Then, a \emph{judgment} shall be interpreted as an object in the presheaf
-category $\Define{\mathbb{J}}{\Sets^{\OpCat{\mathcal{W}}}}$; in other words, a
+renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
+category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
 judgment is identified with the intensional set of its renaming-sensitive
-derivations.
+derivations. $\mathbb{J}$, like all presheaf categories, is a topos.
 
@@
   hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ using the
-  exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses the semantic consequence
-  of $\mathcal{J}_2$ from $\mathcal{J}_1$:
+  exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
+  the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
   \begin{align*}
     \ADefine{
-      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(w)
+      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
     }{
-      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(w)
+      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
     }\\
@@
       \Hom{\mathbb{J}}{
-        \Yoneda{w}\times\mathcal{J}_1
+        \Yoneda{\Upsilon}\times\mathcal{J}_1
       }{
@@
     \OMode{
-      \int_{w'}{{{w'}\succeq{w}}\Rightarrow{{{\mathcal{J}}_{2}(w')}^{{\mathcal{J}}_{1}(w')}}}
+      \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
+        \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
     }
@@
 \begin{definition}[General Judgment]
-  For a metavariable context $\IsMetaCtx{\Omega}$, we can define the set
+  For a metavariable context $\IsMetaCtx{\Theta}$, we can define the set
   of its substitutions as follows:
   \[
-    \Define{{\square\Omega}}{
-      \prod_{(\MV{m}:v)\in\Omega}
+    \Define{{\square\Theta}}{
+      \prod_{(\MV{m}:v)\in\Theta}
         \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
@@
 
-  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Omega}}$, we can define
-  the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows, where
-  $\Of{\Yoneda{-}}{\mathcal{W}\ra\mathbb{J}}$ is the Yoneda Embedding:
+  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
+  the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows, where
+  $\Of{\Yoneda{-}}{\OpCat{\SCtx}\ra\mathbb{J}}$ is the Yoneda Embedding:
   \begin{align*}
     \ADefine{
-      \left(\GenJ{\Omega}{\mathcal{J}}\right)(w)
+      \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
     }{
       \Hom{\mathbb{J}}{
-        \Yoneda{w}
+        \Yoneda{\Upsilon}
       }{
@@
     \OMode{
-      \int_{(w',\,\vec{E})}{{w'}\succeq{w}\Rightarrow{\mathcal{J}(\vec{E})(w')}}
+      \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
     }
@@
 
-   Intuitively, the general judgment $\GenJ{\Omega}{\mathcal{J}}$ is evident
+   Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
    when the family of judgments $\mathcal{J}$ shall forevermore have a global
-   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Omega}$,
+   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
    $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have
@@
 
-   We will write $\HPGenJ{\Upsilon}{\Gamma}{\Omega}{\mathcal{J}}$ for
-   $\GenJ{\Omega'}{\mathcal{J}}$, where
-   $\Lookup{\Omega'}{\MV{m}}{\MkValence{\Upsilon,\vec{\sigma}}{\Gamma,\vec{\tau}}{\tau}}$
-   if $\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$.
+   We will write $\HPGenJ{\Upsilon}{\Gamma}{\Theta}{\mathcal{J}}$ for
+   $\GenJ{\Theta'}{\mathcal{J}}$, where
+   $\Lookup{\Theta'}{\MV{m}}{\MkValence{\Upsilon,\vec{\sigma}}{\Gamma,\vec{\tau}}{\tau}}$
+   if $\Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$.
    Additionally, for clarity we will often write
@@
 
+\end{definition}
+
+\begin{definition}[Internal Judgments]
+
+  It often happens that the evidence for a judgment is the knowledge of the
+  meaning of a new form of judgment. This can be neatly expressed in our topos
+  using the subobject classifier $\Omega$, where the evidence of $\Omega$ at
+  $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
+  $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.  A
+  \emph{global internal judgment} is a section
+  $\Of{J}{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\Omega}}$, which can
+  be externalized as a judgment $\Of{\Embed{J}}{\mathbb{J}}$ whose derivations
+  at $\Upsilon$ are the arrows
+  $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}}}$ such that the
+  following diagram commutes:
+  \[
+    \begin{tikzcd}
+      \IMode{\Yoneda{\Upsilon}}
+        \arrow[r, dashed, "\OMode{\phi}"]
+        \arrow[rd, "\IMode{J_\Upsilon}"']
+      & \IMode{\mathbbold{1}} \arrow[d, "\IMode{\mathbf{true}}"] \\
+      & \IMode{\Omega}
+    \end{tikzcd}
+  \]
+
+  In most cases, we will say something like ``To know $\mathcal{J}$ is to have
+  a judgment $\mathcal{J'}$'' rather than appealing directly to the subobject
+  classifier.
 \end{definition}
@@
 $\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-copresheaves of closed expressions and closed values over symbol contexts and sorts as
-follows:
+presheaves of closed expressions and closed values as follows:
 \begin{align*}
@@
 
+\begin{remark}
+  If read as a judgment, $\Exprs_\tau$ (resp.\ $\Values_\tau$) means ``there is
+  a closed expression (resp.\ value) of sort $\tau$'', and has as its evidence any closed
+  expression (resp.\ value) of sort $\tau$.
+\end{remark}
+
 Then, a \emph{lazy computation system} (lcs) is a lazy computation language
-$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an indexed
-$\equiv_\alpha$-functional relation
-$\Of{\Downarrow}{\Pow{\Exprs\times\Values\times\NaturalsPsh}}$ where
-$\NaturalsPsh$ is the constant presheaf $\MapsTo{(\Upsilon, \tau)}{\Naturals}$.
-We will write $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ when
-$\Member{\Tuple{M,N,n}}{\Downarrow_\tau(\Upsilon)}$, to mean that $M$ evaluates
-to $N$ in $n$ steps.
+$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an $\SCtx$-indexed
+$\equiv_\alpha$-functional evaluation relation
+$\EvalN{\Upsilon}{\tau}{n}{M}{N}$ presupposing $\Member{n}{\Naturals}$,
+$\Member{M}{\Exprs_\tau(\Upsilon)}$ and $\Member{N}{\Values_\tau(\Upsilon)}$,
+expressing that $M$ evaluates to $N$ in $n$ steps.
+
+%% \begin{remark}
+%%
+%%   Formally, the evaluation relation is given in our logical framework by a
+%%   presheaf of evaluation rules $\Of{\mathfrak{R}^\tau_\Downarrow}{\mathbb{J}}$,
+%%   such that the judgment
+%%   $\HypJ{\Pow{\mathfrak{R}^\tau_\Downarrow}}{\Exprs_\tau, \Values_\tau,
+%%   \NaturalsPsh}$ is globally evident where $\NaturalsPsh$ is the constant
+%%   presheaf $\MapsTo{\Upsilon}{\Naturals}$. In other words, we have a family of
+%%   internal judgments $\Of{[\Downarrow_\tau](\Upsilon)}{\int_{\Upsilon'}
+%%   \Upsilon\inj\Upsilon'\Rightarrow\Pow{\mathfrak{R}^\tau_\Downarrow}}$
+%%
+%%
+%% shall be evident by means of some internal
+%% family of judgments $[\Downarrow_\tau]$, where $\NaturalsPsh$ is the constant
+%% presheaf $\MapsTo{\Upsilon}{\Naturals}$. In other words, we shall forevermore know
+%% what it means to know that some expression evaluates to some value in $n$
+%% steps.
+%% \end{remark}
+%%
+%% We will write $\Eval{\Upsilon}{\tau}{M}{N}$ to mean that there exists an $n$
+%% such that $\EvalN{\Upsilon}{\tau}{n}{M}{N}$; in other words, the judgment is
+%% interpreted as the following presheaf:
+%% \[
+%%   \Define{
+%%     \left(\Upsilon\parallel M\Downarrow_\tau N\right)(w)
+%%   }{
+%%     \int^n[\Downarrow_\tau](w,M,N,n)
+%%   }
+%% \]
+%%
+%%
+%% By this token, the derivations of $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ at
+%% $w$ are the set $[\Downarrow_\tau](w,M,N,n)$; we write
+%% $\Eval{\Upsilon}{\tau}{M}{N}$ for the judgment
+%% $\int^n[\Downarrow_\tau](-,M,N,n)$
+%
 

### 2015-11-15 paper.tex
slight clarification of internal judgment

@@
   $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
-  $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.  A
-  \emph{global internal judgment} is a section
-  $\Of{J}{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\Omega}}$, which can
+  $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.
+
+  A \emph{global internal judgment} is a section
+  $\Of{J}{\int_\Upsilon{\Omega^{\mathbbold{1}}}(\Upsilon)}\cong
+  \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\Omega}}$, which can
   be externalized as a judgment $\Of{\Embed{J}}{\mathbb{J}}$ whose derivations
@@
   following diagram commutes:
+
   \[

### 2015-11-15 paper.tex
Restore (changed version) of lcs definition

@@
 $\Member{M}{\Exprs_\tau(\Upsilon)}$ and $\Member{N}{\Values_\tau(\Upsilon)}$,
-expressing that $M$ evaluates to $N$ in $n$ steps.
+expressing that $M$ evaluates to $N$ in $n$ steps.  We will write
+$\Eval{\Upsilon}{\tau}{M}{N}$ to mean that there exists an $n$ such that
+$\EvalN{\Upsilon}{\tau}{n}{M}{N}$.
 
-%% \begin{remark}
-%%
-%%   Formally, the evaluation relation is given in our logical framework by a
-%%   presheaf of evaluation rules $\Of{\mathfrak{R}^\tau_\Downarrow}{\mathbb{J}}$,
-%%   such that the judgment
-%%   $\HypJ{\Pow{\mathfrak{R}^\tau_\Downarrow}}{\Exprs_\tau, \Values_\tau,
-%%   \NaturalsPsh}$ is globally evident where $\NaturalsPsh$ is the constant
-%%   presheaf $\MapsTo{\Upsilon}{\Naturals}$. In other words, we have a family of
-%%   internal judgments $\Of{[\Downarrow_\tau](\Upsilon)}{\int_{\Upsilon'}
-%%   \Upsilon\inj\Upsilon'\Rightarrow\Pow{\mathfrak{R}^\tau_\Downarrow}}$
-%%
-%%
-%% shall be evident by means of some internal
-%% family of judgments $[\Downarrow_\tau]$, where $\NaturalsPsh$ is the constant
-%% presheaf $\MapsTo{\Upsilon}{\Naturals}$. In other words, we shall forevermore know
-%% what it means to know that some expression evaluates to some value in $n$
-%% steps.
-%% \end{remark}
-%%
-%% We will write $\Eval{\Upsilon}{\tau}{M}{N}$ to mean that there exists an $n$
-%% such that $\EvalN{\Upsilon}{\tau}{n}{M}{N}$; in other words, the judgment is
-%% interpreted as the following presheaf:
-%% \[
-%%   \Define{
-%%     \left(\Upsilon\parallel M\Downarrow_\tau N\right)(w)
-%%   }{
-%%     \int^n[\Downarrow_\tau](w,M,N,n)
-%%   }
-%% \]
-%%
-%%
-%% By this token, the derivations of $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ at
-%% $w$ are the set $[\Downarrow_\tau](w,M,N,n)$; we write
-%% $\Eval{\Upsilon}{\tau}{M}{N}$ for the judgment
-%% $\int^n[\Downarrow_\tau](-,M,N,n)$
-%
+ \begin{remark}
+   Formally, the evaluation relation is given in our logical framework by a
+   global internal judgment $\Of{\Downarrow_\tau}{\int_\SCtx
+   \Omega^{\Exprs_\tau\times\Values_\tau\times\NaturalsPsh}}$, where
+   $\NaturalsPsh$ is the constant presheaf $\MapsTo{\Upsilon}{\Naturals}$.
+   Then, we know $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ just when the
+   externalization $\Embed{\Downarrow_\tau}(M,N,n)$ is evident at $\Upsilon$.
+ \end{remark}
 

### 2015-11-15 paper.tex
More fixing up of the internal judgments to support families of judgments

@@
 
-  A \emph{global internal judgment} is a section
-  $\Of{J}{\int_\Upsilon{\Omega^{\mathbbold{1}}}(\Upsilon)}\cong
-  \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\Omega}}$, which can
-  be externalized as a judgment $\Of{\Embed{J}}{\mathbb{J}}$ whose derivations
-  at $\Upsilon$ are the arrows
-  $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}}}$ such that the
-  following diagram commutes:
-
+  A \emph{global internal (family of) judgments} is a section
+  $\Of{J}{\int_\SCtx {\Omega^{\mathcal{J}}}}\cong
+  \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}\times\mathcal{J}}{\Omega}}$,
+  which can be externalized as a ``total'' judgment $\Of{\Embed{J}}{\mathbb{J}}$
+  for whose derivations at $\Upsilon$ are the arrows
+  $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}\times\mathcal{J}}}$
+  such that the following diagram commutes:
   \[
-    \begin{tikzcd}
+    \begin{tikzcd}[cramped,sep=huge]
       \IMode{\Yoneda{\Upsilon}}
         \arrow[r, dashed, "\OMode{\phi}"]
-        \arrow[rd, "\IMode{J_\Upsilon}"']
-      & \IMode{\mathbbold{1}} \arrow[d, "\IMode{\mathbf{true}}"] \\
+        \arrow[rd, "\IMode{J_\Upsilon\Pair{1}{\pi_2\circ\phi}}"']
+      & \IMode{\mathbbold{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
       & \IMode{\Omega}
@@
 
-  In most cases, we will say something like ``To know $\mathcal{J}$ is to have
-  a judgment $\mathcal{J'}$'' rather than appealing directly to the subobject
-  classifier.
+  Then, to say $\IMode{\Upsilon}\parallel\Embed{J}(E)$ where
+  $\Member{E}{\mathcal{J}(\Upsilon)}$ is to assert that we have evidence
+  $\Member{F}{\Embed{J}(\Upsilon)}$ such that $\IsEq{\pi_2(F)}{E}$.
+
+  When asserting the existence of an internal judgment, in most cases we will
+  say something like ``To know $\mathcal{J}$ is to have a judgment
+  $\mathcal{J'}$'' rather than appealing directly to the subobject classifier.
 \end{definition}
@@
    $\NaturalsPsh$ is the constant presheaf $\MapsTo{\Upsilon}{\Naturals}$.
-   Then, we know $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ just when the
-   externalization $\Embed{\Downarrow_\tau}(M,N,n)$ is evident at $\Upsilon$.
  \end{remark}

### 2015-11-20 paper.tex
Remove formalization of evaluation relation (#5)
Switch to informal meaning explanations for the LF (#5)

--- 2015-11-15 paper.tex
@@
 
-For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
-a generic logical framework. Recall from~\cite{sterling-morrison:2015} that
-$\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
-renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
-category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
-judgment is identified with the intensional set of its renaming-sensitive
-derivations. $\mathbb{J}$, like all presheaf categories, is a topos.
+For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
+deploy a generic logical framework equipped with higher-order judgments. A
+judgment is defined by giving it a \emph{meaning explanation}, which is a
+specification of the conditions under which it may be asserted (sc.\ its
+evidence).
+
+%% Recall from~\cite{sterling-morrison:2015} that
+%% $\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
+%% renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
+%% category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
+%% judgment is identified with the intensional set of its renaming-sensitive
+%% derivations. $\mathbb{J}$, like all presheaf categories, is a topos.
 
@@
   For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
-  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ using the
-  exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
-  the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
-  \begin{align*}
-    \ADefine{
-      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
-    }{
-      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
-    }\\
-    &\cong
-    \OMode{
-      \Hom{\mathbb{J}}{
-        \Yoneda{\Upsilon}\times\mathcal{J}_1
-      }{
-        \mathcal{J}_2
-      }
-    }\\
-    &\cong
-    \OMode{
-      \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
-        \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
-    }
-  \end{align*}
-
-  In other words, the hypothetical judgment
-  $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
-  be an effective transformation of evidence for the antecedent into evidence
-  for the consequent. It is important to keep in mind that this is a
+  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$, whose evidence
+  shall be an effective transformation of any evidence for $\mathcal{J}_1$ into
+  evidence for $\mathcal{J}_2$. It is important to keep in mind that this is a
   \emph{semantic} consequence and therefore expresses admissibilities, in
@@
   derivabilities~\cite{harper:2016}. We will not have any need for the latter.
+
+  %% using the
+  %% exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
+  %% the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
+  %% \begin{align*}
+  %%   \ADefine{
+  %%     \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
+  %%   }{
+  %%     \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
+  %%   }\\
+  %%   &\cong
+  %%   \OMode{
+  %%     \Hom{\mathbb{J}}{
+  %%       \Yoneda{\Upsilon}\times\mathcal{J}_1
+  %%     }{
+  %%       \mathcal{J}_2
+  %%     }
+  %%   }\\
+  %%   &\cong
+  %%   \OMode{
+  %%     \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
+  %%       \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
+  %%   }
+  %% \end{align*}
+
+  %% In other words, the hypothetical judgment
+  %% $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
+  %% be an effective transformation of evidence for the antecedent into evidence
+  %% for the consequent.
 
@@
 
-  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
-  the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows, where
-  $\Of{\Yoneda{-}}{\OpCat{\SCtx}\ra\mathbb{J}}$ is the Yoneda Embedding:
-  \begin{align*}
-    \ADefine{
-      \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
-    }{
-      \Hom{\mathbb{J}}{
-        \Yoneda{\Upsilon}
-      }{
-        \varprojlim\mathcal{J}
-      }
-    }\\
-    &\cong
-    \OMode{
-      \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
-    }
-  \end{align*}
+  For any $\square{\Theta}$-indexed family of judgments $\mathcal{J}$, we can
+  express the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as being evident
+  when $\mathcal{J}(\vec{E})$ is evident for each
+  $\Of{\vec{E}}{\square\Theta}$; in other words, the evidence of the general
+  judgment is an effective transformation of $\Theta$-substitutions to their
+  fibres in $\mathcal{J}$.  Recall that metavariables have \emph{valences}
+  rather than sorts; as such, the general judgment allows us to quantify over
+  terms of higher type.
 
-   Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
-   when the family of judgments $\mathcal{J}$ shall forevermore have a global
-   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
-   $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have
-   \emph{valences} rather than sorts; as such, the general judgment allows us
-   to quantify over terms of higher type.
+%%
+%%  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
+%%  the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows, where
+%%  $\Of{\Yoneda{-}}{\OpCat{\SCtx}\ra\mathbb{J}}$ is the Yoneda Embedding:
+%%  \begin{align*}
+%%    \ADefine{
+%%      \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
+%%    }{
+%%      \Hom{\mathbb{J}}{
+%%        \Yoneda{\Upsilon}
+%%      }{
+%%        \varprojlim\mathcal{J}
+%%      }
+%%    }\\
+%%    &\cong
+%%    \OMode{
+%%      \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
+%%    }
+%%  \end{align*}
+%%
+%%   Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
+%%   when the family of judgments $\mathcal{J}$ shall forevermore have a global
+%%   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
+%%   $\mathcal{J}(\vec{E})$ shall be evident).
 
@@
 
-\begin{definition}[Internal Judgments]
-
-  It often happens that the evidence for a judgment is the knowledge of the
-  meaning of a new form of judgment. This can be neatly expressed in our topos
-  using the subobject classifier $\Omega$, where the evidence of $\Omega$ at
-  $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
-  $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.
-
-  A \emph{global internal judgment} is a section
-  $\Of{J}{\int_\Upsilon{\Omega^{\mathbbold{1}}}(\Upsilon)}\cong
-  \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\Omega}}$, which can
-  be externalized as a judgment $\Of{\Embed{J}}{\mathbb{J}}$ whose derivations
-  at $\Upsilon$ are the arrows
-  $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}}}$ such that the
-  following diagram commutes:
-
-  \[
-    \begin{tikzcd}
-      \IMode{\Yoneda{\Upsilon}}
-        \arrow[r, dashed, "\OMode{\phi}"]
-        \arrow[rd, "\IMode{J_\Upsilon}"']
-      & \IMode{\mathbbold{1}} \arrow[d, "\IMode{\mathbf{true}}"] \\
-      & \IMode{\Omega}
-    \end{tikzcd}
-  \]
-
-  In most cases, we will say something like ``To know $\mathcal{J}$ is to have
-  a judgment $\mathcal{J'}$'' rather than appealing directly to the subobject
-  classifier.
-\end{definition}
+%% \begin{definition}[Internal Judgments]
+%%
+%%   It often happens that the evidence for a judgment is the knowledge of the
+%%   meaning of a new form of judgment. This can be neatly expressed in our topos
+%%   using the subobject classifier $\Omega$, where the evidence of $\Omega$ at
+%%   $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
+%%   $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.
+%%
+%%   A \emph{global internal (family of) judgments} is a section
+%%   $\Of{J}{\int_\SCtx {\Omega^{\mathcal{J}}}}\cong
+%%   \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}\times\mathcal{J}}{\Omega}}$,
+%%   which can be externalized as a ``total'' judgment $\Of{\Embed{J}}{\mathbb{J}}$
+%%   for whose derivations at $\Upsilon$ are the arrows
+%%   $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}\times\mathcal{J}}}$
+%%   such that the following diagram commutes:
+%%   \[
+%%     \begin{tikzcd}[cramped,sep=huge]
+%%       \IMode{\Yoneda{\Upsilon}}
+%%         \arrow[r, dashed, "\OMode{\phi}"]
+%%         \arrow[rd, "\IMode{J_\Upsilon\Pair{1}{\pi_2\circ\phi}}"']
+%%       & \IMode{\mathbbold{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
+%%       & \IMode{\Omega}
+%%     \end{tikzcd}
+%%   \]
+%%
+%%   Then, to say $\IMode{\Upsilon}\parallel\Embed{J}(E)$ where
+%%   $\Member{E}{\mathcal{J}(\Upsilon)}$ is to assert that we have evidence
+%%   $\Member{F}{\Embed{J}(\Upsilon)}$ such that $\IsEq{\pi_2(F)}{E}$.
+%%
+%%   When asserting the existence of an internal judgment, in most cases we will
+%%   say something like ``To know $\mathcal{J}$ is to have a judgment
+%%   $\mathcal{J'}$'' rather than appealing directly to the subobject classifier.
+%% \end{definition}
 
@@
 
- \begin{remark}
-   Formally, the evaluation relation is given in our logical framework by a
-   global internal judgment $\Of{\Downarrow_\tau}{\int_\SCtx
-   \Omega^{\Exprs_\tau\times\Values_\tau\times\NaturalsPsh}}$, where
-   $\NaturalsPsh$ is the constant presheaf $\MapsTo{\Upsilon}{\Naturals}$.
-   Then, we know $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ just when the
-   externalization $\Embed{\Downarrow_\tau}(M,N,n)$ is evident at $\Upsilon$.
- \end{remark}
-
 \ifdraft{}{

### 2015-11-20 paper.tex
Document renaming convention, stub appendix (#5)

@@
 \usepackage{url}
+\usepackage[titletoc,title]{appendix}
 \usepackage{todonotes}
@@
 
-%% Recall from~\cite{sterling-morrison:2015} that
-%% $\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
-%% renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
+\paragraph{Renaming convention} Because the evidence of a judgment will often
+contain symbols $u,v,\dots$ we informally adopt the convention that the
+evidence of a judgment shall always be subject to fresh renamings of its free
+symbols (i.e.\ the validity of a judgment shall not depend on choice of names).
+This apparatus is developed formally in Appendix~\ref{appendix:lf}, where the
+logical framework is cast as a topos of covariant presheaves on symbol contexts
+$\SCtx$.
+
+%% Then, a \emph{judgment} shall be interpreted as an object in the presheaf
 %% category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
@@
 
+\newpage
+\begin{appendices}
+\section{Formal Definition of the Logical Framework}
+\label{appendix:lf}
+\end{appendices}
+
 \ifdraft{}{
+  \newpage
   \nocite{maclane:1971}

### 2015-11-20 paper.tex
Move the formal definitions of higher order judgments into appendix (#5)

--- 2015-11-15 paper.tex
@@
 
-For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
-a generic logical framework. Recall from~\cite{sterling-morrison:2015} that
-$\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
-renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
-category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
-judgment is identified with the intensional set of its renaming-sensitive
-derivations. $\mathbb{J}$, like all presheaf categories, is a topos.
+For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
+deploy a generic logical framework equipped with higher-order judgments. A
+judgment is always defined using a \emph{meaning explanation}, which is a
+specification of the conditions under which it may be asserted (sc.\ its
+evidence) \cite{martin-lof:1996}.
+
+\paragraph{Renaming convention} Because the evidence of a judgment will often
+contain symbols $u,v,\dots$ we informally adopt the convention that the
+evidence of a judgment shall always be subject to fresh renamings of its free
+symbols (i.e.\ the validity of a judgment shall not depend on choice of names).
+This apparatus is developed formally in Appendix~\ref{appendix:lf}, where the
+logical framework is cast as a topos of covariant presheaves on symbol contexts
+$\SCtx$.
 
@@
 
-  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
-  the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows, where
-  $\Of{\Yoneda{-}}{\OpCat{\SCtx}\ra\mathbb{J}}$ is the Yoneda Embedding:
-  \begin{align*}
-    \ADefine{
-      \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
-    }{
-      \Hom{\mathbb{J}}{
-        \Yoneda{\Upsilon}
-      }{
-        \varprojlim\mathcal{J}
-      }
-    }\\
-    &\cong
-    \OMode{
-      \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
-    }
-  \end{align*}
-
-   Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
-   when the family of judgments $\mathcal{J}$ shall forevermore have a global
-   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
-   $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have
-   \emph{valences} rather than sorts; as such, the general judgment allows us
-   to quantify over terms of higher type.
+  For any $\square{\Theta}$-indexed family of judgments $\mathcal{J}$, we can
+  express the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as being evident
+  when $\mathcal{J}(\vec{E})$ is evident for each
+  $\Of{\vec{E}}{\square\Theta}$; in other words, the evidence of the general
+  judgment is an effective transformation of $\Theta$-substitutions to their
+  fibres in $\mathcal{J}$.  Recall that metavariables have \emph{valences}
+  rather than sorts; as such, the general judgment allows us to quantify over
+  terms of higher type.
 
@@
 
-\end{definition}
-
-\begin{definition}[Internal Judgments]
-
-  It often happens that the evidence for a judgment is the knowledge of the
-  meaning of a new form of judgment. This can be neatly expressed in our topos
-  using the subobject classifier $\Omega$, where the evidence of $\Omega$ at
-  $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
-  $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.
-
-  A \emph{global internal (family of) judgments} is a section
-  $\Of{J}{\int_\SCtx {\Omega^{\mathcal{J}}}}\cong
-  \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}\times\mathcal{J}}{\Omega}}$,
-  which can be externalized as a ``total'' judgment $\Of{\Embed{J}}{\mathbb{J}}$
-  for whose derivations at $\Upsilon$ are the arrows
-  $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}\times\mathcal{J}}}$
-  such that the following diagram commutes:
-  \[
-    \begin{tikzcd}[cramped,sep=huge]
-      \IMode{\Yoneda{\Upsilon}}
-        \arrow[r, dashed, "\OMode{\phi}"]
-        \arrow[rd, "\IMode{J_\Upsilon\Pair{1}{\pi_2\circ\phi}}"']
-      & \IMode{\mathbbold{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
-      & \IMode{\Omega}
-    \end{tikzcd}
-  \]
-
-  Then, to say $\IMode{\Upsilon}\parallel\Embed{J}(E)$ where
-  $\Member{E}{\mathcal{J}(\Upsilon)}$ is to assert that we have evidence
-  $\Member{F}{\Embed{J}(\Upsilon)}$ such that $\IsEq{\pi_2(F)}{E}$.
-
-  When asserting the existence of an internal judgment, in most cases we will
-  say something like ``To know $\mathcal{J}$ is to have a judgment
-  $\mathcal{J'}$'' rather than appealing directly to the subobject classifier.
 \end{definition}
@@
 
- \begin{remark}
-   Formally, the evaluation relation is given in our logical framework by a
-   global internal judgment $\Of{\Downarrow_\tau}{\int_\SCtx
-   \Omega^{\Exprs_\tau\times\Values_\tau\times\NaturalsPsh}}$, where
-   $\NaturalsPsh$ is the constant presheaf $\MapsTo{\Upsilon}{\Naturals}$.
- \end{remark}
+\newpage
+\begin{appendices}
+\section{Formal Definition of the Logical Framework}
+\label{appendix:lf}
+
+As noted earlier, we must be cautious in the definitions of both atomic and
+higher-order judgments that their evidence shall always respect the
+\emph{renaming convention}. We do so explicitly in this appendix by developing
+our logical framework as a presheaf topos over symbol contexts.
+
+A judgment is, then, an object in the presheaf category
+$\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, judgments are identified
+with the intensional set of its renameable evidence. $\mathbb{J}$, like all
+presheaf categories, is a topos. The task at hand, then, is to show how to
+replace the informal meaning explanations that we gave in
+section~\ref{section:ambient-lf} with precise definitions in the topos
+$\mathbb{J}$.
+
+\subsection{Hypothetical Judgment}
+
+Hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is defined using the
+exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
+the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
+\begin{align*}
+  \ADefine{
+    \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
+  }{
+    \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
+  }\\
+  &\cong
+  \OMode{
+    \Hom{\mathbb{J}}{
+      \Yoneda{\Upsilon}\times\mathcal{J}_1
+    }{
+      \mathcal{J}_2
+    }
+  }\\
+  &\cong
+  \OMode{
+    \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
+      \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
+  }
+\end{align*}
+
+In other words, the hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$
+is evident when there shall ``forevermore'' (i.e. for any applicable renaming
+of symbols) be an effective transformation of evidence for the antecedent into
+evidence for the consequent.
+
+\subsection{General Judgment}
+
+For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
+the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows:
+\begin{align*}
+  \ADefine{
+    \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
+  }{
+    \Hom{\mathbb{J}}{
+      \Yoneda{\Upsilon}
+    }{
+      \varprojlim\mathcal{J}
+    }
+  }\\
+  &\cong
+  \OMode{
+    \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
+  }
+\end{align*}
+
+ Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
+ when the family of judgments $\mathcal{J}$ shall forevermore have a global
+ section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
+ $\mathcal{J}(\vec{E})$ shall be evident).
+
+\subsection{Internal Judgments}
+It often happens that the evidence for a judgment is the knowledge of the
+meaning of a new form of judgment; this is the case when we say ``To know
+$\mathcal{J}$ is to know \emph{the meaning of} another judgment
+$\mathcal{J}'$.'' This can be neatly expressed in our topos using the subobject
+classifier $\Omega$, where the evidence of $\Omega$ at $\Upsilon$ gives rise to
+an \emph{internal judgment} which is defined at all $\Upsilon'$ for which we
+have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.
+
+A \emph{global internal (family of) judgments} is a section
+$\Of{J}{\int_\SCtx {\Omega^{\mathcal{J}}}}\cong
+\IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}\times\mathcal{J}}{\Omega}}$,
+which can be externalized as a ``total'' judgment $\Of{\Embed{J}}{\mathbb{J}}$
+for whose derivations at $\Upsilon$ are the arrows
+$\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}\times\mathcal{J}}}$
+such that the following diagram commutes:
+\[
+  \begin{tikzcd}[cramped,sep=huge]
+    \IMode{\Yoneda{\Upsilon}}
+      \arrow[r, dashed, "\OMode{\phi}"]
+      \arrow[rd, "\IMode{J_\Upsilon\Pair{1}{\pi_2\circ\phi}}"']
+    & \IMode{\mathbbold{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
+    & \IMode{\Omega}
+  \end{tikzcd}
+\]
+
+Then, to say $\IMode{\Upsilon}\parallel\Embed{J}(E)$ where
+$\Member{E}{\mathcal{J}(\Upsilon)}$ is to assert that we have evidence
+$\Member{F}{\Embed{J}(\Upsilon)}$ such that $\IsEq{\pi_2(F)}{E}$.
+
+\end{appendices}
 
 \ifdraft{}{
+  \newpage
   \nocite{maclane:1971}

### 2015-11-20 paper.tex
fix up a missing label
Define applicative approximation (#3)
@jozefg

@@
 ]{microtype}
-\DeclareSymbolFont{bbold}{U}{bbold}{m}{n}
-\DeclareSymbolFontAlphabet{\mathbbold}{bbold}
+\usepackage{bbold}
 
@@
 }
-\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
+\newcommand\MkBTm[3]{\mathbb{\lambda}\{#1\}[#2].\, #3}
 \newcommand\MkArity[2]{(#1)\, #2}
@@
 
-\newcommand\Rename[2]{%
-  #2\upharpoonright#1%
-}
-
-\newcommand\IsRenaming[3]{%
-  \Rename{\IMode{#1}}{\IMode{#2}}%
-  \leadsto\OMode{#3}%
+\newcommand\Rename[3]{%
+  \left\{#1\, /\, #2\right\}#3%
 }
@@
 \newcommand\Eval[4]{\EvalN{#1}{#2}{}{#3}{#4}}
-\newcommand\Exprs{\mathcal{E}}
-\newcommand\Values{\mathcal{V}}
+\newcommand\Exprs{\mathbf{E}}
+\newcommand\Values{\mathbf{V}}
+\newcommand\BTms{\mathbf{B}}
 \newcommand\Naturals{\mathbb{N}}
@@
 \section{The Ambient Logical Framework}
+\label{section:ambient-lf}
 
@@
 \emph{lazy computation language} is an abt signature
-$\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished copresheaf
-$\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
+$\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished
+copresheaf $\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
 \emph{canonical} operators such that $\IsSubsetEq{\CanOperators}{\Operators}$.
@@
 $\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-presheaves of closed expressions and closed values as follows:
+covariant presheaves on $\SCtx\times\Ctx$ of open expressions, open values, and open bound terms
+as follows:
 \begin{align*}
-  \ADefine{\Exprs_\tau(\Upsilon)}{
-    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{M}{\tau}}
+  \ADefine{\Exprs_\tau(\Upsilon\parallel\Gamma)}{
+    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\Gamma}{M}{\tau}}
   }\\
-  \ADefine{\Values_\tau(\Upsilon)}{
+  \ADefine{\Values_\tau(\Upsilon\parallel\Gamma)}{
     \MkSet{
       M\equiv\App{\vartheta}{\vec{E}}
-      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
+      \mid M\in\Exprs_\tau(\Upsilon\parallel\Gamma)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
     }
+  }\\
+  \ADefine{
+    \BTms_{v}(\Upsilon\parallel\Gamma)
+  }{
+    \MkSet{E \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\Gamma}{E}{v}}
   }
@@
 
-\begin{remark}
-  If read as a judgment, $\Exprs_\tau$ (resp.\ $\Values_\tau$) means ``there is
-  a closed expression (resp.\ value) of sort $\tau$'', and has as its evidence any closed
-  expression (resp.\ value) of sort $\tau$.
-\end{remark}
-
-Then, a \emph{lazy computation system} (lcs) is a lazy computation language
+We'll write $\Exprs_\tau(\Upsilon)$, $\Values_\tau(\Upsilon)$ and
+$\BTms_v(\Upsilon)$ for $\Exprs_\tau(\Upsilon\parallel\cdot)$,
+$\Values_\tau(\Upsilon\parallel\cdot)$ and $\BTms_v(\Upsilon\parallel\cdot)$
+respectively, viewed as covariant presheaves on $\SCtx$. Then, a \emph{lazy
+computation system} (lcs) is a lazy computation language
 $\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an $\SCtx$-indexed
@@
 $\EvalN{\Upsilon}{\tau}{n}{M}{N}$.
+
+In any topos $\mathcal{E}$, for an object $X$ we can define the object of
+relations on $X$ as the exponential $\Define{\Pow{X}}{\Omega^X}$. Then, the
+data of such a relation is contained in a natural transformation
+$\Of{R}{\Hom{\mathcal{E}}{\mathbb{1}}{\Pow{X}}}$.
+
+\newcommand\SBinRel[4]{\IMode{#2}\parallel\IMode{#3}\mathrel{#1}\IMode{#4}}
+
+Fix a sort-indexed family of binary relations on closed expressions
+$\Of{R_\tau}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$; we will
+also write the relation as a judgment scheme
+$\SBinRel{R_\tau}{\Upsilon}{M}{N}$. Now, we can always extend $R_\tau$ to a new
+family of relations
+$\Of{R_v}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for any valence
+$\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
+follows:
+\[
+  \infer{
+    \SBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
+      \MkBTm{\vec{u}}{\vec{x}}{M}
+    }{
+      \MkBTm{\vec{v}}{\vec{y}}{N}
+    }
+  }{
+    \begin{array}{l}
+      \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon}\cup\vec{u}\cup\vec{v}}.\quad
+      \forall\Of{\vec{X}}{\Exprs_\tau^{[\vec{\tau}]}(\Upsilon,\vec{w}:\vec{\sigma})}.\\
+      \SBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{
+        \Subst{\vec{X}}{\vec{x}}{\Rename{\vec{w}}{\vec{u}}{M}}
+      }{
+        \Subst{\vec{X}}{\vec{y}}{\Rename{\vec{w}}{\vec{v}}{N}}
+      }
+    \end{array}
+  }
+\]
+
+As a convenience, we'll also define this judgment over vectors of bound terms and valences:
+\[
+  \infer{
+    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
+  }{
+    \forall\Member{(E,F,v)}{(\vec{E},\vec{F},\vec{v})}.\
+    \SBinRel{R_v}{\Upsilon}{E}{F}
+  }
+\]
+
+Now, we can extend $R_\tau$ to a new relation $[R_\tau]$ on closed expressions
+which respects a single ``layer'' of computation, defined pointwise in the following way:
+\begin{gather*}
+  \infer{
+    \SBinRel{[R_\tau]}{\Upsilon}{M}{N}
+  }{
+    \Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}} &
+    \Eval{\Upsilon}{\tau}{N}{\App{\vartheta}{\Vec{F}}} &
+    \Lookup{\CanOperators}{\vartheta}{\MkArity{\vec{v}}{\tau}} &
+    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
+  }
+\end{gather*}
+
+Because $[-_\tau]$ is monotonic, it has a greatest fixed point, which we will
+call \emph{applicative approximation},
+$\Of{\preceq_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
+written pointwise as $\SBinRel{\preceq_\tau}{\Upsilon}{M}{N}$.
+
+%%% Now, let us consider relations on open terms; in general, it will be easiest
+%%% for us to situate our constructions on relations in the presheaf topos
+%%% $\Define{\widehat{\HypCat}}{\Sets^{\SCtx\times\Ctx}}$. Then, for any two
+%%% objects objects $\Of{X,Y}{\widehat{\HypCat}}$, we can form the object of
+%%% relations between them as $\Define{\Pow{X\times Y}}{\Omega^{X\times Y}}$.
+%%%
+%% . A binary relation $R_\tau$
+%% between open expressions of sort $\tau$ can be thought of as a global section
+%% of the exponential object
+%% $\Match{\Pow{\Exprs_\tau\times\Exprs_\tau}}{\Omega^{\Exprs_\tau\times\Exprs_\tau}}$
+%% of the presheaf topos $\Define{\widehat{\HypCat}}{\Sets^{\SCtx\times\Ctx}}$
 
@@
 for whose derivations at $\Upsilon$ are the arrows
-$\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}\times\mathcal{J}}}$
+$\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbb{1}\times\mathcal{J}}}$
 such that the following diagram commutes:
@@
       \arrow[rd, "\IMode{J_\Upsilon\Pair{1}{\pi_2\circ\phi}}"']
-    & \IMode{\mathbbold{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
+    & \IMode{\mathbb{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
     & \IMode{\Omega}

### 2015-11-20 paper.tex
Correct computational approximation definition (bad notation still)

@@
 
-\newcommand\HypJ[2]{#1\ \ (#2)}
+\newcommand\HypJ[2]{#1\ \ \left(#2\right)}
 \newcommand\GenJ[2]{\vert_{\IMode{#1}}\; #2}
@@
 
+\begin{remark}
   In any topos $\mathcal{E}$, for an object $X$ we can define the object of
@@
   $\Of{R}{\Hom{\mathcal{E}}{\mathbb{1}}{\Pow{X}}}$.
+\end{remark}
 
@@
 
-As a convenience, we'll also define this judgment over vectors of bound terms and valences:
+As a matter of convenience, we'll also define this judgment over vectors of bound terms and valences:
 \[
@@
   }{
-    \Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}} &
-    \Eval{\Upsilon}{\tau}{N}{\App{\vartheta}{\Vec{F}}} &
-    \Lookup{\CanOperators}{\vartheta}{\MkArity{\vec{v}}{\tau}} &
-    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
+    \left(\Eval{\Upsilon}{\tau}{N}{\App{\vartheta}{\Vec{F}}}
+    \quad
+    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}\right)
+    \quad
+    \text{supposing}
+    \quad
+    \left(
+      \Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}},
+      \Lookup{\CanOperators}{\vartheta}{\MkArity{\vec{v}}{\tau}}
+    \right)
   }
@@
 
+\begin{definition}[Computational Approximation]
 Because $[-_\tau]$ is monotonic, it has a greatest fixed point, which we will
-call \emph{applicative approximation},
-$\Of{\preceq_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
-written pointwise as $\SBinRel{\preceq_\tau}{\Upsilon}{M}{N}$.
+call \emph{computational approximation},
+$\Of{\preccurlyeq_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
+written pointwise as $\SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}$.
+\end{definition}
+
+\begin{definition}[Computational Equivalence]
+  Two terms are said to be \emph{computationally equivalent} when they approximate each other:
+  \[
+    \infer{
+      \SBinRel{\sim_\tau}{\Upsilon}{M}{N}
+    }{
+      \SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N} &
+      \SBinRel{\preccurlyeq_\tau}{\Upsilon}{N}{M}
+    }
+  \]
+\end{definition}
 

### 2015-11-20 paper.tex
Make definition of approximation precise
This is, incidentally, a case where having that presheaf machinery would
have been very useful.

@@
 Now, we can extend $R_\tau$ to a new relation $[R_\tau]$ on closed expressions
-which respects a single ``layer'' of computation, defined pointwise in the following way:
-\begin{gather*}
-  \infer{
-    \SBinRel{[R_\tau]}{\Upsilon}{M}{N}
-  }{
-    \left(\Eval{\Upsilon}{\tau}{N}{\App{\vartheta}{\Vec{F}}}
-    \quad
-    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}\right)
-    \quad
-    \text{supposing}
-    \quad
-    \left(
-      \Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}},
-      \Lookup{\CanOperators}{\vartheta}{\MkArity{\vec{v}}{\tau}}
-    \right)
-  }
-\end{gather*}
+which respects a single ``layer'' of computation. We will say
+$\SBinRel{[R_\tau]}{\Upsilon}{M}{N}$ when, supposing
+$\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}}$ such that
+$\Lookup{\CanOperators}{\vartheta}{\MkArity{\vec{v}}{\tau}}$, for any
+$\Of{\rho}{\Upsilon\inj\Upsilon'}$, we have
+$\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
+$\SBinRel{R_{\vec{v}}}{\Upsilon'}{\vec{E}\rho}{\vec{F}\rho}$.
 

### 2015-11-20 paper.tex
fix notation

@@
 $\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}}$ such that
-$\Lookup{\CanOperators}{\vartheta}{\MkArity{\vec{v}}{\tau}}$, for any
-$\Of{\rho}{\Upsilon\inj\Upsilon'}$, we have
+$\IMode{\Upsilon}\Vdash_{\IMode{\CanOperators}}\IMode{\vartheta}:\OMode{\MkArity{\vec{v}}{\tau}}$,
+for any $\Of{\rho}{\Upsilon\inj\Upsilon'}$, we have
 $\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and

### 2015-11-22 paper.tex
rename "logical framework" to "judgmental framework"

@@
 
-\section{The Ambient Logical Framework}
+\section{The Ambient Judgmental Framework}
 \label{section:ambient-lf}
@@
 For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
-deploy a generic logical framework equipped with higher-order judgments. A
+deploy a generic judgmental framework equipped with higher-order judgments. A
 judgment is always defined using a \emph{meaning explanation}, which is a
@@
 This apparatus is developed formally in Appendix~\ref{appendix:lf}, where the
-logical framework is cast as a topos of covariant presheaves on symbol contexts
+judgmental framework is cast as a topos of covariant presheaves on symbol contexts
 $\SCtx$.
@@
 \begin{appendices}
-\section{Formal Definition of the Logical Framework}
+\section{Formal Definition of the Judgmental Framework}
 \label{appendix:lf}
@@
 \emph{renaming convention}. We do so explicitly in this appendix by developing
-our logical framework as a presheaf topos over symbol contexts.
+our judgmental framework as a presheaf topos over symbol contexts.
 

### 2015-11-22 paper.tex
Comment out the bulk of the paper pending theory modifications

@@
 
-We will require two judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$,
-which means that $\Phi$ is a symbol context which refines the symbol context
-$\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
-$\Refines{\Phi}{\phi}{\tau}$ (which shall presuppose
-$\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
-
-\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
-\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
-
-The first judgment we define inductively:
-\begin{gather*}
-  \infer[\RefinesCtxNil]{
-    \RefinesCtx{\cdot}{\cdot}
-  }{
-  }\qquad
-  \infer[\RefinesCtxSnoc]{
-    \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
-  }{
-    \RefinesCtx{\Phi}{\Upsilon} &
-    \Refines{\Phi}{\phi}{\tau}
-  }
-\end{gather*}
-
-The second judgment is defined coinductively via the following meaning
-explanation:
-\begin{definition}[Parametric Refinement]
-  To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
-  $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
-  $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
-  what it means for $M$ and $N$ to be identified under $\phi$, supposing
-  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
-  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
-  $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
-  $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
-  $\IMode{M}\sim^\rho_\phi\IMode{O}$.
-\end{definition}
-
-In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
-$\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
-$\Upsilon'$, which partial equivalence relation on the closed $\tau$-sorted
-$\Upsilon'$-terms $\phi$ denotes. At this point, it should be remarked that
-this is very similar to the semantical explanation of typehood given in
-\cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
-setting, and that we have fibred the entire apparatus over collections
-$\Upsilon$ of symbols.
-
 \todo[inline]{%
-  define maps of refined contexts, then change weakening lemma to
-  monotonicity, which is stronger and more useful.
-}
-\begin{thm}[Weakening]
-  If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
-  $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
-\end{thm}
-\begin{proof}
-  This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
-  $\Upsilon\inj\Upsilon,u:\sigma$.
-\end{proof}
-
-When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ \emph{globally}
-refines $\tau$ and may simply write $\GRefines{\phi}{\tau}$. By weakening, if
-$\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
-$\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.
-
-\subsubsection{Ordering and Equality of Parametric Refinement}
-
-We will write $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ to mean that $\phi$ is a
-\emph{subrefinement} of $\psi$, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
-$\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$:
-\begin{definition}
-  To know $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
-  $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
-  \[
-    \GenJ{
-      \MV{m},\MV{n}:\tau
-    }{
-      \HypJ{
-        \IMode{M}\sim^\rho_\psi\IMode{N}
-      }{
-        \IMode{M}\sim^\rho_\phi\IMode{N}
-      }
-    }
-  \]
-\end{definition}
-
-Two refinements are equal when they denote the same PER. That is, we have
-$\EqRefines{\Phi}{\phi}{\psi}{\tau}$ just when both
-$\SubRefines{\Phi}{\phi}{\psi}{\tau}$ and $\SubRefines{\Phi}{\psi}{\phi}{\tau}$.
-
-\subsection{Parametric Equality}
-
-The primary judgment concerning refinements is the parametric equality,
-$\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
-$\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
-$\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
-$M$ and $N$ are identified by $\phi$ at the identity renaming
-$\Of{1}{\IMode{\Upsilon}\inj\OMode{\Upsilon}}$:
-\[
-  \infer{
-    \RMemEq{\Phi}{M}{N}{\phi}
-  }{
-    \IMode{M}\sim^1_\phi\IMode{N}
-  }
-\]
-
-\subsection{Functional Refinement}
-
-\newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
-\newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}
-
-Next, we define functional refinement
-$\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
-simultaneously with parametric context refinement
-$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and functional equality
-$\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$. We will write
-$\HRefines{\Phi}{\Psi}{\phi}{\tau}$ as a shorthand for
-$\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.
-
-\newcommand\DefPRefinesCtxNil[1]{
-  \infer[\PRefinesCtxNil]{
-    \PRefinesCtx{#1}{\cdot}{\cdot}
-  }{
-  }
-}
-\newcommand\DefPRefinesCtxSnoc[6]{
-  \infer[\PRefinesCtxSnoc]{
-    \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
-  }{
-    \PRefinesCtx{#1}{#2}{#5} &
-    \HRefines{#1}{#2}{#4}{#6}
-  }
+  Restore definitions of the refinement framework after they have been fixed!!
 }
 
-Parametric context refinement $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ shall be
-defined as follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$ and
-$\IsVarCtx{\Gamma}$:
-\begin{gather*}
-  \DefPRefinesCtxNil{\Phi}\qquad
-  \DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\phi}{\Gamma}{\tau}
-\end{gather*}
-
-Functional refinement $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ presupposes
-$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
-induction on the evidence for its first presupposition.
-
-\paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
-\[
-  \infer{
-    \HEqRefines{\Phi}{\cdot}{\phi}{\psi}{\tau}
-  }{
-    \EqRefines{\Phi}{\phi}{\psi}{\tau}
-  }
-\]
-
-\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$,
-where $\RefinesCtx{\Phi}{\Upsilon}$.
-
-\begin{gather*}
-  \infer{
-    \HEqRefines{\Phi}{\Psi,x:\chi}{\phi}{\psi}{\tau}
-  }{
-    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
-      \HypJ{
-        \HEqRefines{\Phi}{\Psi}{
-          \Subst{\MV{m}}{x}{\phi}
-        }{
-          \Subst{\MV{n}}{x}{\psi}
-        }{\tau}
-      }{
-        \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\chi}
-      }
-    }
-  }
-\end{gather*}
-
-\subsection{Functional Equality}
-
-Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
-$\RefinesCtx{\Phi}{\Upsilon}$, $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$,
-$\HRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
-and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
-induction on the evidence for the second presupposition
-$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
-
-\paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
-\[
-  \infer{
-    \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
-  }{
-    \RMemEq{\Phi}{M}{N}{\phi}
-  }
-\]
-
-\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
-where $\RefinesCtx{\Phi}{\Upsilon}$.
-\[
-  \infer{
-    \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
-  }{
-    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
-      \HypJ{
-        \HRMemEq{\Phi}{\Psi}{
-          \Subst{\MV{m}}{x}{\phi}
-        }{
-          \Subst{\MV{n}}{x}{\psi}
-        }{\phi}
-      }{
-        \HRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
-      }
-    }
-  }
-\]
-
-\section{Constructions on Refinements}
-
-Refinements can be arranged in a partial order via the
-$\SubRefines{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
-at any sort $\tau$.
-
-We have the bottom refinement $\GRefines{\TyBot{\tau}}{\tau}$, which simply
-contains no elements. Next, the top refinement $\GRefines{\TyTop{\tau}}{\tau}$
-identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
-and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
-$\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
-$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
-\begin{gather*}
-  \infer{
-    \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
-  }{
-    \RMemEq{\Phi}{M}{N}{\phi}
-  }\qquad
-  \infer{
-    \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
-  }{
-    \RMemEq{\Phi}{M}{N}{\psi}
-  }
-\end{gather*}
-
-Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
-$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
-dually:
-\[
-  \infer{
-    \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
-  }{
-    \RMemEq{\Phi}{M}{N}{\phi} &
-    \RMemEq{\Phi}{M}{N}{\psi}
-  }
-\]
-
-It is trivial to verify that these constructions on refinements give rise to a
-lattice.
-
-\section{Multi-Sorted Computational Type Theory}
-
-Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
-you specify how to form its equal canonical inhabitants, or values. Then, the
-inhabitants of the types are explained uniformly by extending this relation
-over an operational semantics such that two terms are equal members of a
-type just when they compute to equal values of that type.
-
-More generally, one may wish to consider types which include computations in
-addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
-is a type, which would include all the members of $\psi$ as well as a divergent
-computation. Another example is the type $\TyTop{\tau}$ which identifies all
-closed terms of sort $\tau$.
-
-In order to ensure that such types are sensible, we must formulate a notion of
-computational equivalence, and then say that a refinement is a \emph{type} just
-when it respects computational equivalence; then, we are able to define types
-as PERs on \emph{programs} in addition to values.
+%% \subsection{Parametric Refinement}
+%%
+%% We will require two judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$,
+%% which means that $\Phi$ is a symbol context which refines the symbol context
+%% $\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
+%% $\Refines{\Phi}{\phi}{\tau}$ (which shall presuppose
+%% $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
+%%
+%% \newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
+%% \newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
+%%
+%% The first judgment we define inductively:
+%% \begin{gather*}
+%%   \infer[\RefinesCtxNil]{
+%%     \RefinesCtx{\cdot}{\cdot}
+%%   }{
+%%   }\qquad
+%%   \infer[\RefinesCtxSnoc]{
+%%     \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
+%%   }{
+%%     \RefinesCtx{\Phi}{\Upsilon} &
+%%     \Refines{\Phi}{\phi}{\tau}
+%%   }
+%% \end{gather*}
+%%
+%% The second judgment is defined coinductively via the following meaning
+%% explanation:
+%% \begin{definition}[Parametric Refinement]
+%%   To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
+%%   $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
+%%   $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
+%%   what it means for $M$ and $N$ to be identified under $\phi$, supposing
+%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
+%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
+%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
+%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
+%%   $\IMode{M}\sim^\rho_\phi\IMode{O}$.
+%% \end{definition}
+%%
+%% In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
+%% $\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
+%% $\Upsilon'$, which partial equivalence relation on the closed $\tau$-sorted
+%% $\Upsilon'$-terms $\phi$ denotes. At this point, it should be remarked that
+%% this is very similar to the semantical explanation of typehood given in
+%% \cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
+%% setting, and that we have fibred the entire apparatus over collections
+%% $\Upsilon$ of symbols.
+%%
+%% \todo[inline]{%
+%%   define maps of refined contexts, then change weakening lemma to
+%%   monotonicity, which is stronger and more useful.
+%% }
+%% \begin{thm}[Weakening]
+%%   If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
+%%   $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
+%% \end{thm}
+%% \begin{proof}
+%%   This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
+%%   $\Upsilon\inj\Upsilon,u:\sigma$.
+%% \end{proof}
+%%
+%% When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ \emph{globally}
+%% refines $\tau$ and may simply write $\GRefines{\phi}{\tau}$. By weakening, if
+%% $\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
+%% $\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.
+%%
+%% \subsubsection{Ordering and Equality of Parametric Refinement}
+%%
+%% We will write $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ to mean that $\phi$ is a
+%% \emph{subrefinement} of $\psi$, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
+%% $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$:
+%% \begin{definition}
+%%   To know $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
+%%   $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
+%%   \[
+%%     \GenJ{
+%%       \MV{m},\MV{n}:\tau
+%%     }{
+%%       \HypJ{
+%%         \IMode{M}\sim^\rho_\psi\IMode{N}
+%%       }{
+%%         \IMode{M}\sim^\rho_\phi\IMode{N}
+%%       }
+%%     }
+%%   \]
+%% \end{definition}
+%%
+%% Two refinements are equal when they denote the same PER. That is, we have
+%% $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ just when both
+%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ and $\SubRefines{\Phi}{\psi}{\phi}{\tau}$.
+%%
+%% \subsection{Parametric Equality}
+%%
+%% The primary judgment concerning refinements is the parametric equality,
+%% $\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
+%% $\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
+%% $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
+%% $M$ and $N$ are identified by $\phi$ at the identity renaming
+%% $\Of{1}{\IMode{\Upsilon}\inj\OMode{\Upsilon}}$:
+%% \[
+%%   \infer{
+%%     \RMemEq{\Phi}{M}{N}{\phi}
+%%   }{
+%%     \IMode{M}\sim^1_\phi\IMode{N}
+%%   }
+%% \]
+%%
+%% \subsection{Functional Refinement}
+%%
+%% \newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
+%% \newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}
+%%
+%% Next, we define functional refinement
+%% $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
+%% simultaneously with parametric context refinement
+%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and functional equality
+%% $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$. We will write
+%% $\HRefines{\Phi}{\Psi}{\phi}{\tau}$ as a shorthand for
+%% $\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.
+%%
+%% \newcommand\DefPRefinesCtxNil[1]{
+%%   \infer[\PRefinesCtxNil]{
+%%     \PRefinesCtx{#1}{\cdot}{\cdot}
+%%   }{
+%%   }
+%% }
+%% \newcommand\DefPRefinesCtxSnoc[6]{
+%%   \infer[\PRefinesCtxSnoc]{
+%%     \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
+%%   }{
+%%     \PRefinesCtx{#1}{#2}{#5} &
+%%     \HRefines{#1}{#2}{#4}{#6}
+%%   }
+%% }
+%%
+%% Parametric context refinement $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ shall be
+%% defined as follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$ and
+%% $\IsVarCtx{\Gamma}$:
+%% \begin{gather*}
+%%   \DefPRefinesCtxNil{\Phi}\qquad
+%%   \DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\phi}{\Gamma}{\tau}
+%% \end{gather*}
+%%
+%% Functional refinement $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ presupposes
+%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
+%% induction on the evidence for its first presupposition.
+%%
+%% \paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
+%% \[
+%%   \infer{
+%%     \HEqRefines{\Phi}{\cdot}{\phi}{\psi}{\tau}
+%%   }{
+%%     \EqRefines{\Phi}{\phi}{\psi}{\tau}
+%%   }
+%% \]
+%%
+%% \paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$,
+%% where $\RefinesCtx{\Phi}{\Upsilon}$.
+%%
+%% \begin{gather*}
+%%   \infer{
+%%     \HEqRefines{\Phi}{\Psi,x:\chi}{\phi}{\psi}{\tau}
+%%   }{
+%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
+%%       \HypJ{
+%%         \HEqRefines{\Phi}{\Psi}{
+%%           \Subst{\MV{m}}{x}{\phi}
+%%         }{
+%%           \Subst{\MV{n}}{x}{\psi}
+%%         }{\tau}
+%%       }{
+%%         \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\chi}
+%%       }
+%%     }
+%%   }
+%% \end{gather*}
+%%
+%% \subsection{Functional Equality}
+%%
+%% Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
+%% $\RefinesCtx{\Phi}{\Upsilon}$, $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$,
+%% $\HRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
+%% and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
+%% induction on the evidence for the second presupposition
+%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
+%%
+%% \paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
+%% \[
+%%   \infer{
+%%     \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
+%%   }{
+%%     \RMemEq{\Phi}{M}{N}{\phi}
+%%   }
+%% \]
+%%
+%% \paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
+%% where $\RefinesCtx{\Phi}{\Upsilon}$.
+%% \[
+%%   \infer{
+%%     \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
+%%   }{
+%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
+%%       \HypJ{
+%%         \HRMemEq{\Phi}{\Psi}{
+%%           \Subst{\MV{m}}{x}{\phi}
+%%         }{
+%%           \Subst{\MV{n}}{x}{\psi}
+%%         }{\phi}
+%%       }{
+%%         \HRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
+%%       }
+%%     }
+%%   }
+%% \]
+%%
+%% \section{Constructions on Refinements}
+%%
+%% Refinements can be arranged in a partial order via the
+%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
+%% at any sort $\tau$.
+%%
+%% We have the bottom refinement $\GRefines{\TyBot{\tau}}{\tau}$, which simply
+%% contains no elements. Next, the top refinement $\GRefines{\TyTop{\tau}}{\tau}$
+%% identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
+%% and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
+%% $\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
+%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
+%% \begin{gather*}
+%%   \infer{
+%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
+%%   }{
+%%     \RMemEq{\Phi}{M}{N}{\phi}
+%%   }\qquad
+%%   \infer{
+%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
+%%   }{
+%%     \RMemEq{\Phi}{M}{N}{\psi}
+%%   }
+%% \end{gather*}
+%%
+%% Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
+%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
+%% dually:
+%% \[
+%%   \infer{
+%%     \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
+%%   }{
+%%     \RMemEq{\Phi}{M}{N}{\phi} &
+%%     \RMemEq{\Phi}{M}{N}{\psi}
+%%   }
+%% \]
+%%
+%% It is trivial to verify that these constructions on refinements give rise to a
+%% lattice.
+%%
+%% \section{Multi-Sorted Computational Type Theory}
+%%
+%% Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
+%% you specify how to form its equal canonical inhabitants, or values. Then, the
+%% inhabitants of the types are explained uniformly by extending this relation
+%% over an operational semantics such that two terms are equal members of a
+%% type just when they compute to equal values of that type.
+%%
+%% More generally, one may wish to consider types which include computations in
+%% addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
+%% is a type, which would include all the members of $\psi$ as well as a divergent
+%% computation. Another example is the type $\TyTop{\tau}$ which identifies all
+%% closed terms of sort $\tau$.
+%%
+%% In order to ensure that such types are sensible, we must formulate a notion of
+%% computational equivalence, and then say that a refinement is a \emph{type} just
+%% when it respects computational equivalence; then, we are able to define types
+%% as PERs on \emph{programs} in addition to values.
 

### 2015-11-22 paper.tex
Define parametric refinement (#7)

@@
 framework, we will define the notion of behavioral refinement by propounding
-several judgments and their semantical explanations.
+several judgments and their semantical explanations. Let us first define the
+presheaf of $\tau$-sorted expressions, $\Exprs_\tau$ as follows
+\[
+  \Define{
+    \Exprs_\tau(\Upsilon\parallel\Gamma)
+  }{
+    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\Gamma}{M}{\tau}}
+  }
+\]
+
+We will also write $\Exprs_\tau$ for the presheaf $\Exprs_\tau(-\parallel\cdot)$ on $\SCtx$.
 
 \subsection{Parametric Refinement}
+
+The first judgment that will concern us is called \emph{parametric refinement},
+$\Refines{\Upsilon}{\phi}{\tau}$, which means that $\phi$ refines the sort
+$\tau$ under the symbolic parameters $\Upsilon$. We define this judgment
+through a meaning explanation in the style of Martin-L\"of as follows:
+
+\newcommand\RefEquate[4]{#1\{\IMode{#2}\}\left(\IMode{#3},\IMode{#4}\right)}
+
+\begin{definition}
+  To know $\Refines{\Upsilon}{\phi}{\tau}$ (presupposing $\IsSymCtx{\Upsilon}$
+  and $\IsSort{\tau}$) is to know, for any $\Of{\rho}{\Upsilon\inj\Upsilon'}$,
+  what it means for any $\Member{M,N}{\Exprs_\tau(\Upsilon')}$ to be equated by
+  $\phi$ (written $\RefEquate{\phi}{\rho}{M}{N}$), such that
+  $\RefEquate{\phi}{\rho}{\cdot}{\cdot}$ is a partial equivalence relation on
+  $\Exprs_\tau(\Upsilon')$. Morever, that for any
+  $\Of{\rho'}{\Upsilon'\inj\Upsilon''}$, from $\RefEquate{\phi}{\rho}{M}{N}$ we
+  may conclude $\RefEquate{\phi}{\rho'\circ\rho}{M\rho'}{N\rho'}$.
+\end{definition}
 

### 2015-11-22 paper.tex
Remark on why pointwise defs are so gnarly, show clean def in the topos
(#7)

@@
 \end{definition}
+
+Then, we say that $\phi$ \emph{globally refines} $\tau$ (written
+$\GRefines{\phi}{\tau}$) when for all $\Upsilon$, we have
+$\Refines{\Upsilon}{\phi}{\tau}$.
+
+\begin{remark}
+
+  At this point, it should be noted that this is very similar to the semantical
+  explanation of typehood given in \cite{martin-lof:1979}, except that we have
+  generalized it to a multi-sorted setting, and that we have fibred the entire
+  apparatus over collections $\Upsilon$ of symbols.
+
+  In fact, the complexity of the above meaning explanation is an artifact of
+  the pointwise style in which we have expressed it. Considered internally to
+  the presheaf topos $\Sets^\SCtx$, a refinement is merely a section of the
+  object of $\mathfrak{per}(\Exprs_\tau)$, defined internally as a subobject of
+  $\Omega^{\Exprs_\tau\times\Exprs_\tau}$:
+  \[
+    \Define{
+      \mathfrak{per}(X)
+    }{
+      \{ \phi : \Omega^{X\times X}
+         \mid
+         symmetric(\phi)
+         \land transitive(\phi)
+      \}
+    }
+  \]
+
+  So, $\Refines{\Upsilon}{\phi}{\tau}$ obtains just when we have
+  $\Member{\phi}{\mathfrak{per}(\Exprs_\tau)(\Upsilon)}$. The benefit of
+  explaining such judgments in terms of the presheaf topos is that we do not
+  need to deal with the complexities of quantifying over renamings, since this
+  is implicit in the definition of the exponential of presheaves,
+  \begin{align*}
+    \ADefine{
+      B^A(\Upsilon)
+    }{
+      \Hom{\Sets^\SCtx}{
+        \Yoneda{\Upsilon}\times A
+      }{
+        B
+      }
+    }\\
+    &\cong
+    \OMode{
+      \int_{\Upsilon'}
+      (\Upsilon\inj\Upsilon')
+      \Rightarrow
+      {B(\Upsilon')}^{A(\Upsilon')}
+    }
+  \end{align*}
+
+\end{remark}
 

### 2015-11-22 paper.tex
typos

@@
   the presheaf topos $\Sets^\SCtx$, a refinement is merely a section of the
-  object of $\mathfrak{per}(\Exprs_\tau)$, defined internally as a subobject of
-  $\Omega^{\Exprs_\tau\times\Exprs_\tau}$:
+  object of $\Exprs_\tau$-PERs $\mathfrak{per}(\Exprs_\tau)$, defined
+  internally as a subobject of $\Omega^{\Exprs_\tau\times\Exprs_\tau}$:
   \[

### 2015-11-22 paper.tex
Define subrefinement relation, equality of refinements

@@
   To know $\Refines{\Upsilon}{\phi}{\tau}$ (presupposing $\IsSymCtx{\Upsilon}$
-  and $\IsSort{\tau}$) is to know, for any $\Of{\rho}{\Upsilon\inj\Upsilon'}$,
+  and $\IsSort{\tau}$) is to know, for any
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd},
   what it means for any $\Member{M,N}{\Exprs_\tau(\Upsilon')}$ to be equated by
@@
   $\Exprs_\tau(\Upsilon')$. Morever, that for any
-  $\Of{\rho'}{\Upsilon'\inj\Upsilon''}$, from $\RefEquate{\phi}{\rho}{M}{N}$ we
-  may conclude $\RefEquate{\phi}{\rho'\circ\rho}{M\rho'}{N\rho'}$.
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon'}\arrow[r,hook,"\IMode{\rho'}"] &\IMode{\Upsilon''}
+  \end{tikzcd},
+  from $\RefEquate{\phi}{\rho}{M}{N}$ we may conclude
+  $\RefEquate{\phi}{\rho'\circ\rho}{M\rho'}{N\rho'}$.
 \end{definition}
@@
 
+\subsubsection{Order and Equality of Parametric Refinements}
+
+We will write $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ to mean that $\phi$ is
+a \emph{subrefinement} of $\psi$.
+
+\begin{definition}
+
+  To know $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ (presupposing
+  $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$) , is
+  to know, for any renamings
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon} \arrow[r, hook, "\IMode{\rho}"]
+    & \IMode{\Upsilon'} \arrow[r, hook, "\IMode{\rho'}"]
+    & \IMode{\Upsilon''}
+  \end{tikzcd},
+  that from $\RefEquate{\phi}{\rho}{M}{N}$ you can conclude
+  $\RefEquate{\psi}{\rho'\circ\rho}{M\rho'}{N\rho'}$ for any
+  $\Member{M,N}{\Exprs_\tau(\Upsilon')}$.
+
+\end{definition}
+
+Two refinements are equal when they denote the same PER. That is, we have
+$\EqRefines{\Upsilon}{\phi}{\psi}{\tau}$ just when both
+$\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ and
+$\SubRefines{\Upsilon}{\psi}{\phi}{\tau}$.
+
 \todo[inline]{%
@@
 
-%% \subsection{Parametric Refinement}
-%%
-%% We will require two judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$,
-%% which means that $\Phi$ is a symbol context which refines the symbol context
-%% $\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
-%% $\Refines{\Phi}{\phi}{\tau}$ (which shall presuppose
-%% $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
-%%
-%% \newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
-%% \newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
-%%
-%% The first judgment we define inductively:
-%% \begin{gather*}
-%%   \infer[\RefinesCtxNil]{
-%%     \RefinesCtx{\cdot}{\cdot}
-%%   }{
-%%   }\qquad
-%%   \infer[\RefinesCtxSnoc]{
-%%     \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
-%%   }{
-%%     \RefinesCtx{\Phi}{\Upsilon} &
-%%     \Refines{\Phi}{\phi}{\tau}
-%%   }
-%% \end{gather*}
-%%
-%% The second judgment is defined coinductively via the following meaning
-%% explanation:
-%% \begin{definition}[Parametric Refinement]
-%%   To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
-%%   $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
-%%   $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
-%%   what it means for $M$ and $N$ to be identified under $\phi$, supposing
-%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
-%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
-%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
-%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
-%%   $\IMode{M}\sim^\rho_\phi\IMode{O}$.
-%% \end{definition}
-%%
-%% In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
-%% $\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
-%% $\Upsilon'$, which partial equivalence relation on the closed $\tau$-sorted
-%% $\Upsilon'$-terms $\phi$ denotes. At this point, it should be remarked that
-%% this is very similar to the semantical explanation of typehood given in
-%% \cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
-%% setting, and that we have fibred the entire apparatus over collections
-%% $\Upsilon$ of symbols.
-%%
-%% \todo[inline]{%
-%%   define maps of refined contexts, then change weakening lemma to
-%%   monotonicity, which is stronger and more useful.
-%% }
-%% \begin{thm}[Weakening]
-%%   If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
-%%   $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
-%% \end{thm}
-%% \begin{proof}
-%%   This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
-%%   $\Upsilon\inj\Upsilon,u:\sigma$.
-%% \end{proof}
-%%
-%% When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ \emph{globally}
-%% refines $\tau$ and may simply write $\GRefines{\phi}{\tau}$. By weakening, if
-%% $\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
-%% $\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.
-%%
-%% \subsubsection{Ordering and Equality of Parametric Refinement}
-%%
-%% We will write $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ to mean that $\phi$ is a
-%% \emph{subrefinement} of $\psi$, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
-%% $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$:
-%% \begin{definition}
-%%   To know $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
-%%   $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
-%%   \[
-%%     \GenJ{
-%%       \MV{m},\MV{n}:\tau
-%%     }{
-%%       \HypJ{
-%%         \IMode{M}\sim^\rho_\psi\IMode{N}
-%%       }{
-%%         \IMode{M}\sim^\rho_\phi\IMode{N}
-%%       }
-%%     }
-%%   \]
-%% \end{definition}
-%%
-%% Two refinements are equal when they denote the same PER. That is, we have
-%% $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ just when both
-%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ and $\SubRefines{\Phi}{\psi}{\phi}{\tau}$.
 %%
@@
 $\IMode{\Upsilon}\Vdash_{\IMode{\CanOperators}}\IMode{\vartheta}:\OMode{\MkArity{\vec{v}}{\tau}}$,
-for any $\Of{\rho}{\Upsilon\inj\Upsilon'}$, we have
-$\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
+for any
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+\end{tikzcd},
+we have $\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
 $\SBinRel{R_{\vec{v}}}{\Upsilon'}{\vec{E}\rho}{\vec{F}\rho}$.

### 2015-11-22 paper.tex
Restore equality of parametric refinements, functional refinement

@@
 
-\todo[inline]{%
-  Restore definitions of the refinement framework after they have been fixed!!
+\subsection{Parametric Equality}
+
+The primary judgment concerning refinements is the parametric equality,
+$\RMemEq{\Upsilon}{M}{N}{\phi}$, which presupposes
+$\Refines{\Upsilon}{\phi}{\tau}$, and $\Member{M,N}{\Exprs_\tau(\Upsilon)}$.
+The meaning of this judgment is that $M$ and $N$ are identified by $\phi$ at $\Upsilon$:
+\[
+  \infer{
+    \RMemEq{\Upsilon}{M}{N}{\phi}
+  }{
+    \RefEquate{\phi}{1}{M}{N}
+  }
+\]
+
+\subsection{Functional Refinement}
+
+\newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
+\newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}
+
+Next, we define functional refinement
+$\HEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
+simultaneously with parametric context refinement
+$\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and functional equality
+$\HRMemEq{\Upsilon}{\Psi}{M}{N}{\phi}$. We will write
+$\HRefines{\Upsilon}{\Psi}{\phi}{\tau}$ as a shorthand for
+$\HEqRefines{\Upsilon}{\Psi}{\phi}{\phi}{\tau}$.
+
+\newcommand\DefPRefinesCtxNil[1]{
+  \infer[\PRefinesCtxNil]{
+    \PRefinesCtx{#1}{\cdot}{\cdot}
+  }{
+  }
+}
+\newcommand\DefPRefinesCtxSnoc[6]{
+  \infer[\PRefinesCtxSnoc]{
+    \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
+  }{
+    \PRefinesCtx{#1}{#2}{#5} &
+    \HRefines{#1}{#2}{#4}{#6}
+  }
 }
 
-%%
-%% \subsection{Parametric Equality}
-%%
-%% The primary judgment concerning refinements is the parametric equality,
-%% $\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
-%% $\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
-%% $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
-%% $M$ and $N$ are identified by $\phi$ at the identity renaming
-%% $\Of{1}{\IMode{\Upsilon}\inj\OMode{\Upsilon}}$:
-%% \[
-%%   \infer{
-%%     \RMemEq{\Phi}{M}{N}{\phi}
-%%   }{
-%%     \IMode{M}\sim^1_\phi\IMode{N}
-%%   }
-%% \]
-%%
-%% \subsection{Functional Refinement}
-%%
-%% \newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
-%% \newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}
-%%
-%% Next, we define functional refinement
-%% $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
-%% simultaneously with parametric context refinement
-%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and functional equality
-%% $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$. We will write
-%% $\HRefines{\Phi}{\Psi}{\phi}{\tau}$ as a shorthand for
-%% $\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.
-%%
-%% \newcommand\DefPRefinesCtxNil[1]{
-%%   \infer[\PRefinesCtxNil]{
-%%     \PRefinesCtx{#1}{\cdot}{\cdot}
-%%   }{
-%%   }
-%% }
-%% \newcommand\DefPRefinesCtxSnoc[6]{
-%%   \infer[\PRefinesCtxSnoc]{
-%%     \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
-%%   }{
-%%     \PRefinesCtx{#1}{#2}{#5} &
-%%     \HRefines{#1}{#2}{#4}{#6}
-%%   }
-%% }
-%%
-%% Parametric context refinement $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ shall be
-%% defined as follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$ and
-%% $\IsVarCtx{\Gamma}$:
-%% \begin{gather*}
-%%   \DefPRefinesCtxNil{\Phi}\qquad
-%%   \DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\phi}{\Gamma}{\tau}
-%% \end{gather*}
-%%
-%% Functional refinement $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ presupposes
-%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
-%% induction on the evidence for its first presupposition.
-%%
-%% \paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
-%% \[
-%%   \infer{
-%%     \HEqRefines{\Phi}{\cdot}{\phi}{\psi}{\tau}
-%%   }{
-%%     \EqRefines{\Phi}{\phi}{\psi}{\tau}
-%%   }
-%% \]
-%%
-%% \paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$,
-%% where $\RefinesCtx{\Phi}{\Upsilon}$.
-%%
-%% \begin{gather*}
-%%   \infer{
-%%     \HEqRefines{\Phi}{\Psi,x:\chi}{\phi}{\psi}{\tau}
-%%   }{
-%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
-%%       \HypJ{
-%%         \HEqRefines{\Phi}{\Psi}{
-%%           \Subst{\MV{m}}{x}{\phi}
-%%         }{
-%%           \Subst{\MV{n}}{x}{\psi}
-%%         }{\tau}
-%%       }{
-%%         \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\chi}
-%%       }
-%%     }
-%%   }
-%% \end{gather*}
+Parametric context refinement $\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ shall be
+defined as follows, presupposing $\IsSymCtx{\Upsilon}$ and
+$\IsVarCtx{\Gamma}$:
+\begin{gather*}
+  \DefPRefinesCtxNil{\Upsilon}\qquad
+  \DefPRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\phi}{\Gamma}{\tau}
+\end{gather*}
+
+Functional refinement $\HEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ presupposes
+$\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
+induction on the evidence for its first presupposition.
+
+\paragraph{Case} $\DefPRefinesCtxNil{\Upsilon}$.
+\[
+  \infer{
+    \HEqRefines{\Upsilon}{\cdot}{\phi}{\psi}{\tau}
+  }{
+    \EqRefines{\Upsilon}{\phi}{\psi}{\tau}
+  }
+\]
+
+\paragraph{Case} $\DefPRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
+
+\begin{gather*}
+  \infer{
+    \HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
+  }{
+    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
+      \HypJ{
+        \HEqRefines{\Upsilon}{\Psi}{
+          \Subst{\MV{m}}{x}{\phi}
+        }{
+          \Subst{\MV{n}}{x}{\psi}
+        }{\tau}
+      }{
+        \HRMemEq{\Upsilon}{\Psi}{\MV{n}}{\MV{n}}{\chi}
+      }
+    }
+  }
+\end{gather*}
+
+\todo[inline]{%
+  Is the above a correct rule? I think that it might be good to be explicit
+  here rather than using the hypothetico-general judgment. In fact, if we are
+  explicit about everything, we can get away with not even explaining the
+  higher-order judgments.
+}
+
 %%

### 2015-11-22 paper.tex
improve definition of parametric equality

@@
 $\RMemEq{\Upsilon}{M}{N}{\phi}$, which presupposes
-$\Refines{\Upsilon}{\phi}{\tau}$, and $\Member{M,N}{\Exprs_\tau(\Upsilon)}$.
-The meaning of this judgment is that $M$ and $N$ are identified by $\phi$ at $\Upsilon$:
+$\Refines{\Upsilon'}{\phi}{\tau}$ for some $\Upsilon'$ such that we have
+\begin{tikzcd}[cramped, sep = small]
+  \IMode{\Upsilon'} \arrow[r, hook, "\IMode{\rho}"] &\IMode{\Upsilon}
+\end{tikzcd},
+and $\Member{M,N}{\Exprs_\tau(\Upsilon)}$. The meaning of this judgment is
+that $M$ and $N$ are identified by $\phi$ at $\Upsilon$:
 \[
@@
   }{
-    \RefEquate{\phi}{1}{M}{N}
+    \RefEquate{\phi}{\rho}{M}{N}
   }

### 2015-11-22 paper.tex
improve rule labels
Make the definition of functional refinement more precise

@@
 Then, we say that $\phi$ \emph{globally refines} $\tau$ (written
-$\GRefines{\phi}{\tau}$) when for all $\Upsilon$, we have
-$\Refines{\Upsilon}{\phi}{\tau}$.
+$\GRefines{\phi}{\tau}$) when we have $\Refines{\cdot}{\phi}{\tau}$.
+Furthermore, when $\Refines{\Upsilon}{\phi}{\tau}$, for any renaming
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+\end{tikzcd}
+we can define a new refinement $\phi\rho$ such that
+$\Refines{\Upsilon'}{\phi\rho}{\tau}$ in the obvious way.
 
@@
 
-\newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
-\newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}
+\newcommand\PRefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
+\newcommand\PRefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
+\newcommand\PFunRefNil{\vDash^\sqsubset_{\mathtt{nil}}}
+\newcommand\PFunRefSnoc{\vDash^\sqsubset_{\mathtt{snoc}}}
 
@@
 
+For a context refinement $\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and a renaming
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+\end{tikzcd},
+we can define a new context refinement
+$\PRefinesCtx{\Upsilon'}{\Psi\rho}{\Gamma}$ by applying $\rho$ pointwise at
+each of the sort refinements in the context.
+
 Functional refinement $\HEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ presupposes
@@
 \[
-  \infer{
+  \infer[\PFunRefNil]{
     \HEqRefines{\Upsilon}{\cdot}{\phi}{\psi}{\tau}
@@
 
-\begin{gather*}
-  \infer{
-    \HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
-  }{
-    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
-      \HypJ{
-        \HEqRefines{\Upsilon}{\Psi}{
-          \Subst{\MV{m}}{x}{\phi}
-        }{
-          \Subst{\MV{n}}{x}{\psi}
-        }{\tau}
-      }{
-        \HRMemEq{\Upsilon}{\Psi}{\MV{n}}{\MV{n}}{\chi}
-      }
-    }
-  }
-\end{gather*}
+\begin{quote}
+  To know $\HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}$ is to know, for any renaming
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}
+  and closed terms $\Member{M,N}{\Exprs_\sigma(\Upsilon')}$, that from
+  $\HRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}$, you can conclude
+  $\HRMemEq{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
 
-\todo[inline]{%
-  Is the above a correct rule? I think that it might be good to be explicit
-  here rather than using the hypothetico-general judgment. In fact, if we are
-  explicit about everything, we can get away with not even explaining the
-  higher-order judgments.
-}
+\end{quote}
 

### 2015-11-22 paper.tex
Entirely remove the ambient logical framework, higher order judgments
I think that things will be clearer if we are fully explicit about the
quantification over renamings, even if we could factor it out into the
presheaf machinery.

@@
 
-\section{Abstract Binding Trees and Symbols}
-
-See~\cite{sterling-morrison:2015} for the development of abstract binding trees
-with symbols. \todo[inline]{give a brief description of the framework, and present its
-rules.}
-
-\section{The Ambient Judgmental Framework}
-\label{section:ambient-lf}
-
 \ifcolored%
+\paragraph{Colors}
 In this paper, we hint at the \emph{modes} of judgments and
@@
 
-For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
-deploy a generic judgmental framework equipped with higher-order judgments. A
-judgment is always defined using a \emph{meaning explanation}, which is a
-specification of the conditions under which it may be asserted (sc.\ its
-evidence) \cite{martin-lof:1996}.
+\section{Abstract Binding Trees and Symbols}
 
-\paragraph{Renaming convention} Because the evidence of a judgment will often
-contain symbols $u,v,\dots$ we informally adopt the convention that the
-evidence of a judgment shall always be subject to fresh renamings of its free
-symbols (i.e.\ the validity of a judgment shall not depend on choice of names).
-This apparatus is developed formally in Appendix~\ref{appendix:lf}, where the
-judgmental framework is cast as a topos of covariant presheaves on symbol contexts
-$\SCtx$.
-
-\begin{definition}[Hypothetical Judgment]
-  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
-  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$, whose evidence
-  shall be an effective transformation of any evidence for $\mathcal{J}_1$ into
-  evidence for $\mathcal{J}_2$. It is important to keep in mind that this is a
-  \emph{semantic} consequence and therefore expresses admissibilities, in
-  contrast to the \emph{logical} consequence which expresses
-  derivabilities~\cite{harper:2016}. We will not have any need for the latter.
-
-  \paragraph{Notation}
-  For a hypothetical judgment with multiple hypotheses, we will write
-  $\HypJ{\mathcal{J}_3}{\mathcal{J}_1, \mathcal{J}_2}$ rather than
-  $\HypJ{\mathcal{J}_3}{\mathcal{J}_1\times\mathcal{J}_2}$ or
-  $\HypJ{(\HypJ{\mathcal{J}_3}{\mathcal{J}_2})}{\mathcal{J}_1}$.
-\end{definition}
-
-\begin{definition}[General Judgment]
-  For a metavariable context $\IsMetaCtx{\Theta}$, we can define the set
-  of its substitutions as follows:
-  \[
-    \Define{{\square\Theta}}{
-      \prod_{(\MV{m}:v)\in\Theta}
-        \MkSet{E\mid\IsAbtUnmoded{\cdot}{\cdot}{\cdot}{E}{v}}
-    }
-  \]
-
-  For any $\square{\Theta}$-indexed family of judgments $\mathcal{J}$, we can
-  express the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as being evident
-  when $\mathcal{J}(\vec{E})$ is evident for each
-  $\Of{\vec{E}}{\square\Theta}$; in other words, the evidence of the general
-  judgment is an effective transformation of $\Theta$-substitutions to their
-  fibres in $\mathcal{J}$.  Recall that metavariables have \emph{valences}
-  rather than sorts; as such, the general judgment allows us to quantify over
-  terms of higher type.
-
-   \paragraph{Notation}
-
-   We will write $\HPGenJ{\Upsilon}{\Gamma}{\Theta}{\mathcal{J}}$ for
-   $\GenJ{\Theta'}{\mathcal{J}}$, where
-   $\Lookup{\Theta'}{\MV{m}}{\MkValence{\Upsilon,\vec{\sigma}}{\Gamma,\vec{\tau}}{\tau}}$
-   if $\Lookup{\Theta}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$.
-   Additionally, for clarity we will often write
-   $\GenJ{\MV{m}:\sigma}{\mathcal{J}}$ rather than
-   $\GenJ{\MV{m}:\MkValence{\cdot}{\cdot}{\sigma}}{\mathcal{J}}$; likewise, in
-   the body of $\mathcal{J}$, when there are contextually salient
-   $\vec{u}$, $\vec{x}$, we will simply write $\MV{m}$ rather than the more proper
-   $\MApp{\MV{m}}{\vec{u}}{\vec{x}}$.
-
-\end{definition}
-
-After this section, we will define a judgment either by rules, or by an
-informal semantical explanation, in both cases leaving its intensional
-character implicit. A judgment is said to be \emph{correct} or \emph{valid}
-just in case it is globally inhabited.
+See~\cite{sterling-morrison:2015} for the development of abstract binding trees
+with symbols. \todo[inline]{give a brief description of the framework, and present its
+rules.}
 
@@
 
-\newpage
-\begin{appendices}
-\section{Formal Definition of the Judgmental Framework}
-\label{appendix:lf}
-
-As noted earlier, we must be cautious in the definitions of both atomic and
-higher-order judgments that their evidence shall always respect the
-\emph{renaming convention}. We do so explicitly in this appendix by developing
-our judgmental framework as a presheaf topos over symbol contexts.
-
-A judgment is, then, an object in the presheaf category
-$\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, judgments are identified
-with the intensional set of its renameable evidence. $\mathbb{J}$, like all
-presheaf categories, is a topos. The task at hand, then, is to show how to
-replace the informal meaning explanations that we gave in
-section~\ref{section:ambient-lf} with precise definitions in the topos
-$\mathbb{J}$.
-
-\subsection{Hypothetical Judgment}
-
-Hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is defined using the
-exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
-the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
-\begin{align*}
-  \ADefine{
-    \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
-  }{
-    \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
-  }\\
-  &\cong
-  \OMode{
-    \Hom{\mathbb{J}}{
-      \Yoneda{\Upsilon}\times\mathcal{J}_1
-    }{
-      \mathcal{J}_2
-    }
-  }\\
-  &\cong
-  \OMode{
-    \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
-      \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
-  }
-\end{align*}
-
-In other words, the hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$
-is evident when there shall ``forevermore'' (i.e. for any applicable renaming
-of symbols) be an effective transformation of evidence for the antecedent into
-evidence for the consequent.
-
-\subsection{General Judgment}
-
-For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
-the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows:
-\begin{align*}
-  \ADefine{
-    \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
-  }{
-    \Hom{\mathbb{J}}{
-      \Yoneda{\Upsilon}
-    }{
-      \varprojlim\mathcal{J}
-    }
-  }\\
-  &\cong
-  \OMode{
-    \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
-  }
-\end{align*}
-
- Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
- when the family of judgments $\mathcal{J}$ shall forevermore have a global
- section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
- $\mathcal{J}(\vec{E})$ shall be evident).
-
-\subsection{Internal Judgments}
-It often happens that the evidence for a judgment is the knowledge of the
-meaning of a new form of judgment; this is the case when we say ``To know
-$\mathcal{J}$ is to know \emph{the meaning of} another judgment
-$\mathcal{J}'$.'' This can be neatly expressed in our topos using the subobject
-classifier $\Omega$, where the evidence of $\Omega$ at $\Upsilon$ gives rise to
-an \emph{internal judgment} which is defined at all $\Upsilon'$ for which we
-have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.
-
-A \emph{global internal (family of) judgments} is a section
-$\Of{J}{\int_\SCtx {\Omega^{\mathcal{J}}}}\cong
-\IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}\times\mathcal{J}}{\Omega}}$,
-which can be externalized as a ``total'' judgment $\Of{\Embed{J}}{\mathbb{J}}$
-for whose derivations at $\Upsilon$ are the arrows
-$\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbb{1}\times\mathcal{J}}}$
-such that the following diagram commutes:
-\[
-  \begin{tikzcd}[cramped,sep=huge]
-    \IMode{\Yoneda{\Upsilon}}
-      \arrow[r, dashed, "\OMode{\phi}"]
-      \arrow[rd, "\IMode{J_\Upsilon\Pair{1}{\pi_2\circ\phi}}"']
-    & \IMode{\mathbb{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
-    & \IMode{\Omega}
-  \end{tikzcd}
-\]
-
-Then, to say $\IMode{\Upsilon}\parallel\Embed{J}(E)$ where
-$\Member{E}{\mathcal{J}(\Upsilon)}$ is to assert that we have evidence
-$\Member{F}{\Embed{J}(\Upsilon)}$ such that $\IsEq{\pi_2(F)}{E}$.
-
-\end{appendices}
-
 \ifdraft{}{

### 2015-11-22 paper.tex
Restore rule-style definition of functional refinement
typos

@@
   $\HRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}$, you can conclude
-  $\HRMemEq{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
-
+  $\HEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
+  In other words:
+  %
+  \[
+    \infer[\PFunRefSnoc]{
+      \HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
+    }{
+      \begin{array}{l}
+        \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}.\
+        \forall \Member{M,N}{\Exprs_\sigma(\Upsilon')}.\\
+        \left(\HRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}\right)
+        \Rightarrow
+        \left(\HEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}\right)
+      \end{array}
+    }
+  \]
 \end{quote}

### 2015-11-22 paper.tex
wording

@@
 \end{tikzcd}
-we can define a new refinement $\phi\rho$ such that
-$\Refines{\Upsilon'}{\phi\rho}{\tau}$ in the obvious way.
+we can clearly define a new refinement $\Refines{\Upsilon'}{\phi\rho}{\tau}$.
 
@@
 $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished
-copresheaf $\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
+presheaf $\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
 \emph{canonical} operators such that $\IsSubsetEq{\CanOperators}{\Operators}$.

### 2015-11-24 paper.tex
Factor out common notation into submodule

@@
 \documentclass[11pt]{article}
+
+\usepackage{notation/modes}
+\usepackage{notation/judgments}
+\usepackage{notation/common}
+\usepackage{notation/abts}
 
@@
 
-\newif\ifcolored%
-\coloredtrue%
-
 %% fonts
@@
 
-\ifcolored%
-  \def\IModeColorName{MidnightBlue}
-  \def\OModeColorName{Maroon}
-\else
-  \def\IModeColorName{Black}
-  \def\OModeColorName{Black}
-\fi
-
-\newcommand\IMode[1]{{\color{\IModeColorName}{#1}}}
-\newcommand\OMode[1]{{\color{\OModeColorName}{#1}}}
-
-\newcommand\HypJ[2]{#1\ \ \left(#2\right)}
-\newcommand\GenJ[2]{\vert_{\IMode{#1}}\; #2}
 \newcommand\HPGenJ[4]{\vert^{\IMode{#1}\parallel\IMode{#2}}_{\IMode{#3}}\; #4}
 
-\newcommand\IsTac[1]{\IMode{#1}\ \textit{tactic}}
-
-\newcommand\MkSet[1]{\left\{#1\right\}}
-\newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
-\newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
-\newcommand\Of[2]{\IMode{#1}: \IMode{#2}}
-
-\newcommand\MkValence[3]{%
-  \ifthenelse{\equal{#1}{}}{}{%
-    \{#1\}%
-  }%
-  \ifthenelse{\equal{#2}{}}{}{%
-    [#2]%
-  }.\, #3%
-}
-\newcommand\MkBTm[3]{\mathbb{\lambda}\{#1\}[#2].\, #3}
-\newcommand\MkArity[2]{(#1)\, #2}
-\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
-\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
-\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
-\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}
-
-\newcommand\ArrId[1]{1_{#1}}
-\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
 \newcommand\MapsTo[2]{\IMode{#1}\mapsto\OMode{#2}}
@@
 \newcommand\FinSets{\mathbb{F}}
-\newcommand\Sorts{\mathcal{S}}
-\newcommand\Valences{\mathcal{V}}
-\newcommand\Arities{\mathcal{A}}
-\newcommand\Operators{\mathcal{O}}
 \newcommand\CanOperators{\mathcal{K}}
-
-\newcommand\Cats{\mathbf{Cat}}
-\newcommand\Discrete[1]{#1_{\equiv}}
-\newcommand\Arr[1]{\mathbf{{#1}^{\ra}}}
-\newcommand\Sets{\mathbf{Set}}
-\newcommand\SCtx{\mathbf{SCtx}}
-\newcommand\Ctx{\mathbf{Ctx}}
-\newcommand\Seq{\mathbf{Seq}}
-\newcommand\HypCat{\mathbf{H}}
-
-\newcommand\OpCat[1]{#1^{\mathsf{op}}}
-
-\newcommand\IsOperator[3]{\IMode{#1}\Vdash\IMode{#2}: \OMode{#3}}
-
-\newcommand\IsMetaCtx[1]{\IMode{#1}\ \textit{mctx}}
-\newcommand\IsVarCtx[1]{\IMode{#1}\ \textit{vctx}}
-\newcommand\IsSymCtx[1]{\IMode{#1}\ \textit{sctx}}
-
-\newcommand\Lookup[3]{\IMode{#1}\ni\IMode{#2}:\OMode{#3}}
-
-\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
-\newcommand\Pair[2]{\left\langle#1, #2 \right\rangle}
-\newcommand\Tuple[1]{\left\langle#1 \right\rangle}
-\newcommand\FLift[2]{
-  \Operators\langle#1%
-  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle%
-}
-\newcommand\PairTimes[2]{\langle#1\, {\scriptstyle{{\times}}}\, #2 \rangle}
-\newcommand\SSingleton{\bullet}
-\newcommand\Forget[1]{|\, #1\, |}
-
-\newcommand\Comma[2]{#1 \downarrow#2}
-\newcommand\Coslice[2]{#1\, /\, #2}
 
@@
 }
-\newcommand\IsAbt[5]{\IsAbtUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}{\IMode{#4}}{\OMode{#5}}}
-\newcommand\IsBTm[5]{\IsAbtUnmoded{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}{\IMode{#4}}{\IMode{#5}}}
 
-\newcommand\MV[1]{\mathfrak{#1}}
-\newcommand\MApp[3]{%
-  #1
-  \ifthenelse{\equal{#2}{}}{}{%
-    \{#2\}%
-  }%
-  \ifthenelse{\equal{#3}{}}{}{%
-    [#3]%
-  }
-}
-\newcommand\App[2]{#1(#2)}
-\newcommand\Dom[1]{\left\vert#1\right\vert}
-\newcommand\FV[1]{\mathbf{FV}\left(#1\right)}
-\newcommand\FS[1]{\mathbf{FS}\left(#1\right)}
-\newcommand\NotIn[2]{\IMode{#1}\notin\IMode{#2}}
-\newcommand\SortExp{\mathsf{exp}}
-
-\newcommand\Hom[3]{#1\!\left[#2,\,#3\right]}
-
-\newcommand\Rename[3]{%
+\newcommand\RenameLit[3]{%
   \left\{#1\, /\, #2\right\}#3%
@@
 
-\newcommand\Subst[3]{%
-  \left[#1\, /\, #2\right]#3%
-}
-
-\newcommand\IsSubst[4]{%
-  \Subst{\IMode{#1}}{\IMode{#2}}{\IMode{#3}}%
-  \leadsto\OMode{#4}%
-}
-
-\newcommand\IsEq[2]{\IMode{#1} = \IMode{#2}}
-\newcommand\IsApart[2]{\IMode{#1} \mathrel{\#} \IMode{#2}}
 \newcommand\IsSubsetEq[2]{\IMode{#1}\subseteq\IMode{#2}}
 
-\newcommand\Define[2]{\IMode{#1}\triangleq\OMode{#2}}
-\newcommand\ADefine[2]{\IMode{#1}&\triangleq\OMode{#2}}
-
-\newcommand\SigFun[1]{\mathcal{F}_{#1}}
-
-\newcommand\SemBrackets[1]{\left\llbracket#1\right\rrbracket}
 \newcommand\Embed[1]{\left\lfloor#1\right\rfloor}
-\newcommand\Yoneda[1]{\boldsymbol{y} (#1)}
 
@@
 both by placing restrictions on membership and by coarsening equivalence. Thus
-far, all developments of behavioral type theory have been built on a
+far, most developments of behavioral type theory have been built on a
 \emph{unityped} syntactic framework, and so the relation to type refinements
@@
   $\phi$ (written $\RefEquate{\phi}{\rho}{M}{N}$), such that
-  $\RefEquate{\phi}{\rho}{\cdot}{\cdot}$ is a partial equivalence relation on
+  $\RefEquate{\phi}{\rho}{-}{-}$ is a partial equivalence relation on
   $\Exprs_\tau(\Upsilon')$. Morever, that for any
@@
       \SBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{
-        \Subst{\vec{X}}{\vec{x}}{\Rename{\vec{w}}{\vec{u}}{M}}
+        \Subst{\vec{X}}{\vec{x}}{\RenameLit{\vec{w}}{\vec{u}}{M}}
       }{
-        \Subst{\vec{X}}{\vec{y}}{\Rename{\vec{w}}{\vec{v}}{N}}
+        \Subst{\vec{X}}{\vec{y}}{\RenameLit{\vec{w}}{\vec{v}}{N}}
       }

### 2015-11-24 paper.tex
wordsmithing

@@
 perspective, formal type theories can do little more than negotiate matters of
-grammar, and therefore may serve as a syntactic (linguistic) framework
-for mathematical language, being responsible for the management of variable
-binding and substitution. On the other hand, the meaning of mathematical
-statements shall be specified \emph{behaviorally} in the semantic type
-theory.
+grammar, and therefore may serve as a linguistic framework for mathematical
+expression, being responsible for the management of variable binding and
+substitution. On the other hand, mathematical objects find their \emph{meaning}
+in the behavioral type theory, and are subject to an extensional equality.
 
@@
 has been difficult to see. In this paper, we contribute a full theory of
-behavioral refinements over multi-sorted abstract binding trees, a simple
-formal type theory~\cite{harper:2016, sterling-morrison:2015}; this hybrid
-system allows the deployment of a Nuprl-style type theory over any signature of
-sorts and operators.
+behavioral refinements over multi-sorted abstract binding trees with symbolic
+parameters, a simple but interesting formal type theory~\cite{harper:2016,
+sterling-morrison:2015}; this hybrid system allows the deployment of a
+Nuprl-style type theory over any signature of sorts and operators.
 

### 2015-11-26 paper.tex
rename & document macros

@@
 
+% parametric refinement
 \newcommand\Refines[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset\OMode{#3}}
-\newcommand\GRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}
-\newcommand\RefinesCtx[2]{\IMode{#1}\sqsubset^\star\OMode{#2}}
+
+% global parametric refinement
+\newcommand\GloballyRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}
+
+% equality of parametric refinement
 \newcommand\EqRefines[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\sqsubset\OMode{#4}}
 
-\newcommand\HRefines[4]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}\sqsubset\OMode{#4}}
-\newcommand\HEqRefines[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\sqsubset\OMode{#5}}
-\newcommand\PRefinesCtx[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset^\star\OMode{#3}}
+% parametric context refinement
+\newcommand\RefinesCtx[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset^\star\OMode{#3}}
 
+% functional refinement
+\newcommand\FRefines[4]{%
+  \IMode{#1}\parallel\IMode{#2}%
+  \vDash\IMode{#3}\sqsubset\OMode{#4}%
+}
+
+% equality of functional refinement
+\newcommand\FEqRefines[5]{%
+  \IMode{#1}\parallel\IMode{#2}%
+  \vDash\IMode{#3}=\IMode{#4}
+  \sqsubset\OMode{#5}%
+}
+
+% parametric equality of refinement members
 \newcommand\RMemEq[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\in\IMode{#4}}
-\newcommand\HRMemEq[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\in\IMode{#5}}
 
+% functional equality of refinement members
+\newcommand\FRMemEq[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\in\IMode{#5}}
+
+% parametric subrefinement
+\newcommand\IsSubrefinement[4]{%
+  \IMode{#1}\parallel\IMode{#2}%
+  \subseteq\IMode{#3}\sqsubset\OMode{#4}%
+}
+
+% power object
 \newcommand\Pow[1]{\wp\left(#1\right)}
 
-\newcommand\SubRefines[4]{\IMode{#1}\parallel\IMode{#2}\subseteq\IMode{#3}\sqsubset\OMode{#4}}
 \newcommand\TyBase[1]{\mathtt{Base}_{#1}}
@@
 Our position is that these views of type theory are not in conflict, but rather
-merely describe two distinct layers in a single, harmonious system. From this
+describe two distinct layers in a single, harmonious system. From this
 perspective, formal type theories can do little more than negotiate matters of
@@
 Then, we say that $\phi$ \emph{globally refines} $\tau$ (written
-$\GRefines{\phi}{\tau}$) when we have $\Refines{\cdot}{\phi}{\tau}$.
+$\GloballyRefines{\phi}{\tau}$) when we have $\Refines{\cdot}{\phi}{\tau}$.
 Furthermore, when $\Refines{\Upsilon}{\phi}{\tau}$, for any renaming
@@
 
-We will write $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ to mean that $\phi$ is
+We will write $\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ to mean that $\phi$ is
 a \emph{subrefinement} of $\psi$.
@@
 
-  To know $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ (presupposing
+  To know $\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ (presupposing
   $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$) , is
@@
 $\EqRefines{\Upsilon}{\phi}{\psi}{\tau}$ just when both
-$\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ and
-$\SubRefines{\Upsilon}{\psi}{\phi}{\tau}$.
+$\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ and
+$\IsSubrefinement{\Upsilon}{\psi}{\phi}{\tau}$.
 
@@
 
-\newcommand\PRefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
-\newcommand\PRefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
+\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
+\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
 \newcommand\PFunRefNil{\vDash^\sqsubset_{\mathtt{nil}}}
@@
 Next, we define functional refinement
-$\HEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
+$\FEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
 simultaneously with parametric context refinement
-$\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and functional equality
-$\HRMemEq{\Upsilon}{\Psi}{M}{N}{\phi}$. We will write
-$\HRefines{\Upsilon}{\Psi}{\phi}{\tau}$ as a shorthand for
-$\HEqRefines{\Upsilon}{\Psi}{\phi}{\phi}{\tau}$.
+$\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and functional equality
+$\FRMemEq{\Upsilon}{\Psi}{M}{N}{\phi}$. We will write
+$\FRefines{\Upsilon}{\Psi}{\phi}{\tau}$ as a shorthand for
+$\FEqRefines{\Upsilon}{\Psi}{\phi}{\phi}{\tau}$.
 
-\newcommand\DefPRefinesCtxNil[1]{
-  \infer[\PRefinesCtxNil]{
-    \PRefinesCtx{#1}{\cdot}{\cdot}
+\newcommand\DefRefinesCtxNil[1]{
+  \infer[\RefinesCtxNil]{
+    \RefinesCtx{#1}{\cdot}{\cdot}
   }{
@@
 }
-\newcommand\DefPRefinesCtxSnoc[6]{
-  \infer[\PRefinesCtxSnoc]{
-    \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
+\newcommand\DefRefinesCtxSnoc[6]{
+  \infer[\RefinesCtxSnoc]{
+    \RefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
   }{
-    \PRefinesCtx{#1}{#2}{#5} &
-    \HRefines{#1}{#2}{#4}{#6}
+    \RefinesCtx{#1}{#2}{#5} &
+    \FRefines{#1}{#2}{#4}{#6}
   }
@@
 
-Parametric context refinement $\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ shall be
+Parametric context refinement $\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ shall be
 defined as follows, presupposing $\IsSymCtx{\Upsilon}$ and
@@
 \begin{gather*}
-  \DefPRefinesCtxNil{\Upsilon}\qquad
-  \DefPRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\phi}{\Gamma}{\tau}
+  \DefRefinesCtxNil{\Upsilon}\qquad
+  \DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\phi}{\Gamma}{\tau}
 \end{gather*}
 
-For a context refinement $\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and a renaming
+For a context refinement $\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and a renaming
 \begin{tikzcd}[cramped]
@@
 we can define a new context refinement
-$\PRefinesCtx{\Upsilon'}{\Psi\rho}{\Gamma}$ by applying $\rho$ pointwise at
+$\RefinesCtx{\Upsilon'}{\Psi\rho}{\Gamma}$ by applying $\rho$ pointwise at
 each of the sort refinements in the context.
 
-Functional refinement $\HEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ presupposes
-$\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
+Functional refinement $\FEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ presupposes
+$\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
 induction on the evidence for its first presupposition.
 
-\paragraph{Case} $\DefPRefinesCtxNil{\Upsilon}$.
+\paragraph{Case} $\DefRefinesCtxNil{\Upsilon}$.
 \[
   \infer[\PFunRefNil]{
-    \HEqRefines{\Upsilon}{\cdot}{\phi}{\psi}{\tau}
+    \FEqRefines{\Upsilon}{\cdot}{\phi}{\psi}{\tau}
   }{
@@
 
-\paragraph{Case} $\DefPRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
+\paragraph{Case} $\DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
 
 \begin{quote}
-  To know $\HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}$ is to know, for any renaming
+  To know $\FEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}$ is to know, for any renaming
   \begin{tikzcd}[cramped]
@@
   and closed terms $\Member{M,N}{\Exprs_\sigma(\Upsilon')}$, that from
-  $\HRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}$, you can conclude
-  $\HEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
+  $\FRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}$, you can conclude
+  $\FEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
   In other words:
@@
     \infer[\PFunRefSnoc]{
-      \HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
+      \FEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
     }{
@@
         \forall \Member{M,N}{\Exprs_\sigma(\Upsilon')}.\\
-        \left(\HRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}\right)
+        \left(\FRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}\right)
         \Rightarrow
-        \left(\HEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}\right)
+        \left(\FEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}\right)
       \end{array}
@@
 %%
-%% Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
-%% $\RefinesCtx{\Phi}{\Upsilon}$, $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$,
-%% $\HRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
+%% Functional equality $\FRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
+%% $\RefinesCtx{\Phi}{\Upsilon}$, $\RefinesCtx{\Phi}{\Psi}{\Gamma}$,
+%% $\FRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
 %% and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
 %% induction on the evidence for the second presupposition
-%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
+%% $\RefinesCtx{\Phi}{\Psi}{\Gamma}$:
 %%
-%% \paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
+%% \paragraph{Case} $\DefRefinesCtxNil{\Phi}$.
 %% \[
 %%   \infer{
-%%     \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
+%%     \FRMemEq{\Phi}{\cdot}{M}{N}{\phi}
 %%   }{
@@
 %%
-%% \paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
+%% \paragraph{Case} $\DefRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
 %% where $\RefinesCtx{\Phi}{\Upsilon}$.
@@
 %%   \infer{
-%%     \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
+%%     \FRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
 %%   }{
@@
 %%       \HypJ{
-%%         \HRMemEq{\Phi}{\Psi}{
+%%         \FRMemEq{\Phi}{\Psi}{
 %%           \Subst{\MV{m}}{x}{\phi}
@@
 %%       }{
-%%         \HRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
+%%         \FRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
 %%       }
@@
 %% Refinements can be arranged in a partial order via the
-%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
+%% $\IsSubrefinement{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
 %% at any sort $\tau$.
 %%
-%% We have the bottom refinement $\GRefines{\TyBot{\tau}}{\tau}$, which simply
-%% contains no elements. Next, the top refinement $\GRefines{\TyTop{\tau}}{\tau}$
+%% We have the bottom refinement $\GloballyRefines{\TyBot{\tau}}{\tau}$, which simply
+%% contains no elements. Next, the top refinement $\GloballyRefines{\TyTop{\tau}}{\tau}$
 %% identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$

### 2015-11-26 paper.tex
Move some macros to a new package
First cut definition of functional equality; fix bugs

@@
 \usepackage{notation/abts}
+\usepackage{refinements}
 
@@
 
+% power object
+\newcommand\Pow[1]{\wp\left(#1\right)}
+
 \newcommand\MapsTo[2]{\IMode{#1}\mapsto\OMode{#2}}
+\newcommand\IsSubsetEq[2]{\IMode{#1}\subseteq\IMode{#2}}
 
@@
 }
-
-\newcommand\IsSubsetEq[2]{\IMode{#1}\subseteq\IMode{#2}}
-
-\newcommand\Embed[1]{\left\lfloor#1\right\rfloor}
-
-% parametric refinement
-\newcommand\Refines[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset\OMode{#3}}
-
-% global parametric refinement
-\newcommand\GloballyRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}
-
-% equality of parametric refinement
-\newcommand\EqRefines[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\sqsubset\OMode{#4}}
-
-% parametric context refinement
-\newcommand\RefinesCtx[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset^\star\OMode{#3}}
-
-% functional refinement
-\newcommand\FRefines[4]{%
-  \IMode{#1}\parallel\IMode{#2}%
-  \vDash\IMode{#3}\sqsubset\OMode{#4}%
-}
-
-% equality of functional refinement
-\newcommand\FEqRefines[5]{%
-  \IMode{#1}\parallel\IMode{#2}%
-  \vDash\IMode{#3}=\IMode{#4}
-  \sqsubset\OMode{#5}%
-}
-
-% parametric equality of refinement members
-\newcommand\RMemEq[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\in\IMode{#4}}
-
-% functional equality of refinement members
-\newcommand\FRMemEq[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\in\IMode{#5}}
-
-% parametric subrefinement
-\newcommand\IsSubrefinement[4]{%
-  \IMode{#1}\parallel\IMode{#2}%
-  \subseteq\IMode{#3}\sqsubset\OMode{#4}%
-}
-
-% power object
-\newcommand\Pow[1]{\wp\left(#1\right)}
 
@@
   \end{tikzcd}
-  and closed terms $\Member{M,N}{\Exprs_\sigma(\Upsilon')}$, that from
-  $\FRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}$, you can conclude
-  $\FEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
+  and closed terms $\Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}$, that from
+  $\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}$, you can conclude
+  $\FEqRefines{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{\phi\rho}}{\Subst{M_1}{x}{\psi\rho}}{\tau}$.
   In other words:
@@
         \end{tikzcd}.\
-        \forall \Member{M,N}{\Exprs_\sigma(\Upsilon')}.\\
-        \left(\FRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}\right)
+        \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
+        \left(\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}\right)
         \Rightarrow
-        \left(\FEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}\right)
+        \left(\FEqRefines{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{\phi\rho}}{\Subst{M_1}{x}{\psi\rho}}{\tau}\right)
       \end{array}
@@
 
-%%
-%% \subsection{Functional Equality}
-%%
-%% Functional equality $\FRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
-%% $\RefinesCtx{\Phi}{\Upsilon}$, $\RefinesCtx{\Phi}{\Psi}{\Gamma}$,
-%% $\FRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
-%% and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
-%% induction on the evidence for the second presupposition
-%% $\RefinesCtx{\Phi}{\Psi}{\Gamma}$:
-%%
-%% \paragraph{Case} $\DefRefinesCtxNil{\Phi}$.
-%% \[
-%%   \infer{
-%%     \FRMemEq{\Phi}{\cdot}{M}{N}{\phi}
-%%   }{
-%%     \RMemEq{\Phi}{M}{N}{\phi}
-%%   }
-%% \]
-%%
-%% \paragraph{Case} $\DefRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
-%% where $\RefinesCtx{\Phi}{\Upsilon}$.
-%% \[
-%%   \infer{
-%%     \FRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
-%%   }{
-%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
-%%       \HypJ{
-%%         \FRMemEq{\Phi}{\Psi}{
-%%           \Subst{\MV{m}}{x}{\phi}
-%%         }{
-%%           \Subst{\MV{n}}{x}{\psi}
-%%         }{\phi}
-%%       }{
-%%         \FRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
-%%       }
-%%     }
-%%   }
-%% \]
+\subsection{Functional Equality}
+
+Functional equality $\FRMemEq{\Upsilon}{\Psi}{M}{N}{\phi}$ presupposes
+$\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$, $\FRefines{\Upsilon}{\Psi}{\phi}{\tau}$,
+and $\Member{M,N}{\Exprs_\tau(\Upsilon)}$; this judgment is defined by
+induction on the evidence for the first presupposition
+$\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$:
+
+\paragraph{Case} $\DefRefinesCtxNil{\Upsilon}$.
+\[
+  \infer{
+    \FRMemEq{\Upsilon}{\cdot}{N_0}{N_1}{\phi}
+  }{
+    \RMemEq{\Upsilon}{N_0}{N_1}{\phi}
+  }
+\]
+
+\paragraph{Case} $\DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
+\begin{quote}
+  To know $\FRMemEq{\Upsilon}{\Psi,x:\chi}{N_0}{N_1}{\phi}$ is to know, for any renaming
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}
+  and closed terms $\Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}$, that from
+  $\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}$, you can conclude
+  $\FRMemEq{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{N_0\rho}}{\Subst{M_1}{x}{N_1\rho}}{\Subst{M_0}{x}{\phi\rho}}$.
+  In other words:
+  %
+  \[
+    \infer{
+      \FRMemEq{\Upsilon}{\Psi,x:\psi}{N_0}{N_1}{\phi}
+    }{
+      \begin{array}{l}
+        \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}.\
+        \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
+        \left(\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}\right)
+        \Rightarrow
+        \left(%
+          \FRMemEq{\Upsilon'}{\Psi\rho}{%
+            \Subst{M_0}{x}{N_0\rho}
+          }{%
+            \Subst{M_1}{x}{N_1\rho}%
+          }{\Subst{M_0}{x}{\phi\rho}
+          }%
+        \right)
+      \end{array}
+    }
+  \]
+\end{quote}
 %%

### 2015-11-26 paper.tex
Correct presupposition of functional equality

@@
 
-Functional equality $\FRMemEq{\Upsilon}{\Psi}{M}{N}{\phi}$ presupposes
+Functional equality $\FRMemEq{\Upsilon}{\Psi}{N_0}{N_1}{\phi}$ presupposes
 $\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$, $\FRefines{\Upsilon}{\Psi}{\phi}{\tau}$,
-and $\Member{M,N}{\Exprs_\tau(\Upsilon)}$; this judgment is defined by
+and $\Member{N_0,N_1}{\Exprs_\tau(\Upsilon\parallel\Gamma)}$; this judgment is defined by
 induction on the evidence for the first presupposition

### 2015-11-26 paper.tex
remove reundant definition

@@
 $\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-covariant presheaves on $\SCtx\times\Ctx$ of open expressions, open values, and open bound terms
+covariant presheaves on $\SCtx\times\Ctx$ of open values, and open bound terms
 as follows:
 \begin{align*}
-  \ADefine{\Exprs_\tau(\Upsilon\parallel\Gamma)}{
-    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\Gamma}{M}{\tau}}
-  }\\
   \ADefine{\Values_\tau(\Upsilon\parallel\Gamma)}{
@@
 
-We'll write $\Exprs_\tau(\Upsilon)$, $\Values_\tau(\Upsilon)$ and
-$\BTms_v(\Upsilon)$ for $\Exprs_\tau(\Upsilon\parallel\cdot)$,
+We'll write $\Values_\tau(\Upsilon)$ and $\BTms_v(\Upsilon)$ for
 $\Values_\tau(\Upsilon\parallel\cdot)$ and $\BTms_v(\Upsilon\parallel\cdot)$

### 2015-11-26 paper.tex
Update notation
wording

@@
 
-\newcommand\Symbols{\mathbf{I}}
-\newcommand\SymSets{\mathbb{F}_\Symbols}
-\newcommand\FinSets{\mathbb{F}}
 \newcommand\CanOperators{\mathcal{K}}
@@
 models for behavioral type theory interpret the types as partial equivalence
-relations on only closed terms, and the meaning of the hypothetical judgment is
-defined separately and uniformly in the logical relations style.
+relations (PER) on only closed terms, and the meaning of the hypothetical
+judgment is defined separately and uniformly in the logical relations style.
 
@@
     \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
-  \end{tikzcd},
-  what it means for any $\Member{M,N}{\Exprs_\tau(\Upsilon')}$ to be equated by
+  \end{tikzcd}
+, what it means for any $\Member{M,N}{\Exprs_\tau(\Upsilon')}$ to be equated by
   $\phi$ (written $\RefEquate{\phi}{\rho}{M}{N}$), such that
-  $\RefEquate{\phi}{\rho}{-}{-}$ is a partial equivalence relation on
+  $\RefEquate{\phi}{\rho}{-}{-}$ is a partial equivalence relation (PER) on
   $\Exprs_\tau(\Upsilon')$. Morever, that for any
@@
     \IMode{\Upsilon'}\arrow[r,hook,"\IMode{\rho'}"] &\IMode{\Upsilon''}
-  \end{tikzcd},
+  \end{tikzcd}
+,
   from $\RefEquate{\phi}{\rho}{M}{N}$ we may conclude
@@
   At this point, it should be noted that this is very similar to the semantical
-  explanation of typehood given in \cite{martin-lof:1979}, except that we have
+  explanation of typehood given in~\cite{martin-lof:1979}, except that we have
   generalized it to a multi-sorted setting, and that we have fibred the entire
@@
 
+  It should be noted that PERs can be understood as 0-semigroupoids. In other
+  words, these are 1-groupoids without identity morphisms and without coherence
+  laws regarding composition and inverse operations. This distinction is
+  especially relevant in formal type theory where groupoids ``one dimension
+  down'' are not sets but setoids (i.e., groupoids sans coherence). Thus, a PER
+  can formally be thought of as a kind of partial setoid that facilitates
+  constructions on subsets in addition to quotients. \todo[inline]{Say something
+    about internal (semi)-groupoids?}
+
 \end{remark}
@@
   To know $\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ (presupposing
-  $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$) , is
+  $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$), is
   to know, for any renamings
@@
     & \IMode{\Upsilon''}
-  \end{tikzcd},
-  that from $\RefEquate{\phi}{\rho}{M}{N}$ you can conclude
+  \end{tikzcd}
+, that from $\RefEquate{\phi}{\rho}{M}{N}$ you can conclude
   $\RefEquate{\psi}{\rho'\circ\rho}{M\rho'}{N\rho'}$ for any
@@
 
-Two refinements are equal when they denote the same PER. That is, we have
+Two refinements are equal when they denote the same PER.\@That is, we have
 $\EqRefines{\Upsilon}{\phi}{\psi}{\tau}$ just when both
@@
   \IMode{\Upsilon'} \arrow[r, hook, "\IMode{\rho}"] &\IMode{\Upsilon}
-\end{tikzcd},
-and $\Member{M,N}{\Exprs_\tau(\Upsilon)}$. The meaning of this judgment is
+\end{tikzcd}
+, and $\Member{M,N}{\Exprs_\tau(\Upsilon)}$. The meaning of this judgment is
 that $M$ and $N$ are identified by $\phi$ at $\Upsilon$:
@@
   \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
-\end{tikzcd},
-we can define a new context refinement
+\end{tikzcd}
+, we can define a new context refinement
 $\RefinesCtx{\Upsilon'}{\Psi\rho}{\Gamma}$ by applying $\rho$ pointwise at
@@
           \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
-        \end{tikzcd}.\
-        \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
+        \end{tikzcd}
+        .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
         \left(\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}\right)
@@
   relations on $X$ as the exponential $\Define{\Pow{X}}{\Omega^X}$. Then, the
-  data of such a relation is contained in a natural transformation
+  data of such a relation is contained in a global section
   $\Of{R}{\Hom{\mathcal{E}}{\mathbb{1}}{\Pow{X}}}$.
@@
   \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
-\end{tikzcd},
-we have $\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
+\end{tikzcd}
+, we have $\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
 $\SBinRel{R_{\vec{v}}}{\Upsilon'}{\vec{E}\rho}{\vec{F}\rho}$.

### 2015-11-27 paper.tex
Introduce PER abbreviation

--- 2015-11-26 paper.tex
@@
   $\phi$ (written $\RefEquate{\phi}{\rho}{M}{N}$), such that
-  $\RefEquate{\phi}{\rho}{-}{-}$ is a partial equivalence relation on
+  $\RefEquate{\phi}{\rho}{-}{-}$ is a partial equivalence relation (PER) on
   $\Exprs_\tau(\Upsilon')$. Morever, that for any

### 2015-11-27 paper.tex
Say something about 0-semigroupoids

@@
   \end{align*}
+
+  It should be noted that PERs can be understood as 0-semigroupoids. In other
+  words, these are 1-groupoids without identity morphisms and without coherence
+  laws regarding composition and inverse operations. This distinction is
+  especially relevant in formal type theory where groupoids ``one dimension
+  down'' are not sets but setoids (i.e., groupoids sans coherence). Thus, a PER
+  can formally be thought of as a kind of partial setoid that facilitates
+  constructions on subsets in addition to quotients. \todo[inline]{Say something
+    about internal (semi)groupoids?}
 

### 2015-11-27 paper.tex
Fix chktex warnings
Restore "Constructions on Refinements"

--- 2015-11-26 paper.tex
@@
 \end{quote}
-%%
-%% \section{Constructions on Refinements}
-%%
-%% Refinements can be arranged in a partial order via the
-%% $\IsSubrefinement{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
-%% at any sort $\tau$.
-%%
-%% We have the bottom refinement $\GloballyRefines{\TyBot{\tau}}{\tau}$, which simply
-%% contains no elements. Next, the top refinement $\GloballyRefines{\TyTop{\tau}}{\tau}$
-%% identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
-%% and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
-%% $\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
-%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
-%% \begin{gather*}
-%%   \infer{
-%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
-%%   }{
-%%     \RMemEq{\Phi}{M}{N}{\phi}
-%%   }\qquad
-%%   \infer{
-%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
-%%   }{
-%%     \RMemEq{\Phi}{M}{N}{\psi}
-%%   }
-%% \end{gather*}
-%%
-%% Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
-%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
-%% dually:
-%% \[
-%%   \infer{
-%%     \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
-%%   }{
-%%     \RMemEq{\Phi}{M}{N}{\phi} &
-%%     \RMemEq{\Phi}{M}{N}{\psi}
-%%   }
-%% \]
-%%
-%% It is trivial to verify that these constructions on refinements give rise to a
-%% lattice.
-%%
-%% \section{Multi-Sorted Computational Type Theory}
-%%
-%% Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
-%% you specify how to form its equal canonical inhabitants, or values. Then, the
-%% inhabitants of the types are explained uniformly by extending this relation
-%% over an operational semantics such that two terms are equal members of a
-%% type just when they compute to equal values of that type.
-%%
-%% More generally, one may wish to consider types which include computations in
-%% addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
-%% is a type, which would include all the members of $\psi$ as well as a divergent
-%% computation. Another example is the type $\TyTop{\tau}$ which identifies all
-%% closed terms of sort $\tau$.
-%%
-%% In order to ensure that such types are sensible, we must formulate a notion of
-%% computational equivalence, and then say that a refinement is a \emph{type} just
-%% when it respects computational equivalence; then, we are able to define types
-%% as PERs on \emph{programs} in addition to values.
+
+\section{Constructions on Refinements}
+
+Refinements can be arranged in a partial order via the
+$\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
+at any sort $\tau$.
+
+We have the bottom refinement $\GloballyRefines{\TyBot{\tau}}{\tau}$, which
+simply contains no elements. Next, the top refinement
+$\GloballyRefines{\TyTop{\tau}}{\tau}$ identifies all terms
+$\Member{M,N}{\Exprs_\tau(\Upsilon)}$.  The join
+$\Refines{\Upsilon}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
+$\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$ is defined as
+follows:
+
+\begin{gather*}
+  \infer{
+    \RMemEq{\Upsilon}{M}{N}{\phi\TyJoin{\tau}\psi}
+  }{
+    \RMemEq{\Upsilon}{M}{N}{\phi}
+  }\qquad
+  \infer{
+    \RMemEq{\Upsilon}{M}{N}{\phi\TyJoin{\tau}\psi}
+  }{
+    \RMemEq{\Upsilon}{M}{N}{\psi}
+  }
+\end{gather*}
+
+Finally the meet $\Refines{\Upsilon}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
+$\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$ is defined
+dually:
+\[
+  \infer{
+    \RMemEq{\Upsilon}{M}{N}{\phi\TyMeet{\tau}\psi}
+  }{
+    \RMemEq{\Upsilon}{M}{N}{\phi} &
+    \RMemEq{\Upsilon}{M}{N}{\psi}
+  }
+\]
+
+It is trivial to verify that these constructions on refinements give rise to a
+lattice.
+
+\section{Multi-Sorted Computational Type Theory}
+
+Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
+you specify how to form its equal canonical inhabitants, or values. Then, the
+inhabitants of the types are explained uniformly by extending this relation
+over an operational semantics such that two terms are equal members of a
+type just when they compute to equal values of that type.
+
+More generally, one may wish to consider types which include computations in
+addition to values; an example of such a type would be $\psi_\bot$ when $\psi$
+is a type, which would include all the members of $\psi$ as well as a divergent
+computation. Another example is the type $\TyTop{\tau}$ which identifies all
+closed terms of sort $\tau$.
+
+In order to ensure that such types are sensible, we must formulate a notion of
+computational equivalence, and then say that a refinement is a \emph{type} just
+when it respects computational equivalence; then, we are able to define types
+as PERs on \emph{programs} in addition to values.
 

### 2015-11-27 paper.tex
remove dead comment
Pluralization

@@
 models for behavioral type theory interpret the types as partial equivalence
-relations (PER) on only closed terms, and the meaning of the hypothetical
+relations (PERs) on only closed terms, and the meaning of the hypothetical
 judgment is defined separately and uniformly in the logical relations style.
@@
 
-%%% Now, let us consider relations on open terms; in general, it will be easiest
-%%% for us to situate our constructions on relations in the presheaf topos
-%%% $\Define{\widehat{\HypCat}}{\Sets^{\SCtx\times\Ctx}}$. Then, for any two
-%%% objects objects $\Of{X,Y}{\widehat{\HypCat}}$, we can form the object of
-%%% relations between them as $\Define{\Pow{X\times Y}}{\Omega^{X\times Y}}$.
-%%%
-%% . A binary relation $R_\tau$
-%% between open expressions of sort $\tau$ can be thought of as a global section
-%% of the exponential object
-%% $\Match{\Pow{\Exprs_\tau\times\Exprs_\tau}}{\Omega^{\Exprs_\tau\times\Exprs_\tau}}$
-%% of the presheaf topos $\Define{\widehat{\HypCat}}{\Sets^{\SCtx\times\Ctx}}$
-
 \ifdraft{}{

### 2015-12-19 paper.tex
Add todo note
Fix fishy/incorrect rule
Explain some notation
correct renaming notation

@@
 \usepackage{tikz-cd}
-\usepackage{tikzpfeile}
+\usepackage[length=.5]{tikzpfeile}
 \usepackage{url}
@@
 \newtheorem{example}[thm]{Example}
+\theoremstyle{notation}
+\newtheorem{notation}[thm]{Notation}
+\newtheorem*{notation*}{Notation}
 \theoremstyle{remark}
@@
   #4:\OMode{#5}%
-}
-
-\newcommand\RenameLit[3]{%
-  \left\{#1\, /\, #2\right\}#3%
 }
@@
   from $\RefEquate{\phi}{\rho}{M}{N}$ we may conclude
-  $\RefEquate{\phi}{\rho'\circ\rho}{M\rho'}{N\rho'}$.
+  $\RefEquate{\phi}{\rho'\circ\rho}{\Rename{\rho'}{M}}{\Rename{\rho'}{N}}$.
 \end{definition}
@@
 \end{tikzcd}
-we can clearly define a new refinement $\Refines{\Upsilon'}{\phi\rho}{\tau}$.
+we can clearly define a new refinement $\Refines{\Upsilon'}{\Rename{\rho}{\rho}}{\tau}$.
 
@@
   , that from $\RefEquate{\phi}{\rho}{M}{N}$ you can conclude
-  $\RefEquate{\psi}{\rho'\circ\rho}{M\rho'}{N\rho'}$ for any
+  $\RefEquate{\psi}{\rho'\circ\rho}{\Rename{\rho'}{M}}{\Rename{\rho'}{N}}$ for any
   $\Member{M,N}{\Exprs_\tau(\Upsilon')}$.
@@
 , we can define a new context refinement
-$\RefinesCtx{\Upsilon'}{\Psi\rho}{\Gamma}$ by applying $\rho$ pointwise at
+$\RefinesCtx{\Upsilon'}{\Rename{\rho}{\Psi}}{\Gamma}$ by applying $\rho$ pointwise at
 each of the sort refinements in the context.
@@
   and closed terms $\Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}$, that from
-  $\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}$, you can conclude
-  $\FEqRefines{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{\phi\rho}}{\Subst{M_1}{x}{\psi\rho}}{\tau}$.
+  $\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}$, you can conclude
+  $\FEqRefines{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{\phi}}}{\Subst{M_1}{x}{\Rename{\rho}{\psi}}}{\tau}$.
   In other words:
@@
         .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
-        \left(\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}\right)
+        \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)
         \Rightarrow
-        \left(\FEqRefines{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{\phi\rho}}{\Subst{M_1}{x}{\psi\rho}}{\tau}\right)
+        \left(\FEqRefines{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{\phi}}}{\Subst{M_1}{x}{\Rename{\rho}{\psi}}}{\tau}\right)
       \end{array}
@@
   and closed terms $\Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}$, that from
-  $\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}$, you can conclude
-  $\FRMemEq{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{N_0\rho}}{\Subst{M_1}{x}{N_1\rho}}{\Subst{M_0}{x}{\phi\rho}}$.
+  $\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}$, you can conclude
+  $\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{N_0}}}{\Subst{M_1}{x}{\Rename{\rho}{N_1}}}{\Subst{M_0}{x}{\phi\rho}}$.
   In other words:
@@
         .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
-        \left(\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}\right)
+        \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)
         \Rightarrow
         \left(%
-          \FRMemEq{\Upsilon'}{\Psi\rho}{%
-            \Subst{M_0}{x}{N_0\rho}
+          \FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{%
+            \Subst{M_0}{x}{\Rename{\rho}{N_0}}
           }{%
-            \Subst{M_1}{x}{N_1\rho}%
+            \Subst{M_1}{x}{\Rename{\rho}{N_1}}%
           }{\Subst{M_0}{x}{\phi\rho}
@@
       \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon}\cup\vec{u}\cup\vec{v}}.\quad
-      \forall\Of{\vec{X}}{\Exprs_\tau^{[\vec{\tau}]}(\Upsilon,\vec{w}:\vec{\sigma})}.\\
-      \SBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{
-        \Subst{\vec{X}}{\vec{x}}{\RenameLit{\vec{w}}{\vec{u}}{M}}
+      \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+           \IMode{\Upsilon,\vec{w}:\vec{\sigma}} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}.\quad
+      \forall\Of{\vec{X}}{\Exprs^{[\vec{\tau}]}(\Upsilon')}.\\
+      \quad
+      \SBinRel{R_\tau}{\Upsilon'}{
+        \Subst{\vec{X}}{\vec{x}}{
+          \Rename{(\rho\circ \vec{u}\mapsto\vec{w})}{M}
+        }
       }{
-        \Subst{\vec{X}}{\vec{y}}{\RenameLit{\vec{w}}{\vec{v}}{N}}
+        \Subst{\vec{X}}{\vec{y}}{
+          \Rename{(\rho\circ \vec{u}\mapsto\vec{w})}{M}
+        }
       }
@@
 
-As a matter of convenience, we'll also define this judgment over vectors of bound terms and valences:
+\begin{notation*}
+  For a family of copresheaves $\HypJ{\Of{X_\tau}{\Sets^\SCtx}}{\Member{\tau}{\Sorts}}$, we will
+  use $X^{[\vec{\tau}]}$ as a shorthand for the following iterated product, as
+  in \cite{sterling-morrison:2015}:
+
+  \[
+    \Define{
+      X^{[\vec{\tau}]}
+    }{
+      \prod_{\tau\in\vec{\tau}}
+        X_\tau
+    }
+  \]
+\end{notation*}
+
+As a matter of convenience, we'll also define this judgment over vectors of
+bound terms and valences:
 \[
@@
 \end{tikzcd}
-, we have $\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
-$\SBinRel{R_{\vec{v}}}{\Upsilon'}{\vec{E}\rho}{\vec{F}\rho}$.
+, we have $\Eval{\Upsilon'}{\tau}{\Rename{\rho}{N}}{\App{\Rename{\rho}{\vartheta}}{\Rename{\rho}{\Vec{F}}}}$ and
+$\SBinRel{R_{\vec{v}}}{\Upsilon'}{\Rename{\rho}{\vec{E}}}{\Rename{\rho}{\vec{F}}}$.
 

### 2015-12-19 paper.tex
for lcs, no need to talk about syntax of open terms

@@
 $\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-covariant presheaves on $\SCtx\times\Ctx$ of open values, and open bound terms
+covariant presheaves on $\SCtx$ of closed values, and closed bound terms
 as follows:
 \begin{align*}
-  \ADefine{\Values_\tau(\Upsilon\parallel\Gamma)}{
+  \ADefine{\Values_\tau(\Upsilon)}{
     \MkSet{
       M\equiv\App{\vartheta}{\vec{E}}
-      \mid M\in\Exprs_\tau(\Upsilon\parallel\Gamma)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
+      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
     }
@@
   \ADefine{
-    \BTms_{v}(\Upsilon\parallel\Gamma)
+    \BTms_{v}(\Upsilon)
   }{
-    \MkSet{E \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\Gamma}{E}{v}}
+    \MkSet{E \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{E}{v}}
   }
@@
 
-We'll write $\Values_\tau(\Upsilon)$ and $\BTms_v(\Upsilon)$ for
-$\Values_\tau(\Upsilon\parallel\cdot)$ and $\BTms_v(\Upsilon\parallel\cdot)$
-respectively, viewed as covariant presheaves on $\SCtx$. Then, a \emph{lazy
-computation system} (lcs) is a lazy computation language
+Then, a \emph{lazy computation system} (lcs) is a lazy computation language
 $\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an $\SCtx$-indexed

### 2015-12-19 paper.tex
clearer wording

@@
 \emph{canonical} operators such that $\IsSubsetEq{\CanOperators}{\Operators}$.
+
 For a lazy computation language
-$\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
-covariant presheaves on $\SCtx$ of closed values, and closed bound terms
-as follows:
+$\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define
+sort-indexed (resp.\ valence-indexed) copresheaves on $\SCtx$ of closed values
+(resp\. bound terms) as follows:
 \begin{align*}

### 2015-12-19 paper.tex
inevitable ⇒ possible

@@
 sheaves over contexts of hypotheses, and as such, a proof theoretic
-interpretation of the hypothetical judgment is inevitable.
+interpretation of the hypothetical judgment is possible.
 

### 2015-12-19 paper.tex
Add motivation / remark about symbols; discuss Allen's atoms

@@
 expression, being responsible for the management of variable binding and
-substitution. On the other hand, mathematical objects find their \emph{meaning}
-in the behavioral type theory, and are subject to an extensional equality.
+substitution, among other things. On the other hand, mathematical objects find
+\emph{meaning} in their extension within the behavioral type theory, and are
+subject to an extensional equality.
 
@@
 \emph{unityped} syntactic framework, and so the relation to type refinements
-has been difficult to see. In this paper, we contribute a full theory of
-behavioral refinements over multi-sorted abstract binding trees with symbolic
-parameters, a simple but interesting formal type theory~\cite{harper:2016,
-sterling-morrison:2015}; this hybrid system allows the deployment of a
-Nuprl-style type theory over any signature of sorts and operators.
+has been difficult to see.
+
+In this paper, we contribute a full theory of behavioral refinements over
+multi-sorted nominal abstract binding trees with symbolic parameters, a simple
+but interesting formal type theory~\cite{harper:2016, sterling-morrison:2015};
+this hybrid system allows the deployment of a Nuprl-style type theory over any
+signature of sorts and operators.
+
+Owing to the \emph{nominal} aspect of our abstract binding tree (abt) framework, this
+development constitutes a possible alternative to Allen's semantics for
+unguessable atoms in Nuprl~\cite{allen:2006}; Allen proposed a
+``supervaluation'' semantics, that explained sequents involving atoms (symbols)
+by quantifying over the possible ways to interpret the class of atoms. Our
+semantics seem to be a more direct and concrete way to approach the problem,
+which is licensed in part by the nominal character of the abt framework.
 

### 2015-12-19 paper.tex
Give meaning explanation for typehood in CTT

@@
 \subsection{Parametric Refinement}
+\label{sec:parametric-refinement}
 
@@
 
-\begin{remark}
+\begin{remark}\label{rem:internal}
 
@@
   generalized it to a multi-sorted setting, and that we have fibred the entire
-  apparatus over collections $\Upsilon$ of symbols.
+  apparatus over collections $\Upsilon$ of symbols; furthermore, whereas
+  Martin-L\"of defines types in terms of evaluation to canonical form, we have
+  remained agnostic on this point as far as refinements are concerned. We will
+  come back to the notions of \emph{type} and \emph{computation} in
+  section~\ref{sec:ctt}.
 
@@
   the presheaf topos $\Sets^\SCtx$, a refinement is merely a section of the
-  object of $\Exprs_\tau$-PERs $\mathfrak{per}(\Exprs_\tau)$, defined
+  object of $\Exprs_\tau$-PERs $\PERs{\Exprs_\tau}$, defined
   internally as a subobject of $\Omega^{\Exprs_\tau\times\Exprs_\tau}$:
@@
     \Define{
-      \mathfrak{per}(X)
+      \PERs{X}
     }{
@@
   So, $\Refines{\Upsilon}{\phi}{\tau}$ obtains just when we have
-  $\Member{\phi}{\mathfrak{per}(\Exprs_\tau)(\Upsilon)}$. The benefit of
+  $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of
   explaining such judgments in terms of the presheaf topos is that we do not
@@
 \section{Multi-Sorted Computational Type Theory}
+\label{sec:ctt}
 
@@
 
+\subsection{When is a refinement a type?}
+
+As we remarked in section~\ref{sec:parametric-refinement}, the notion of
+\emph{refinement} that we described is very much akin to a multi-sorted
+generalization of Martin-L\"of's explanation of types
+in~\cite{martin-lof:1979}, except that we have not required an operational
+semantics to be present prior to the definition of refinements.
+
+Types, on the other hand, are intrinsically tied up with computation.  In order
+to make this concrete, let us first define what it takes for a refinement to be
+a type, in the sense of Martin-L\"of---that is, a \emph{value type}. In the
+simple case, where we have a global refinement $\GloballyRefines{\phi}{\tau}$,
+we will say $\GlobalIsValueType{\phi}{\tau}$ when there exists a PER $\psi$ on
+\emph{values} that factors uniquely through $\phi$.
+
+In general, we must explain typehood for parametric refinements
+$\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
+Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
+transformation
+$\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
+Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
+is a unique $\psi$ such that the following diagram commutes, where
+$\lambda(\chi)$ is the exponential transpose of $\chi$:
+\[
+  \IsValueType{\Upsilon}{\phi}{\tau}
+  \triangleq
+  \begin{tikzcd}[sep=huge]
+    \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
+      \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
+      \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
+    \IMode{\Omega}\\
+    \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
+      \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
+  \end{tikzcd}
+\]
+
+Contrary to Martin-L\"of, however, we wish to consider a full range of
+\emph{computational types}, including those which are not necessarily defined
+by their values; a general apparatus for computation-respecting refinements will
+allow us to define, for instance, a type for partial functions.
+
+Then, for any refinement $\Refines{\Upsilon}{\phi}{\tau}$, we say
+$\IsType{\Upsilon}{\phi}{\tau}$ just when for any
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+\end{tikzcd}%
+, and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
+$\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
+can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
+Stated internally in the topos, the meaning of parametric typehood for
+$\phi$ is perhaps a bit more clear:
+\[
+  \Define{
+    \IsType{-}{\phi}{\tau}
+  }{
+    \prod_{L,M,N\in\Exprs_\tau}
+    {\phi(L,N)}^{L\sim_\tau M \times \phi(M,N)}
+  }
+\]
+
 \ifdraft{}{

### 2015-12-19 paper.tex
comment out stuff pending fix

@@
 
-In general, we must explain typehood for parametric refinements
-$\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
-Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
-transformation
-$\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
-Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
-is a unique $\psi$ such that the following diagram commutes, where
-$\lambda(\chi)$ is the exponential transpose of $\chi$:
-\[
-  \IsValueType{\Upsilon}{\phi}{\tau}
-  \triangleq
-  \begin{tikzcd}[sep=huge]
-    \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
-      \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
-      \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
-    \IMode{\Omega}\\
-    \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
-      \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
-  \end{tikzcd}
-\]
-
-Contrary to Martin-L\"of, however, we wish to consider a full range of
-\emph{computational types}, including those which are not necessarily defined
-by their values; a general apparatus for computation-respecting refinements will
-allow us to define, for instance, a type for partial functions.
-
-Then, for any refinement $\Refines{\Upsilon}{\phi}{\tau}$, we say
-$\IsType{\Upsilon}{\phi}{\tau}$ just when for any
-\begin{tikzcd}[cramped]
-  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
-\end{tikzcd}%
-, and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
-$\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
-can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
-Stated internally in the topos, the meaning of parametric typehood for
-$\phi$ is perhaps a bit more clear:
-\[
-  \Define{
-    \IsType{-}{\phi}{\tau}
-  }{
-    \prod_{L,M,N\in\Exprs_\tau}
-    {\phi(L,N)}^{L\sim_\tau M \times \phi(M,N)}
-  }
-\]
+%% In general, we must explain typehood for parametric refinements
+%% $\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
+%% Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
+%% transformation
+%% $\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
+%% Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
+%% is a unique $\psi$ such that the following diagram commutes, where
+%% $\lambda(\chi)$ is the exponential transpose of $\chi$:
+%% \[
+%%   \IsValueType{\Upsilon}{\phi}{\tau}
+%%   \triangleq
+%%   \begin{tikzcd}[sep=huge]
+%%     \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
+%%       \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
+%%       \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
+%%     \IMode{\Omega}\\
+%%     \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
+%%       \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
+%%   \end{tikzcd}
+%% \]
+%%
+%% Equivalently, and perhaps more simply, we can express value typehood internally
+%% in our topos as a presheaf:
+%% \[
+%%   \Define{
+%%     \IsValueType{-}{\phi}{\tau}
+%%   }{
+%%     \sum_{\psi\in\Pow{\Values_\tau^2}}
+%%     \prod_{M,N\in\Values_\tau}
+%%     {\phi(M,N)}^{\psi(M,N)}
+%%   }
+%% \]
+%%
+%% Contrary to Martin-L\"of, however, we wish to consider a full range of
+%% \emph{computational types}, including those which are not necessarily defined
+%% by their values; a general apparatus for computation-respecting refinements will
+%% allow us to define, for instance, a type for partial functions.
+%%
+%% Then, for any refinement $\Refines{\Upsilon}{\phi}{\tau}$, we say
+%% $\IsType{\Upsilon}{\phi}{\tau}$ just when for any
+%% \begin{tikzcd}[cramped]
+%%   \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+%% \end{tikzcd}%
+%% , and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
+%% $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
+%% can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
+%% Stated internally as a presheaf in the topos, the meaning of parametric typehood for
+%% $\phi$ is perhaps a bit more clear:
+%% \[
+%%   \Define{
+%%     \IsType{-}{\phi}{\tau}
+%%   }{
+%%     \prod_{L,M,N\in\Exprs_\tau}
+%%     {\phi(L,N)}^{L\sim_\tau M \times \phi(M,N)}
+%%   }
+%% \]
+%%
+%% \begin{thm}
+%%   For any refinement $\Refines{\Upsilon}{\phi}{\tau}$, from
+%%   $\IsValueType{\Upsilon}{\phi}{\tau}$ we may conclude
+%%   $\IsType{\Upsilon}{\phi}{\tau}$.
+%% \end{thm}
+%% \begin{proof}
+%%   Fix a renaming
+%%   \begin{tikzcd}[cramped]
+%%     \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+%%   \end{tikzcd}%
+%%   , terms $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$ such that
+%%   $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$;
+%%   by hypothesis, we also have some
+%%   $\Of{\psi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau^2}}}$,
+%%   factoring through $\phi$.
+%%
+%%   We need to show that $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
+%%
+%% \end{proof}
 

### 2015-12-19 paper.tex
restore general typehood definition

@@
 
-Types, on the other hand, are intrinsically tied up with computation.  In order
-to make this concrete, let us first define what it takes for a refinement to be
-a type, in the sense of Martin-L\"of---that is, a \emph{value type}. In the
-simple case, where we have a global refinement $\GloballyRefines{\phi}{\tau}$,
-we will say $\GlobalIsValueType{\phi}{\tau}$ when there exists a PER $\psi$ on
-\emph{values} that factors uniquely through $\phi$.
+% In order
+% to make this concrete, let us first define what it takes for a refinement to be
+% a type, in the sense of Martin-L\"of---that is, a \emph{value type}. In the
+% simple case, where we have a global refinement $\GloballyRefines{\phi}{\tau}$,
+% we will say $\GlobalIsValueType{\phi}{\tau}$ when there exists a PER $\psi$ on
+% \emph{values} that factors uniquely through $\phi$.
 
@@
 %% Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
-%% is a unique $\psi$ such that the following diagram commutes, where
+%% is a unique $\Of{\psi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau}}}$ such that
+%% the following diagram commutes, where
 %% $\lambda(\chi)$ is the exponential transpose of $\chi$:
@@
 %% \]
-%%
-%% Contrary to Martin-L\"of, however, we wish to consider a full range of
-%% \emph{computational types}, including those which are not necessarily defined
-%% by their values; a general apparatus for computation-respecting refinements will
-%% allow us to define, for instance, a type for partial functions.
-%%
-%% Then, for any refinement $\Refines{\Upsilon}{\phi}{\tau}$, we say
-%% $\IsType{\Upsilon}{\phi}{\tau}$ just when for any
-%% \begin{tikzcd}[cramped]
-%%   \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
-%% \end{tikzcd}%
-%% , and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
-%% $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
-%% can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
-%% Stated internally as a presheaf in the topos, the meaning of parametric typehood for
-%% $\phi$ is perhaps a bit more clear:
-%% \[
-%%   \Define{
-%%     \IsType{-}{\phi}{\tau}
-%%   }{
-%%     \prod_{L,M,N\in\Exprs_\tau}
-%%     {\phi(L,N)}^{L\sim_\tau M \times \phi(M,N)}
-%%   }
-%% \]
-%%
+
+Following Martin-L\"of, a type is a refinement which behaves well with respect
+to computation, in some specified sense; contrary to Martin-L\"of, however, we wish to
+consider a full range of \emph{computational types}, including those which are
+not necessarily defined by their values; a general apparatus for
+computation-respecting refinements will allow us to define, for instance, a
+type for partial functions.
+
+\begin{definition}[Typehood]
+  For any refinement $\Refines{\Upsilon}{\phi}{\tau}$, we say
+  $\IsType{\Upsilon}{\phi}{\tau}$ just when for any
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}%
+  , and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
+  $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
+  can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
+\end{definition}
+
+Stated internally as a presheaf in the topos, the meaning of parametric typehood for
+$\phi$ is perhaps a bit more clear:
+\[
+  \Define{
+    \IsType{-}{\phi}{\tau}
+  }{
+    \prod_{L,M,N\in\Exprs_\tau}
+    {\phi(L,N)}^{L\sim_\tau M \times \phi(M,N)}
+  }
+\]
+
 %% \begin{thm}

### 2015-12-19 paper.tex
finish definitions of type hood, value typehood

@@
 
-%% \begin{thm}
-%%   For any refinement $\Refines{\Upsilon}{\phi}{\tau}$, from
-%%   $\IsValueType{\Upsilon}{\phi}{\tau}$ we may conclude
-%%   $\IsType{\Upsilon}{\phi}{\tau}$.
-%% \end{thm}
-%% \begin{proof}
-%%   Fix a renaming
-%%   \begin{tikzcd}[cramped]
-%%     \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
-%%   \end{tikzcd}%
-%%   , terms $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$ such that
-%%   $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$;
-%%   by hypothesis, we also have some
-%%   $\Of{\psi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau^2}}}$,
-%%   factoring through $\phi$.
-%%
-%%   We need to show that $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
-%%
-%% \end{proof}
+Now, let us return briefly to Martin-L\"of's notion of type, which we call a
+\emph{value type}. The minimal data of such a type's definition consists in a
+partial equivalence relation on values,
+$\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.
+It is easy to turn such a relation into a computational type
+$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows:
+\[
+  \begin{tikzcd}[sep=huge]
+    \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
+      \arrow[r, "\IMode{\lambda(\phi)}"]
+      \arrow[d, swap, "\IMode{\Pair{\ArrId{}}{\iota}}"]
+      &
+    \IMode{\Omega}\\
+    \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
+    \arrow[ur, swap, dotted, "\OMode{\lambda(\VToCType(\phi))}"]
+  \end{tikzcd}
+\]
+where $\VToCType(\phi)$ is the computational extension of $\phi$ in the following sense:
+\begin{align*}
+  \ADefine{
+    \VToCType(\phi)(M,N)
+  }{
+    \forall M', N' \in \Values_\tau.\
+    M\sim_\tau M' \land N\sim_\tau N' \implies \phi(M',N')
+  }
+\end{align*}
+
+\begin{thm}
+When viewed as an internal functor
+$\PERs{\Values_\tau}\to\PERs{\Exprs_\tau}$, $\VToCType$ is the left adjoint to
+the canonical inclusion
+$\Of{\iota}{\PERs{\Exprs_\tau}\inj\PERs{\Values_\tau}}$.
+\end{thm}
+\begin{proof}
+  We will exhibit a counit $\Of{\epsilon}{\VToCType\circ\iota\to
+  \ArrId{\PERs{\Values_\tau}}}$ and a unit
+  $\Of{\eta}{\ArrId{\PERs{\Exprs_\tau}}\to\iota\circ\VToCType}$ for
+  the adjunction $\IMode{\VToCType}\dashv\IMode{\iota}$.
+
+  The counit witnesses the fact that for any refinement
+  $\Of{\phi}{\PERs{\Exprs_\tau}}$ and terms $\Member{M,N}{\Exprs_\tau}$, if
+  $\VToCType(\iota(\phi))(M,N)$, then $\phi(M,N)$; this is evidently the case,
+  because the premise only obtains if $M,N$ are computationally equal to
+  values.  The unit witnesses the fact that for any value PER
+  $\Of{\psi}{\PERs{\Values_\tau}}$ and values $\Member{M,N}{\Values_\tau}$, if
+  $\psi(M,N)$, then $\iota(\VToCType(\psi))(M,N)$; this is clearly true, since
+  $\VToCType(\psi)$ can do nothing but coarsen $\psi$'s relation on values.
+
+\end{proof}
+
+\begin{definition}[Closure]
+  For any computational type $\IsType{\Upsilon}{\phi}{\tau}$, let
+  $\Define{\underline{\phi}}{\VToCType(\iota(\phi))}$ be called $\phi$'s
+  \emph{closure}, which we will use shortly in order to give a formal definition
+  to the notion of value types.
+\end{definition}
+
+\begin{definition}[Value Types]
+  We can give a simple characterization of Martin-L\"of's \emph{value types} in
+  terms of our computation types. For any type $\IsType{\Upsilon}{\phi}{\tau}$,
+  we will say $\IsValueType{\Upsilon}{\phi}{\tau}$ in case $\phi$ is equal to
+  its closure, i.e.\ $\EqRefines{\Upsilon}{\phi}{\underline{\phi}}{\tau}$.
+\end{definition}
 

### 2015-12-19 paper.tex
Fix backwards yoneda ⇒ coyoneda

@@
   So, $\Refines{\Upsilon}{\phi}{\tau}$ obtains just when we have
-  $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of
-  explaining such judgments in terms of the presheaf topos is that we do not
-  need to deal with the complexities of quantifying over renamings, since this
-  is implicit in the definition of the exponential of presheaves,
+  $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of explaining such
+  judgments in terms of the presheaf topos is that we do not need to deal with
+  the complexities of quantifying over renamings, since this is implicit in the
+  definition of the exponential of presheaves, where
+  $\Define{\Coyoneda{\Upsilon}}{\Upsilon\inj-}$ is the co-yoneda embedding:
   \begin{align*}
@@
       \Hom{\Sets^\SCtx}{
-        \Yoneda{\Upsilon}\times A
+        \Coyoneda{\Upsilon}\times A
       }{
@@
 %% transformation
-%% $\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
+%% $\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
 %% Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
-%% is a unique $\Of{\psi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau}}}$ such that
+%% is a unique $\Of{\psi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$ such that
 %% the following diagram commutes, where
@@
 %%   \begin{tikzcd}[sep=huge]
-%%     \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
+%%     \IMode{\Coyoneda{\Upsilon}\times\Values_\tau^2}
 %%       \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
@@
 %%     \IMode{\Omega}\\
-%%     \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
+%%     \IMode{\Coyoneda{\Upsilon}\times\Exprs_\tau^2}
 %%       \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
@@
 partial equivalence relation on values,
-$\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.
+$\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.
 It is easy to turn such a relation into a computational type
@@
   \begin{tikzcd}[sep=huge]
-    \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
+    \IMode{\Coyoneda{\Upsilon}\times\Values_\tau^2}
       \arrow[r, "\IMode{\lambda(\phi)}"]
@@
     \IMode{\Omega}\\
-    \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
+    \IMode{\Coyoneda{\Upsilon}\times\Exprs_\tau^2}
     \arrow[ur, swap, dotted, "\OMode{\lambda(\VToCType(\phi))}"]

### 2015-12-19 paper.tex
dead code
fix layout & notation

@@
   $\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}$, you can conclude
-  $\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{N_0}}}{\Subst{M_1}{x}{\Rename{\rho}{N_1}}}{\Subst{M_0}{x}{\phi\rho}}$.
+  $\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{N_0}}}{\Subst{M_1}{x}{\Rename{\rho}{N_1}}}{\Subst{M_0}{x}{\Rename{\rho}{\phi}}}$.
   In other words:
@@
         .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
-        \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)
-        \Rightarrow
+        \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)\\
+        \quad\Rightarrow
         \left(%
@@
             \Subst{M_1}{x}{\Rename{\rho}{N_1}}%
-          }{\Subst{M_0}{x}{\phi\rho}
+          }{\Subst{M_0}{x}{\Rename{\rho}{\phi}}
           }%
@@
 
-% In order
-% to make this concrete, let us first define what it takes for a refinement to be
-% a type, in the sense of Martin-L\"of---that is, a \emph{value type}. In the
-% simple case, where we have a global refinement $\GloballyRefines{\phi}{\tau}$,
-% we will say $\GlobalIsValueType{\phi}{\tau}$ when there exists a PER $\psi$ on
-% \emph{values} that factors uniquely through $\phi$.
-
-%% In general, we must explain typehood for parametric refinements
-%% $\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
-%% Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
-%% transformation
-%% $\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
-%% Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
-%% is a unique $\Of{\psi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$ such that
-%% the following diagram commutes, where
-%% $\lambda(\chi)$ is the exponential transpose of $\chi$:
-%% \[
-%%   \IsValueType{\Upsilon}{\phi}{\tau}
-%%   \triangleq
-%%   \begin{tikzcd}[sep=huge]
-%%     \IMode{\Coyoneda{\Upsilon}\times\Values_\tau^2}
-%%       \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
-%%       \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
-%%     \IMode{\Omega}\\
-%%     \IMode{\Coyoneda{\Upsilon}\times\Exprs_\tau^2}
-%%       \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
-%%   \end{tikzcd}
-%% \]
-%%
-%% Equivalently, and perhaps more simply, we can express value typehood internally
-%% in our topos as a presheaf:
-%% \[
-%%   \Define{
-%%     \IsValueType{-}{\phi}{\tau}
-%%   }{
-%%     \sum_{\psi\in\Pow{\Values_\tau^2}}
-%%     \prod_{M,N\in\Values_\tau}
-%%     {\phi(M,N)}^{\psi(M,N)}
-%%   }
-%% \]
-
 Following Martin-L\"of, a type is a refinement which behaves well with respect

### 2015-12-19 paper.tex
wording

@@
 
-Stated internally as a presheaf in the topos, the meaning of parametric typehood for
+Stated internally as an object in the topos, the meaning of parametric typehood for
 $\phi$ is perhaps a bit more clear:
@@
   We can give a simple characterization of Martin-L\"of's \emph{value types} in
-  terms of our computation types. For any type $\IsType{\Upsilon}{\phi}{\tau}$,
+  terms of our more general computational types. For any type $\IsType{\Upsilon}{\phi}{\tau}$,
   we will say $\IsValueType{\Upsilon}{\phi}{\tau}$ in case $\phi$ is equal to

### 2015-12-19 paper.tex
fix up adjunction

@@
 
+Let us write $\Types{\tau}$ for the object of types,
+\[
+  \Define{
+    \Types{\tau}
+  }{
+    \MkSet{
+      \phi\in\PERs{\Exprs_\tau}
+      \mid
+      -\parallel\phi\ \textit{type}_\tau
+    }
+  }
+\]
+
 Now, let us return briefly to Martin-L\"of's notion of type, which we call a
@@
 When viewed as an internal functor
-$\PERs{\Values_\tau}\to\PERs{\Exprs_\tau}$, $\VToCType$ is the left adjoint to
+$\PERs{\Values_\tau}\to\Types{\tau}$, $\VToCType$ is the left adjoint to
 the canonical inclusion
-$\Of{\iota}{\PERs{\Exprs_\tau}\inj\PERs{\Values_\tau}}$.
+$\Of{\iota}{\Types{\tau}\inj\PERs{\Values_\tau}}$.
 \end{thm}
@@
   \ArrId{\PERs{\Values_\tau}}}$ and a unit
-  $\Of{\eta}{\ArrId{\PERs{\Exprs_\tau}}\to\iota\circ\VToCType}$ for
+  $\Of{\eta}{\ArrId{\Types{\tau}}\to\iota\circ\VToCType}$ for
   the adjunction $\IMode{\VToCType}\dashv\IMode{\iota}$.
@@
   The counit witnesses the fact that for any refinement
-  $\Of{\phi}{\PERs{\Exprs_\tau}}$ and terms $\Member{M,N}{\Exprs_\tau}$, if
+  $\Of{\phi}{\Types{\tau}}$ and terms $\Member{M,N}{\Exprs_\tau}$, if
   $\VToCType(\iota(\phi))(M,N)$, then $\phi(M,N)$; this is evidently the case,
   because the premise only obtains if $M,N$ are computationally equal to
-  values.  The unit witnesses the fact that for any value PER
+  values, and $\phi$ respects computational equivalence by virtue of being a
+  type.  The unit witnesses the fact that for any value PER
   $\Of{\psi}{\PERs{\Values_\tau}}$ and values $\Member{M,N}{\Values_\tau}$, if

### 2015-12-19 paper.tex
unambiguous naming
closure -> value closure

@@
       \arrow[r, "\IMode{\lambda(\phi)}"]
-      \arrow[d, swap, "\IMode{\Pair{\ArrId{}}{\iota}}"]
+      \arrow[d, hook, swap, "\IMode{\Pair{\ArrId{}}{i}}"]
       &
@@
 
-\begin{definition}[Closure]
+\begin{definition}[Value Closure]
   For any computational type $\IsType{\Upsilon}{\phi}{\tau}$, let
   $\Define{\underline{\phi}}{\VToCType(\iota(\phi))}$ be called $\phi$'s
-  \emph{closure}, which we will use shortly in order to give a formal definition
+  \emph{value closure}, which we will use shortly in order to give a formal definition
   to the notion of value types.
@@
   we will say $\IsValueType{\Upsilon}{\phi}{\tau}$ in case $\phi$ is equal to
-  its closure, i.e.\ $\EqRefines{\Upsilon}{\phi}{\underline{\phi}}{\tau}$.
+  its value closure, i.e.\ $\EqRefines{\Upsilon}{\phi}{\underline{\phi}}{\tau}$.
 \end{definition}

### 2015-12-19 paper.tex
layout
notation
Update K, O to be families of copresheaves, following abt paper

@@
 several judgments and their semantical explanations. Let us first define the
-presheaf of $\tau$-sorted expressions, $\Exprs_\tau$ as follows
+copresheaf of $\tau$-sorted expressions, $\Exprs_\tau$ as follows
 \[
@@
 
-We will also write $\Exprs_\tau$ for the presheaf $\Exprs_\tau(-\parallel\cdot)$ on $\SCtx$.
+We will also write $\Exprs_\tau$ for the copresheaf $\Exprs_\tau(-\parallel\cdot)$ on $\SCtx$.
 
@@
   the pointwise style in which we have expressed it. Considered internally to
-  the presheaf topos $\Sets^\SCtx$, a refinement is merely a section of the
+  the copresheaf topos $\Sets^\SCtx$, a refinement is merely a section of the
   object of $\Exprs_\tau$-PERs $\PERs{\Exprs_\tau}$, defined
@@
   $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of explaining such
-  judgments in terms of the presheaf topos is that we do not need to deal with
+  judgments as objects in the topos is that we do not need to deal with
   the complexities of quantifying over renamings, since this is implicit in the
@@
       (\Upsilon\inj\Upsilon')
-      \Rightarrow
+      \implies
       {B(\Upsilon')}^{A(\Upsilon')}
@@
         .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
-        \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)
-        \Rightarrow
+        \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)\\
+        \quad\implies
         \left(\FEqRefines{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{\phi}}}{\Subst{M_1}{x}{\Rename{\rho}{\psi}}}{\tau}\right)
@@
         \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)\\
-        \quad\Rightarrow
+        \quad\implies
         \left(%
@@
 $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished
-presheaf $\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
-\emph{canonical} operators such that $\IsSubsetEq{\CanOperators}{\Operators}$.
+arity-indexed family of copresheaves
+$\HypJ{\Of{\CanOperators_a}{\Sets^\SCtx}}{\Member{a}{\Arities}}$ of
+\emph{canonical} operators such that
+$\IsSubsetEq{\CanOperators_a}{\Operators_a}$ for each arity $a$.
 

### 2015-12-20 paper.tex
add section on families of refinements

@@
 \newcommand\TyBot[1]{\bot_{#1}}
-\newcommand\TyMeet[1]{\cap_{#1}}
-\newcommand\TyJoin[1]{\cup_{#1}}
 
@@
 $\GloballyRefines{\TyTop{\tau}}{\tau}$ identifies all terms
-$\Member{M,N}{\Exprs_\tau(\Upsilon)}$.  The join
-$\Refines{\Upsilon}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
+$\Member{M,N}{\Exprs_\tau(\Upsilon)}$.  The join (union)
+$\Refines{\Upsilon}{\phi\cup\psi}{\tau}$ of two refinements
 $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$ is defined as
@@
   \infer{
-    \RMemEq{\Upsilon}{M}{N}{\phi\TyJoin{\tau}\psi}
+    \RMemEq{\Upsilon}{M}{N}{\phi\cup\psi}
   }{
@@
 
-Finally the meet $\Refines{\Upsilon}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
+Finally the meet (intersection) $\Refines{\Upsilon}{\phi\cap\psi}{\tau}$ of two refinements
 $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$ is defined
@@
   \infer{
-    \RMemEq{\Upsilon}{M}{N}{\phi\TyMeet{\tau}\psi}
+    \RMemEq{\Upsilon}{M}{N}{\phi\cap\psi}
   }{
@@
 lattice.
+
+\subsection{Families of Refinements}
+
+Given a family of refinements (i.e. a functional refinement)
+$\FRefines{\Upsilon}{x:\phi}{\psi}{\tau}$ such that
+$\Refines{\Upsilon}{\phi}{\sigma}$, including the family intersection, the
+dependent intersection, the family union, and the set comprehension.
+
+\paragraph{Family intersection}
+The intersection $\Refines{\Upsilon}{\bigcap_{(x:\phi)}\psi}{\tau}$ of a family
+of refinements is defined by the following rule, for any
+\begin{tikzcd}[cramped, ampersand replacement = \&]
+  \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+\end{tikzcd}:
+\[
+  \infer={
+    \RMemEq{\Upsilon'}{M}{N}{
+      \bigcap_{(x:\phi)}\psi
+    }
+  }{
+    \FRMemEq{\Upsilon'}{x:\phi}{M}{N}{\Subst{M}{x}{\psi}}
+  }
+\]
+
+The intersection of a family of refinements is a kind of uniform universal
+quantification, in the sense that the terms themselves bear no trace of the
+quantification, but are guaranteed in the metalogic to be related in all fibres
+of $\psi$.
+
+In practice, the family intersection is perhaps the most-used type constructor
+in Nuprl developments, as it facilitates the simultaneous specification and
+implementation of programs in type theory, without causing details of the
+program's verification to leak into the extracted algorithm. Additionally, the
+family intersection can be used to implement coinductive types.
+
+\paragraph{Dependent intersection}
+The dependent intersection $\Refines{\Upsilon}{(x:\phi)\cap\psi}{\sigma}$
+contains the members $M$ of $\phi$ which are also members of
+$\Subst{M}{x}{\psi}$:
+\[
+  \infer={
+    \RMemEq{\Upsilon'}{M}{N}{
+      (x:\phi)\cap\psi
+    }
+  }{
+    \RMemEq{\Upsilon'}{M}{N}{\phi} &
+    \RMemEq{\Upsilon'}{M}{N}{
+      \Subst{M}{x}{\psi}
+    }
+  }
+\]
+
+The dependent intersection, first introduced by Kopylov in \cite{kopylov:2000},
+are a generalization of the binary intersection to the dependent case, not to
+be confused with the intersection of a family of sets. The dependent
+intersection has many applications, including the compositional representation
+of \emph{dependent records} or modules in Type Theory.
+
+\paragraph{Family union}
+The union $\Refines{\Upsilon}{\bigcup_{(x:\phi)}\psi}{\tau}$ of a family of
+refinements is defined as by the following rule:
+\[
+  \infer{
+    \RMemEq{\Upsilon'}{M}{N}{
+      \bigcup_{(x:\phi)}\psi
+    }
+  }{
+    \RMemEq{\Upsilon'}{L_0}{L_1}{\phi} &
+    \RMemEq{\Upsilon'}{M}{N}{
+      \Subst{L_0}{x}{\psi}
+    }
+  }
+\]
+
+The union of a family of refinements can be used to implement inductive types,
+among other things.
+
+\paragraph{Set comprehension}
+
+We can also use a family of refinements to define something analogous to set
+comprehension, $\Refines{\Upsilon'}{\MkSet{x:\phi\mid\psi}}{\sigma}$, defined
+as follows:
+\[
+  \infer{
+    \RMemEq{\Upsilon'}{M}{N}{
+      \MkSet{x:\phi\mid\psi}
+    }
+  }{
+    \RMemEq{\Upsilon'}{M}{N}{\phi} &
+    \RMemEq{\Upsilon'}{L_0}{L_1}{
+      \Subst{M}{x}{\psi}
+    }
+  }
+\]
+
+The set comprehension $\MkSet{x:\phi\mid\psi}$ contains just those elements of
+$\phi$ such that their fibres of $\psi$ are inhabited; this can be seen as an
+``implicit'' version of existential quantification, and can be used to
+simultaneously specify and implement program without the evidence for $\psi$
+intruding into the extracted algorithm.
 

### 2015-12-20 paper.tex
correct outdated notation
Explain why we're defining refinements for arbitrary rho : Y -> Y'

@@
 $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$ is defined as
-follows:
+follows, for any
+\begin{tikzcd}[cramped, ampersand replacement = \&]
+  \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+\end{tikzcd}:
 
@@
   \infer{
-    \RMemEq{\Upsilon}{M}{N}{\phi\cup\psi}
+    \RMemEq{\Upsilon'}{M}{N}{\phi\cup\psi}
   }{
-    \RMemEq{\Upsilon}{M}{N}{\phi}
+    \RMemEq{\Upsilon'}{M}{N}{\phi}
   }\qquad
   \infer{
-    \RMemEq{\Upsilon}{M}{N}{\phi\cup\psi}
+    \RMemEq{\Upsilon'}{M}{N}{\phi\cup\psi}
   }{
-    \RMemEq{\Upsilon}{M}{N}{\psi}
+    \RMemEq{\Upsilon'}{M}{N}{\psi}
   }
@@
   \infer{
-    \RMemEq{\Upsilon}{M}{N}{\phi\cap\psi}
+    \RMemEq{\Upsilon'}{M}{N}{\phi\cap\psi}
   }{
-    \RMemEq{\Upsilon}{M}{N}{\phi} &
-    \RMemEq{\Upsilon}{M}{N}{\psi}
+    \RMemEq{\Upsilon'}{M}{N}{\phi} &
+    \RMemEq{\Upsilon'}{M}{N}{\psi}
   }
@@
 lattice.
+
+\begin{remark}
+  Observe that \emph{definition} of a refinement
+  $\Refines{\Upsilon}{\phi}{\tau}$ is given in terms of an arbitrary
+  $\Upsilon'$ that is related to $\Upsilon$ by a renaming. This is because
+  $\Upsilon$ is the stage at which the refinement has come into being, but the
+  definition of the refinement must be applicable at all subsequent stages.
+\end{remark}
 
@@
       M\equiv\App{\vartheta}{\vec{E}}
-      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
+      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators_a(\Upsilon)\ni\vartheta
     }

### 2015-12-21 paper.tex
Explain λ notation
Define approx-star, extensionality
(close #8)

@@
 \newcommand\SBinRel[4]{\IMode{#2}\parallel\IMode{#3}\mathrel{#1}\IMode{#4}}
+\newcommand\OSBinRel[5]{\IMode{#2}\parallel\IMode{#3}\vDash\IMode{#4}\mathrel{#1}\IMode{#5}}
+
+\begin{notation*}
+  For a family of copresheaves $\HypJ{\Of{X_i}{\Sets^\SCtx}}{\Member{i}{I}}$, we will
+  use $X^{[\vec{\i}]}$ as a shorthand for the following iterated product, as
+  in \cite{sterling-morrison:2015}:
+
+  \[
+    \Define{
+      X^{[\vec{\i}]}
+    }{
+      \prod_{i\in\vec{\i}}
+        X_i
+    }
+  \]
+\end{notation*}
 
@@
 also write the relation as a judgment scheme
-$\SBinRel{R_\tau}{\Upsilon}{M}{N}$. Now, we can always extend $R_\tau$ to a new
-family of relations
-$\Of{R_v}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for any valence
+$\SBinRel{R_\tau}{\Upsilon}{M}{N}$. First of all, we can lift such a relation
+to operate on open terms by quantifying over substitutions, as follows:
+\[
+  \infer{
+    \OSBinRel{R_\tau}{\Upsilon}{\vec{x}:\vec{\sigma}}{M}{N}
+  }{
+    \forall
+      \begin{tikzcd}[cramped, ampersand replacement = \&]
+         \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+      \end{tikzcd}.
+    \quad
+    \forall \Of{\vec{X}}{\Exprs^{[\vec{\sigma}]}(\Upsilon')}.
+    \quad
+    \SBinRel{R_\tau}{\Upsilon'}{
+      \Subst{\vec{X}}{\vec{x}}{\Rename{\rho}{M}}
+    }{
+      \Subst{\vec{X}}{\vec{x}}{\Rename{\rho}{M}}
+    }
+  }
+\]
+
+Next, we can always extend $R_\tau$ to a new
+family of valenced-indexed relations
+$\Of{R_v}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for
 $\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
@@
       \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon}\cup\vec{u}\cup\vec{v}}.\quad
-      \forall
-        \begin{tikzcd}[cramped, ampersand replacement = \&]
-           \IMode{\Upsilon,\vec{w}:\vec{\sigma}} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
-        \end{tikzcd}.\quad
-      \forall\Of{\vec{X}}{\Exprs^{[\vec{\tau}]}(\Upsilon')}.\\
-      \quad
-      \SBinRel{R_\tau}{\Upsilon'}{
-        \Subst{\vec{X}}{\vec{x}}{
-          \Rename{(\rho\circ \vec{u}\mapsto\vec{w})}{M}
+      \forall\IMode{\vec{z}}\mathrel{\#}\IMode{\vec{x}\cup\vec{y}}.\\
+      \OSBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{\vec{z}:\vec{\tau}}{
+        \Subst{\vec{z}}{\vec{x}}{
+          \Rename{(\vec{u}\mapsto\vec{w})}{M}
         }
       }{
-        \Subst{\vec{X}}{\vec{y}}{
-          \Rename{(\rho\circ \vec{u}\mapsto\vec{w})}{M}
+        \Subst{\vec{z}}{\vec{y}}{
+          \Rename{(\vec{v}\mapsto\vec{w})}{M}
         }
@@
 \]
-
-\begin{notation*}
-  For a family of copresheaves $\HypJ{\Of{X_\tau}{\Sets^\SCtx}}{\Member{\tau}{\Sorts}}$, we will
-  use $X^{[\vec{\tau}]}$ as a shorthand for the following iterated product, as
-  in \cite{sterling-morrison:2015}:
-
-  \[
-    \Define{
-      X^{[\vec{\tau}]}
-    }{
-      \prod_{\tau\in\vec{\tau}}
-        X_\tau
-    }
-  \]
-\end{notation*}
 
@@
   \]
+\end{definition}
+
+\subsection{Extensionality and Observational Congruence}
+
+A sort-indexed family of binary relations
+$\HypJ{\Of{R_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}}{\Member{\tau}{\Sorts}}$
+is a \emph{congruence} when for all operators
+$\IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$, the following rule
+obtains:
+\[
+  \infer{
+    \SBinRel{R_\tau}{\Upsilon}{\vartheta(\vec{E})}{\vartheta(\vec{F})}
+  }{
+    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
+  }
+  \tag{Congruence}
+\]
+
+In general, computational equivalence is not a congruence unless we impose a
+certain extensionality condition on operators. First, let us define a new
+version of approximation $\preccurlyeq^\star_\tau$ over open terms:
+\begin{gather*}
+  \infer[\textit{var}]{
+    \OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{x}{y}
+  }{
+    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{x}{y}
+  }
+  \qquad
+  \infer[\textit{app}]{
+    \OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{
+      \App{\vartheta}{\vec{E}}
+    }{
+      N
+    }
+  }{
+    \IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}} &
+    \OSBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon}{\Gamma}{
+      \vec{E}
+    }{
+      \vec{F}
+    } &
+    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{
+      \App{\vartheta}{\vec{F}}
+    }{
+      N
+    }
+  }
+\end{gather*}
+
+We will write $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{M}{N}$ for
+$\OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\cdot}{M}{N}$.  By definition,
+$\preccurlyeq^\star_\tau$ is a congruence; to show that $\sim_\tau$ is a
+congruence, it suffices to show that the following inference is valid:
+\[
+  \infer[?]{
+    \OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{M}{N}
+  }{
+    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{M}{N}
+  }
+\]
+
+We will not reproduce the full proof of the above rule here (see
+\cite{howe:1989}), but will simply remark that it holds only for what is called
+an \emph{extensional lcs}, an lcs in which each operator is extensional in the
+following sense:
+
+\begin{definition}[Extensionality]
+
+  An operator $\IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$ is
+  \emph{extensional} when, for any
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}%
+  , for all $\Member{\vec{E},\vec{F}}{\BTms^{[\vec{v}]}(\Upsilon')}$ and
+  $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$, if we can conclude
+  $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{M}{\App{\vartheta}{\vec{F}}}$
+  from the following conditions:
+  %
+  \begin{enumerate}
+    \item
+      We have $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{N}$.
+
+    \item
+      We have $\SBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{F}}$.
+
+    \item
+      For any
+      \begin{tikzcd}[cramped]
+        \IMode{\Upsilon'} \arrow[r,hook,"\IMode{\rho'}"] &\IMode{\Upsilon''}
+      \end{tikzcd}%
+      , for all $\Member{\sigma}{\Sorts}$ and $\Member{X,X',Y}{\Exprs_\sigma(\Upsilon'')}$, if
+      from $\EvalN{\Upsilon''}{\sigma}{k'}{X}{X'}$ for $\IMode{k'}<\IMode{k}$ and
+      $\SBinRel{\preccurlyeq^\star_\sigma}{\Upsilon''}{X}{Y}$, we can conclude
+      $\SBinRel{\preccurlyeq^\star_\sigma}{\Upsilon''}{X'}{Y}$.
+  \end{enumerate}
+
 \end{definition}
@@
 partial equivalence relation on values,
-$\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.
-It is easy to turn such a relation into a computational type
-$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows:
+$\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.  It
+is easy to turn such a relation into a computational type
+$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows, where $\lambda(\chi)$
+is the exponential transpose of $\chi$:
 \[

### 2015-12-21 paper.tex
prove approx-refl, and extensionality of canonical operators.

@@
 
-\newcommand\CanOperators{\mathcal{K}}
-
 \newcommand\IsAbtUnmoded[5]{
@@
 Then, a \emph{lazy computation system} (lcs) is a lazy computation language
-$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an $\SCtx$-indexed
+$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with a partial, $\SCtx$-indexed
 $\equiv_\alpha$-functional evaluation relation
@@
   For a family of copresheaves $\HypJ{\Of{X_i}{\Sets^\SCtx}}{\Member{i}{I}}$, we will
-  use $X^{[\vec{\i}]}$ as a shorthand for the following iterated product, as
+  use $X^{[\vec{i}]}$ as a shorthand for the following iterated product, as
   in \cite{sterling-morrison:2015}:
@@
     \Define{
-      X^{[\vec{\i}]}
-    }{
-      \prod_{i\in\vec{\i}}
+      X^{[\vec{i}]}
+    }{
+      \prod_{i\in\vec{i}}
         X_i
@@
 $\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}}$ such that
-$\IMode{\Upsilon}\Vdash_{\IMode{\CanOperators}}\IMode{\vartheta}:\OMode{\MkArity{\vec{v}}{\tau}}$,
+$\IsCanOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$,
 for any
@@
 
+\begin{thm}[Reflexivity]\label{thm:approx-refl}
+  For all $\Member{M}{\Exprs_\tau(\Upsilon)}$, we have $\SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{M}$.
+\end{thm}
+\begin{proof}
+  The proof proceeds by coinduction, using the relation
+  $\IMode{M,N}\mapsto\OMode{M = N}$ as our motive; it, then, suffices to show
+  that for all $M$, $\SBinRel{[=]}{\Upsilon}{M}{M}$. Expanding the definition
+  of $[-]$, this means that we have to show that when
+  $\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\vec{E}}}$, then
+  $\IsEq{\vec{E}}{\vec{E}}$; but this is always the case, because of the
+  functionality of the evaluation relation.
+\end{proof}
+
 \begin{definition}[Computational Equivalence]
@@
     \IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}} &
+    \exists\IMode{\vec{F}}. &
     \OSBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon}{\Gamma}{
@@
 \begin{definition}[Extensionality]
+  \label{def:extensionality}
 
@@
   $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$, if we can conclude
-  $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{M}{\App{\vartheta}{\vec{F}}}$
+  $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{M}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
   from the following conditions:
@@
     \item
-      We have $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{N}$.
+      We have $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{M}$.
 
@@
 \end{definition}
+
+\begin{thm}
+  Every canonical operator
+  $\IsCanOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$ is
+  extensional.
+\end{thm}
+
+\begin{proof}
+  Fix
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}%
+  and $\Member{\vec{E},\vec{F}}{\BTms^{[\vec{v}]}(\Upsilon')}$ and
+  $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$.
+  %
+  By hypothesis (1) from Definition~\ref{def:extensionality} above, we have
+  $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{M}$;
+  because $M$ is a closed term, by the idempotency of evaluation we have
+  $\Match{M}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}$.
+
+  It suffices, then, to show that
+  $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$.
+  %
+  Backward chaining through the \textit{app}-rule of $\preccurlyeq^\star$, we must
+  come up with a certain $\vec{G}$ such that the following subgoals obtain:
+  \begin{enumerate}
+    \item $\SBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{G}}$
+    \item $\SBinRel{\preccurlyeq_\tau}{\Upsilon'}{\App{\Rename{\rho}{\vartheta}}{\vec{G}}}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
+  \end{enumerate}
+
+  Setting $\Define{\vec{G}}{\vec{F}}$, we discharge the first subgoal through
+  hypothesis (2); the second hypothesis holds because of the reflexivity of
+  approximation (Theorem~\ref{thm:approx-refl}).
+
+\end{proof}
 

### 2015-12-21 paper.tex
fix typo
typo
improve wording

@@
 The union $\Refines{\Upsilon}{\bigcup_{(x:\phi)}\psi}{\tau}$ of a family of
-refinements is defined as by the following rule:
+refinements is defined by the following rule:
 \[
@@
   The proof proceeds by coinduction, using the relation
-  $\IMode{M,N}\mapsto\OMode{M = N}$ as our motive; it, then, suffices to show
+  $\IMode{M,N}\mapsto\OMode{M = N}$ as our motive; then, it suffices to show
   that for all $M$, $\SBinRel{[=]}{\Upsilon}{M}{M}$. Expanding the definition
@@
   $\IsEq{\vec{E}}{\vec{E}}$; but this is always the case, because of the
-  functionality of the evaluation relation.
+  functionality/determinacy of the evaluation relation.
 \end{proof}
@@
   \infer[?]{
+    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{M}{N}
+  }{
     \OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{M}{N}
-  }{
-    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{M}{N}
   }

### 2015-12-21 paper.tex
Switch to large power object, large subobject classifier

@@
 
-% power object
 \newcommand\Pow[1]{\wp\left(#1\right)}
@@
 expression, being responsible for the management of variable binding and
-substitution, among other things. On the other hand, mathematical objects find
+substitution. On the other hand, mathematical objects find
 \emph{meaning} in their extension within the behavioral type theory, and are
@@
 rules.}
+
+\section{Generalized Predicative Mathematical Foundations}
+
+Before we continue, it makes sense to remark briefly on the strength of the
+mathematical foundations that we are assuming for the development of our
+framework. As in Martin-L\"of's meaning-theoretic judgmental method, our
+apparatus is meant to be understood in a constructive and predicative
+metatheory; therefore, by $\Sets$, we do not mean the classical category of
+sets, but rather the intuitionistic one, which has lacks the powerset
+operation.
+
+Therefore, in our setting, $\Sets$ is not an elementary topos, and nor is any
+category of $\Sets$-valued (pre)sheaves. However, for a category of classes
+$\Classes$, we can express a power class operation
+$\Define{\Pow{X}}{\MkSet{Y\mid Y\subseteq X, Y\in\Sets}}$ which gives the
+proper class of sub\emph{sets} of a class $X$ \cite{gambino:2002}.
+
+A category of $\Classes$-valued presheaves then is a $\Pi W M$-pretopos (a
+predicative topos with dependent function objects, wellfounded trees, and
+non-wellfounded trees); such a pretopos also admits a large subobject
+classifier $\Omega(c)$, defined as the class of small sieves on
+$c$,\footnote{A sieve $S$ on $c$ is a subobject of the Yoneda
+embedding $\Yoneda{c}$, and $S$ is \emph{small} if $S(d)$ is a set for every
+stage $d$.} whence we can form the large power object
+$\Define{\Pow{X}}{\Omega^X}$ for any object $X$.
+
+In practice, the shift to a predicative topos with a large power object was a
+mere technical nuisance, as all the constructions we needed to perform carried
+over perfectly to this setting. For the remainder of this paper, we assume the
+semantic universe $\Define{\SemU}{\Classes^\SCtx}$, the category of
+copresheaves on symbol contexts. Every judgment $\mathcal{J}$ which we define
+in this paper should be understood as such a presheaf, which is at each stage
+$\Upsilon$ the class of derivations of $\mathcal{J}$ using symbols in
+$\Upsilon$; we will also make use of \emph{internal} forms of judgment, which
+are sections of $\Pow{X}$ for some object $X$.
 
@@
   the pointwise style in which we have expressed it. Considered internally to
-  the copresheaf topos $\Sets^\SCtx$, a refinement is merely a section of the
+  the semantic universe $\SemU$, a refinement is merely a section of the
   object of $\Exprs_\tau$-PERs $\PERs{\Exprs_\tau}$, defined
-  internally as a subobject of $\Omega^{\Exprs_\tau\times\Exprs_\tau}$:
+  internally as a subobject of $\Pow{\Exprs_\tau\times\Exprs_\tau}$:
   \[
@@
     }{
-      \{ \phi : \Omega^{X\times X}
+      \{ \phi : \Pow{X\times X}
          \mid
@@
   $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of explaining such
-  judgments as objects in the topos is that we do not need to deal with
+  judgments as copresheaves is that we do not need to deal with
   the complexities of quantifying over renamings, since this is implicit in the
@@
 quantification, in the sense that the terms themselves bear no trace of the
-quantification, but are guaranteed in the metalogic to be related in all fibres
+quantification, but are guaranteed in the metalogic to be related in all fibers
 of $\psi$.
 
-In practice, the family intersection is perhaps the most-used type constructor
-in Nuprl developments, as it facilitates the simultaneous specification and
-implementation of programs in type theory, without causing details of the
-program's verification to leak into the extracted algorithm. Additionally, the
-family intersection can be used to implement coinductive types.
+In practice, the family intersection is perhaps the most ubiquitous type
+constructor in Nuprl developments, as it facilitates the simultaneous
+specification and implementation of programs in type theory, without causing
+details of the program's verification to leak into the extracted algorithm.
+Additionally, the family intersection can be used to implement coinductive
+types.
 
@@
 The set comprehension $\MkSet{x:\phi\mid\psi}$ contains just those elements of
-$\phi$ such that their fibres of $\psi$ are inhabited; this can be seen as an
+$\phi$ such that their fibers of $\psi$ are inhabited; this can be seen as an
 ``implicit'' version of existential quantification, and can be used to
@@
 $\EvalN{\Upsilon}{\tau}{n}{M}{N}$.
-
-\begin{remark}
-  In any topos $\mathcal{E}$, for an object $X$ we can define the object of
-  relations on $X$ as the exponential $\Define{\Pow{X}}{\Omega^X}$. Then, the
-  data of such a relation is contained in a global section
-  $\Of{R}{\Hom{\mathcal{E}}{\mathbb{1}}{\Pow{X}}}$.
-\end{remark}
 
@@
 
-Stated internally as an object in the topos, the meaning of parametric typehood for
+Stated internally as an object in $\SemU$, the meaning of parametric typehood for
 $\phi$ is perhaps a bit more clear:

### 2015-12-21 paper.tex
update a bunch of instances of Sets^Sctx

@@
     }{
-      \Hom{\Sets^\SCtx}{
+      \Hom{\SemU}{
         \Coyoneda{\Upsilon}\times A
@@
 \begin{notation*}
-  For a family of copresheaves $\HypJ{\Of{X_i}{\Sets^\SCtx}}{\Member{i}{I}}$, we will
+  For a family of copresheaves $\HypJ{\Of{X_i}{\SemU}}{\Member{i}{I}}$, we will
   use $X^{[\vec{i}]}$ as a shorthand for the following iterated product, as
@@
 Fix a sort-indexed family of binary relations on closed expressions
-$\Of{R_\tau}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$; we will
+$\Of{R_\tau}{\Hom{\SemU}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$; we will
 also write the relation as a judgment scheme
@@
 family of valenced-indexed relations
-$\Of{R_v}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for
+$\Of{R_v}{\Hom{\SemU}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for
 $\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
@@
   call \emph{computational approximation},
-  $\Of{\preccurlyeq_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
+  $\Of{\preccurlyeq_\tau}{\Hom{\SemU}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
   written pointwise as $\SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}$.
@@
 A sort-indexed family of binary relations
-$\HypJ{\Of{R_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}}{\Member{\tau}{\Sorts}}$
+$\HypJ{\Of{R_\tau}{\Hom{\SemU}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}}{\Member{\tau}{\Sorts}}$
 is a \emph{congruence} when for all operators
@@
 partial equivalence relation on values,
-$\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.  It
+$\Of{\phi}{\Hom{\SemU}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.  It
 is easy to turn such a relation into a computational type

### 2015-12-22 paper.tex
correct a typo
adjust section heading name
explain our purpose---meaning explanation, not just model

@@
 
-\section{Generalized Predicative Mathematical Foundations}
+\section{Generalized Predicative Foundations}
 
@@
 are sections of $\Pow{X}$ for some object $X$.
+
+Our purpose in fixing a semantic universe is emphatically \emph{not} to give a
+mathematical interpretation or model of type theory inside set theory, though
+it is possible to read our construction in this metamathematical light---however,
+the vocabulary of the pretopos $\SemU$ gives us a perspicuous language in which
+to propound what we consider to be an informal (pre-mathematical) meaning
+explanation in the sense of Heyting~\cite{heyting:1956} and
+Martin-L\"of~\cite{martin-lof:1979}.
+\footnote{
+  Both Gambino~\cite{gambino:2002} and Dybjer~\cite{dybjer:2012} give lucid
+  explanations of the distinction between \emph{meaning theory} and \emph{model theory}.
+}
 
@@
 \end{tikzcd}
-we can clearly define a new refinement $\Refines{\Upsilon'}{\Rename{\rho}{\rho}}{\tau}$.
+we can clearly define a new refinement $\Refines{\Upsilon'}{\Rename{\rho}{\phi}}{\tau}$.
 

### 2015-12-22 paper.tex
wording

--- 2015-12-21 paper.tex
@@
 are sections of $\Pow{X}$ for some object $X$.
+
+Our purpose in fixing a semantic universe is emphatically \emph{not} to give a
+mathematical interpretation or model of type theory inside set theory, though
+it is possible to read our construction in this metamathematical light---however,
+the vocabulary of the pretopos $\SemU$ gives us a perspicuous notation in which
+to propound what we consider to be an informal (pre-mathematical) meaning
+explanation in the sense of Heyting~\cite{heyting:1956} and
+Martin-L\"of~\cite{martin-lof:1979}.
+\footnote{
+  Both Gambino~\cite{gambino:2002} and Dybjer~\cite{dybjer:2012} give lucid
+  explanations of the distinction between \emph{meaning theory} and \emph{model theory}.
+}
 

### 2015-12-22 paper.tex
more wording

--- 2015-12-21 paper.tex
@@
 copresheaves on symbol contexts. Every judgment $\mathcal{J}$ which we define
-in this paper should be understood as such a presheaf, which is at each stage
+in this paper should be understood as such a copresheaf, which is at each stage
 $\Upsilon$ the class of derivations of $\mathcal{J}$ using symbols in
@@
 are sections of $\Pow{X}$ for some object $X$.
+
+Our purpose in fixing a semantic universe is emphatically \emph{not} to give an
+interpretation or model of type theory inside set theory, though it is possible
+to read our construction in this metamathematical light---however, the
+vocabulary of the pretopos $\SemU$ gives us a perspicuous notation in which to
+propound what we consider to be an informal (pre-mathematical) meaning
+explanation in the sense of Heyting~\cite{heyting:1956} and
+Martin-L\"of~\cite{martin-lof:1979}.
+\footnote{
+  Both Gambino~\cite{gambino:2002} and Dybjer~\cite{dybjer:2012} give lucid
+  explanations of the distinction between \emph{meaning theory} and \emph{model theory}.
+}
 

### 2015-12-22 paper.tex
typo

@@
 The dependent intersection, first introduced by Kopylov in \cite{kopylov:2000},
-are a generalization of the binary intersection to the dependent case, not to
+is a generalization of the binary intersection to the dependent case, not to
 be confused with the intersection of a family of sets. The dependent

### 2015-12-22 paper.tex
notation updates

@@
 
+\newcommand\PFunEqNil{\vDash^=_{\mathtt{nil}}}
+\newcommand\PFunEqSnoc{\vDash^=_{\mathtt{snoc}}}
+
 Functional equality $\FRMemEq{\Upsilon}{\Psi}{N_0}{N_1}{\phi}$ presupposes
@@
 \[
-  \infer{
+  \infer[\PFunEqNil]{
     \FRMemEq{\Upsilon}{\cdot}{N_0}{N_1}{\phi}
@@
   \[
-    \infer{
+    \infer[\PFunEqSnoc]{
       \FRMemEq{\Upsilon}{\Psi,x:\psi}{N_0}{N_1}{\phi}
@@
 \[
-  \infer{
+  \infer={
     \RMemEq{\Upsilon'}{M}{N}{\phi\cap\psi}
@@
 \[
-  \infer{
+  \infer={
     \RMemEq{\Upsilon'}{M}{N}{
@@
 
-\newcommand\SBinRel[4]{\IMode{#2}\parallel\IMode{#3}\mathrel{#1}\IMode{#4}}
-\newcommand\OSBinRel[5]{\IMode{#2}\parallel\IMode{#3}\vDash\IMode{#4}\mathrel{#1}\IMode{#5}}
+\newcommand\PBinRel[4]{\IMode{#2}\parallel\IMode{#3}\mathrel{#1}\IMode{#4}}
+\newcommand\HPBinRel[5]{\IMode{#2}\parallel\IMode{#3}\vDash\IMode{#4}\mathrel{#1}\IMode{#5}}
 
@@
 also write the relation as a judgment scheme
-$\SBinRel{R_\tau}{\Upsilon}{M}{N}$. First of all, we can lift such a relation
+$\PBinRel{R_\tau}{\Upsilon}{M}{N}$. First of all, we can lift such a relation
 to operate on open terms by quantifying over substitutions, as follows:
@@
   \infer{
-    \OSBinRel{R_\tau}{\Upsilon}{\vec{x}:\vec{\sigma}}{M}{N}
+    \HPBinRel{R_\tau}{\Upsilon}{\vec{x}:\vec{\sigma}}{M}{N}
   }{
@@
     \quad
-    \SBinRel{R_\tau}{\Upsilon'}{
+    \PBinRel{R_\tau}{\Upsilon'}{
       \Subst{\vec{X}}{\vec{x}}{\Rename{\rho}{M}}
@@
   \infer{
-    \SBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
+    \PBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
       \MkBTm{\vec{u}}{\vec{x}}{M}
@@
       \forall\IMode{\vec{z}}\mathrel{\#}\IMode{\vec{x}\cup\vec{y}}.\\
-      \OSBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{\vec{z}:\vec{\tau}}{
+      \HPBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{\vec{z}:\vec{\tau}}{
         \Subst{\vec{z}}{\vec{x}}{
@@
   \infer{
-    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
+    \PBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
   }{
     \forall\Member{(E,F,v)}{(\vec{E},\vec{F},\vec{v})}.\
-    \SBinRel{R_v}{\Upsilon}{E}{F}
+    \PBinRel{R_v}{\Upsilon}{E}{F}
   }
@@
 which respects a single ``layer'' of computation. We will say
-$\SBinRel{[R_\tau]}{\Upsilon}{M}{N}$ when, supposing
+$\PBinRel{[R_\tau]}{\Upsilon}{M}{N}$ when, supposing
 $\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}}$ such that
@@
 , we have $\Eval{\Upsilon'}{\tau}{\Rename{\rho}{N}}{\App{\Rename{\rho}{\vartheta}}{\Rename{\rho}{\Vec{F}}}}$ and
-$\SBinRel{R_{\vec{v}}}{\Upsilon'}{\Rename{\rho}{\vec{E}}}{\Rename{\rho}{\vec{F}}}$.
+$\PBinRel{R_{\vec{v}}}{\Upsilon'}{\Rename{\rho}{\vec{E}}}{\Rename{\rho}{\vec{F}}}$.
 
@@
   $\Of{\preccurlyeq_\tau}{\Hom{\SemU}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
-  written pointwise as $\SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}$.
+  written pointwise as $\PBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}$.
 \end{definition}
@@
 \begin{thm}[Reflexivity]\label{thm:approx-refl}
-  For all $\Member{M}{\Exprs_\tau(\Upsilon)}$, we have $\SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{M}$.
+  For all $\Member{M}{\Exprs_\tau(\Upsilon)}$, we have $\PBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{M}$.
 \end{thm}
@@
   $\IMode{M,N}\mapsto\OMode{M = N}$ as our motive; then, it suffices to show
-  that for all $M$, $\SBinRel{[=]}{\Upsilon}{M}{M}$. Expanding the definition
+  that for all $M$, $\PBinRel{[=]}{\Upsilon}{M}{M}$. Expanding the definition
   of $[-]$, this means that we have to show that when
@@
     \infer{
-      \SBinRel{\sim_\tau}{\Upsilon}{M}{N}
-    }{
-      \SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N} &
-      \SBinRel{\preccurlyeq_\tau}{\Upsilon}{N}{M}
+      \PBinRel{\sim_\tau}{\Upsilon}{M}{N}
+    }{
+      \PBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N} &
+      \PBinRel{\preccurlyeq_\tau}{\Upsilon}{N}{M}
     }
@@
   \infer{
-    \SBinRel{R_\tau}{\Upsilon}{\vartheta(\vec{E})}{\vartheta(\vec{F})}
-  }{
-    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
+    \PBinRel{R_\tau}{\Upsilon}{\vartheta(\vec{E})}{\vartheta(\vec{F})}
+  }{
+    \PBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
   }
@@
   \infer[\textit{var}]{
-    \OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{x}{y}
-  }{
-    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{x}{y}
+    \HPBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{x}{y}
+  }{
+    \HPBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{x}{y}
   }
@@
   \infer[\textit{app}]{
-    \OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{
+    \HPBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{
       \App{\vartheta}{\vec{E}}
@@
     \exists\IMode{\vec{F}}. &
-    \OSBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon}{\Gamma}{
+    \HPBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon}{\Gamma}{
       \vec{E}
@@
     } &
-    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{
+    \HPBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{
       \App{\vartheta}{\vec{F}}
@@
 
-We will write $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{M}{N}$ for
-$\OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\cdot}{M}{N}$.  By definition,
+We will write $\PBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{M}{N}$ for
+$\HPBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\cdot}{M}{N}$.  By definition,
 $\preccurlyeq^\star_\tau$ is a congruence; to show that $\sim_\tau$ is a
@@
   \infer[?]{
-    \OSBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{M}{N}
-  }{
-    \OSBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{M}{N}
+    \HPBinRel{\preccurlyeq_\tau}{\Upsilon}{\Gamma}{M}{N}
+  }{
+    \HPBinRel{\preccurlyeq^\star_\tau}{\Upsilon}{\Gamma}{M}{N}
   }
@@
   $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$, if we can conclude
-  $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{M}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
+  $\PBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{M}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
   from the following conditions:
@@
     \item
-      We have $\SBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{F}}$.
+      We have $\PBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{F}}$.
 
@@
       from $\EvalN{\Upsilon''}{\sigma}{k'}{X}{X'}$ for $\IMode{k'}<\IMode{k}$ and
-      $\SBinRel{\preccurlyeq^\star_\sigma}{\Upsilon''}{X}{Y}$, we can conclude
-      $\SBinRel{\preccurlyeq^\star_\sigma}{\Upsilon''}{X'}{Y}$.
+      $\PBinRel{\preccurlyeq^\star_\sigma}{\Upsilon''}{X}{Y}$, we can conclude
+      $\PBinRel{\preccurlyeq^\star_\sigma}{\Upsilon''}{X'}{Y}$.
   \end{enumerate}
@@
   It suffices, then, to show that
-  $\SBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$.
+  $\PBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$.
   %
@@
   \begin{enumerate}
-    \item $\SBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{G}}$
-    \item $\SBinRel{\preccurlyeq_\tau}{\Upsilon'}{\App{\Rename{\rho}{\vartheta}}{\vec{G}}}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
+    \item $\PBinRel{\preccurlyeq^\star_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{G}}$
+    \item $\PBinRel{\preccurlyeq_\tau}{\Upsilon'}{\App{\Rename{\rho}{\vartheta}}{\vec{G}}}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
   \end{enumerate}
@@
   , and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
-  $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
+  $\PBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
   can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
@@
   $\Define{\underline{\phi}}{\VToCType(\iota(\phi))}$ be called $\phi$'s
-  \emph{value closure}, which we will use shortly in order to give a formal definition
-  to the notion of value types.
+  \emph{value closure}, which we will use shortly in order to give a formal
+  definition to the notion of value types.
 \end{definition}
@@
   We can give a simple characterization of Martin-L\"of's \emph{value types} in
-  terms of our more general computational types. For any type $\IsType{\Upsilon}{\phi}{\tau}$,
-  we will say $\IsValueType{\Upsilon}{\phi}{\tau}$ in case $\phi$ is equal to
-  its value closure, i.e.\ $\EqRefines{\Upsilon}{\phi}{\underline{\phi}}{\tau}$.
+  terms of our more general computational types. For any type
+  $\IsType{\Upsilon}{\phi}{\tau}$, we will say
+  $\IsValueType{\Upsilon}{\phi}{\tau}$ in case $\phi$ is equal to its value
+  closure, i.e.\ $\EqRefines{\Upsilon}{\phi}{\underline{\phi}}{\tau}$.
 \end{definition}

### 2015-12-22 paper.tex
fix typo

@@
 metatheory; therefore, by $\Sets$, we do not mean the classical category of
-sets, but rather the intuitionistic one, which has lacks the powerset
+sets, but rather the intuitionistic one, which lacks the powerset
 operation.

### 2015-12-22 paper.tex
Add dependent function types

@@
 
+\paragraph{Functional Typehood}
+\newcommand\PTyFamNil{\vDash^\textit{type}_{\mathtt{nil}}}
+\newcommand\PTyFamSnoc{\vDash^\textit{type}_{\mathtt{snoc}}}
+
+Let us define functional typehood
+$\EqTypeFam{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$, presupposing
+$\FEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$. As usual, we
+give our meaning explanation by induction on the evidence for
+the validity of the refined context $\Psi$.
+
+\paragraph{Case} $\DefRefinesCtxNil{\Upsilon}$.
+\[
+  \infer[\PTyFamNil]{
+    \EqTypeFam{\Upsilon}{\cdot}{\phi}{\psi}{\tau}
+  }{
+    \IsType{\Upsilon}{\phi}{\tau}
+  }
+\]
+Note that this case is not trivial, because it is only defined in
+case the presupposition $\FEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$
+obtains. It suffices to require only that $\phi$ be a type, since because
+typehood respects equality of refinement.
+
+\paragraph{Case} $\DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
+
+\begin{quote}
+  To know $\EqTypeFam{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}$ is to know,
+  for any renaming
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}
+  and closed terms $\Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}$, that from
+  $\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}$, you can conclude
+  $\EqTypeFam{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{\phi}}}{\Subst{M_1}{x}{\Rename{\rho}{\psi}}}{\tau}$.
+  In other words:
+  %
+  \[
+    \infer[\PTyFamSnoc]{
+      \EqTypeFam{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
+    }{
+      \begin{array}{l}
+        \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}
+        .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
+        \left(\FRMemEq{\Upsilon'}{\Rename{\rho}{\Psi}}{M_0}{M_1}{\Rename{\rho}{\chi}}\right)\\
+        \quad\implies
+        \left(\EqTypeFam{\Upsilon'}{\Rename{\rho}{\Psi}}{\Subst{M_0}{x}{\Rename{\rho}{\phi}}}{\Subst{M_1}{x}{\Rename{\rho}{\psi}}}{\tau}\right)
+      \end{array}
+    }
+  \]
+\end{quote}
+
+As with refinements, we will write $\IsTypeFam{\Upsilon}{\Psi}{\phi}{\tau}$
+to mean $\EqTypeFam{\Upsilon}{\Psi}{\phi}{\phi}{\tau}$.
+
+\subsection{Example: Dependent Function Types}
+\newcommand\SortArr[2]{#1\rightharpoonup #2}
+\newcommand\OpLambda[2]{\mathtt{lam}_{#1}^{#2}}
+\newcommand\OpAp[2]{\mathtt{ap}_{#1}^{#2}}
+
+Fix an extensional lcs such that its signature contains the following:
+\[
+  \infer{
+    \IsSort{\SortArr{\sigma}{\tau}}
+  }{
+    \IsSort{\sigma} &
+    \IsSort{\tau}
+  }
+  \qquad
+  \infer{
+    \IsOperator{\Upsilon}{
+      \OpAp{\sigma}{\tau}
+    }{
+      \MkArity{
+        \MkValence{}{}{\SortArr{\sigma}{\tau}};
+        \MkValence{}{}{\sigma}
+      }{\tau}
+    }
+  }{
+    \IsSort{\sigma} &
+    \IsSort{\tau}
+  }
+\]
+
+The sort $\SortArr{\sigma}{\tau}$ is intended to classify the (syntax of)
+partial operations, and $\OpAp{\sigma}{\tau}$ is the application operator. Now,
+we can define a refinement on this sort for dependent functions, and then prove
+that it is a type. We intend to define a refinement $\prod_{(x:\phi)}\psi$
+such that the following rule holds:
+\[
+  \infer{
+    \Refines{\Upsilon}{\prod_{(x:\phi)}\psi}{\SortArr{\sigma}{\tau}}
+  }{
+    \Refines{\Upsilon}{\phi}{\sigma} &
+    \FRefines{\Upsilon}{x:\phi}{\psi}{\tau}
+  }
+\]
+
+For any renaming
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}%
+, the meaning of $\prod_{(x:\phi)}\psi$ at
+$\Member{M, N}{\Exprs_{\SortArr{\sigma}{\tau}}(\Upsilon')}$ is given by
+the following rule:
+\[
+  \infer={
+    \RMemEq{\Upsilon'}{M}{N}{
+      \prod_{(x:\phi)}\psi
+    }
+  }{
+    \FRMemEq{\Upsilon'}{x:\phi}{
+      \App{\OpAp{\sigma}{\tau}}{M; x}
+    }{
+      \App{\OpAp{\sigma}{\tau}}{N; x}
+    }{
+      \psi
+    }
+  }
+\]
+
+Note that we have not required an introduction form such as
+$\OpLambda{\sigma}{\tau}$, and have instead opted for a negative
+definition of the dependent function refinement. This is strictly more
+open-ended than the positive definition, because it does not fix in
+advance a syntax for the canonical formation of functions.
+
+\begin{thm}
+  The dependent function refinement
+  $\Refines{\Upsilon}{\prod_{(x:\phi)}\psi}{\SortArr{\sigma}{\tau}}$ is a
+  type assuming $\IsType{\Upsilon}{\phi}{\sigma}$ and
+  $\IsTypeFam{\Upsilon}{x:\phi}{\psi}{\tau}$.
+\end{thm}
+\begin{proof}
+  Fix
+    \begin{tikzcd}[cramped]
+      \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+    \end{tikzcd}%
+  , and $\Member{L,M,N}{\Exprs_{\SortArr{\sigma}{\tau}}(\Upsilon')}$ such that
+  \begin{enumerate}
+    \item $\PBinRel{\sim_{\SortArr{\sigma}{\tau}}}{\Upsilon'}{L}{M}$
+    \item $\FRMemEq{\Upsilon'}{x:\Rename{\rho}{\phi}}{\App{\OpAp{\sigma}{\tau}}{M;x}}{\App{\OpAp{\sigma}{\tau}}{N;x}}{\Rename{\rho}{\psi}}$
+  \end{enumerate}
+%
+  It suffices to show that
+  $\FRMemEq{\Upsilon'}{x:\Rename{\rho}{\phi}}{\App{\OpAp{\sigma}{\tau}}{L;x}}{\App{\OpAp{\sigma}{\tau}}{N;x}}{\Rename{\rho}{\psi}}$.
+%
+  Now fix
+    \begin{tikzcd}[cramped]
+      \IMode{\Upsilon'}\arrow[r,hook,"\IMode{\rho'}"] &\IMode{\Upsilon''}
+    \end{tikzcd}%
+  , and $\Member{X,Y}{\Exprs_\sigma(\Upsilon'')}$ such that
+  $\RMemEq{\Upsilon''}{X}{Y}{\Rename{(\rho'\circ\rho)}{\phi}}$.
+  Then, we have to show that
+  $\RMemEq{\Upsilon''}{%
+    \App{\OpAp{\sigma}{\tau}}{
+      \Rename{\rho'}{L};
+      X
+    }
+  }{
+    \App{\OpAp{\sigma}{\tau}}{
+      \Rename{\rho'}{N};
+      Y
+    }
+  }{\Subst{X}{x}{\Rename{(\rho'\circ\rho)}{\psi}}}$.
+
+  Because $\psi$ is a functional family of types,
+  $\Subst{X}{x}{\Rename{(\rho'\circ\rho)}{\psi}}$ must respect
+  computational equivalence, so by drawing hypothesis (1) along $\rho'$, we
+  can rewrite hypothesis (2) into the goal.
+\end{proof}
+
 \ifdraft{}{

### 2015-12-23 paper.tex
Define the type of atoms

@@
 
-\subsection{Example: Dependent Function Types}
+\subsection{Examples of refinements and their typehood}
+
+We will demonstrate how to define certain refinements, and prove that they are
+types.
+
+\subsubsection{The type of atoms}
+\newcommand\SortAtom{\mathtt{atom}}
+\newcommand\OpToken[1]{\mathtt{tok}[#1]}
+\newcommand\OpMatch[1]{\mathtt{match}_{#1}}
+
+Fix an extensional lcs whose signature contains the following:
+\[
+  \infer{
+    \IsSort{\SortAtom}
+  }{
+  }
+\]
+\[
+  \infer{
+    \IsCanOperator{\Upsilon,a:\SortAtom}{
+      \OpToken{a}
+    }{
+      \MkArity{}{\SortAtom}
+    }
+  }{
+  }
+  \qquad
+  \infer{
+    \IsOperator{\Upsilon}{
+      \OpMatch{\tau}
+    }{
+      \MkArity{
+        \MkValence{}{}{\SortAtom};
+        \MkValence{}{}{\tau};
+        \MkValence{}{}{\tau}
+      }{\tau}
+    }
+  }{
+    \IsSort{\tau}
+  }
+\]
+
+Then, we can define a refinement $\GloballyRefines{\mathbb{A}_0}{\SortAtom}$ as
+follows:
+\[
+  \infer{
+    \RMemEq{\Upsilon}{
+      \App{\OpToken{a}}{}
+    }{
+      \App{\OpToken{a}}{}
+    }{\mathbb{A}_0}
+  }{
+    \IMode{\Upsilon}\ni\IMode{a}:\IMode{\SortAtom}
+  }
+\]
+
+However, it is not the case that $\mathbb{A}_0$ is a type, because it doesn't
+respect computational equivalence. This is easily resolved by defining the type
+of atoms as the computational extension of $\mathbb{A}_0$,
+\[
+  \Define{
+    \mathbb{A}
+  }{
+    \VToCType(\mathbb{A}_0)
+  }
+\]
+
+\begin{thm}
+  For any $\Upsilon$, we have $\IsType{\Upsilon}{\mathbb{A}}{\SortAtom}$.
+\end{thm}
+\begin{proof}
+  Trivial, because the computational extension of any value PER is a type.
+\end{proof}
+
+\subsubsection{Dependent function types}
 \newcommand\SortArr[2]{#1\rightharpoonup #2}

### 2015-12-23 paper.tex
no need to even show the Match operator
Add Base{τ} type

@@
 
+\subsubsection{The type of closed expressions}
+
+The \emph{discrete} refinement on a sort (which equates each closed term only
+with itself, up to alpha equivalence) is not a type in general, because it may
+not respect computational equivalence. However, we can form the next best
+thing, which is called the \emph{base} refinement, which identifies closed
+expressions of a sort up to computational equivalence.
+
+Let $\GloballyRefines{\TyBase{\tau}}{\tau}$ for any sort $\tau$ be defined as
+follows:
+\[
+  \infer{
+    \RMemEq{\Upsilon}{M}{N}{\TyBase{\tau}}
+  }{
+    \PBinRel{\sim_\tau}{\Upsilon}{M}{N}
+  }
+\]
+
+Then, we clearly have $\IsType{\Upsilon}{\TyBase{\tau}}{\tau}$, because it respects
+computational equivalence by definition.
+
 \subsubsection{The type of atoms}
@@
   }
-\]
-\[
+  \qquad
   \infer{
@@
   }{
-  }
-  \qquad
-  \infer{
-    \IsOperator{\Upsilon}{
-      \OpMatch{\tau}
-    }{
-      \MkArity{
-        \MkValence{}{}{\SortAtom};
-        \MkValence{}{}{\tau};
-        \MkValence{}{}{\tau}
-      }{\tau}
-    }
-  }{
-    \IsSort{\tau}
   }

### 2015-12-23 paper.tex
capitalize Yoneda

@@
   definition of the exponential of presheaves, where
-  $\Define{\Coyoneda{\Upsilon}}{\Upsilon\inj-}$ is the co-yoneda embedding:
+  $\Define{\Coyoneda{\Upsilon}}{\Upsilon\inj-}$ is the co-Yoneda embedding:
   \begin{align*}

### 2015-12-24 paper.tex
Compare with Allen's supervaluation semantics (#9)

@@
 We will demonstrate how to define certain refinements, and prove that they are
-types.
+types. The refinements below will differ from the ones defined in Section~
 
@@
 
+\section{Related Work}
+
+In \cite{allen:2006}, Stuart Allen gave a \emph{supervaluation} semantics for
+Nuprl sequents involving atoms (symbols) by quantifying over plenty of choices
+for the interpretation of atoms---possible interpretations including strings,
+finite sets, etc. One benefit of Allen's indirect treatment is that the logical
+theory remains profoundly open-ended, in the sense that it is neither the case
+that the naturals may be injected into the atoms, nor is it the case that they
+cannot.
+
+\newcommand\AtomModel{\mathfrak{A}}
+\newcommand\NuprlSeq[2]{\IMode{#1}\gg\IMode{#2}}
+\newcommand\NuprlInterp[2]{\IMode{#1}\models\IMode{#2}}
+\newcommand\NuprlInterpSeq[3]{\IMode{#1}\models\NuprlSeq{#2}{#3}}
+\newcommand\Strings{\mathbf{String}}
+
+Unlike in our treatment, where the judgments of type theory are intensional
+with respect to symbol contexts, Allen takes a more model-theoretic approach
+and parameterizes the explanation of type theory over a model $\AtomModel$ for
+the abstract atoms.
+
+A Nuprl sequent $\Define{\mathfrak{s}}{H\gg P}$ means ``$P$ is true under
+hypotheses $H$'', and it may contain (abstract) atoms in its syntax; we write
+$\FS{\mathfrak{s}}$ for the set of free atoms of the sequent $\mathfrak{s}$.
+Then, a model $\AtomModel$ consists in a carrier set $\Dom{\AtomModel}$ which
+shall interpret the atoms, and a valuation
+$\NuprlInterp{\AtomModel}{\mathfrak{s}}$ for any Nuprl sequent $\mathfrak{s}$
+which takes its atoms from $\Dom{\AtomModel}$.  A model is called
+\emph{discrete} in case its carrier set has decidable equality.  For a Nuprl
+sequent $\mathfrak{s}$ with atoms in $\AtomModel$ and an injective function
+$\Of{\rho}{\Dom{\AtomModel}\inj\Dom{\AtomModel'}}$, let
+$\Rename{\rho}{\mathfrak{s}}$ be the sequent with atoms in $\Dom{\AtomModel'}$
+got by replacing each atom $a$ with $\rho(a)$.
+
+Then, a supervaluation is given for Nuprl sequents which take atoms in
+$\Strings$, written $\NuprlInterp{*}{\mathfrak{s}}$, which holds if there
+exists a $\Member{k}{\Naturals}$ such that for every discrete model
+$\AtomModel$ having at least $k$ members, for every injection
+$\Of{\rho}{\FS{\mathfrak{s}} \inj \AtomModel}$, we have
+$\NuprlInterp{\AtomModel}{\Rename{\rho}{\mathfrak{s}}}$. Under the
+supervaluation, we can rule out certain observations about atoms which might
+hold in particular models.
+
+For instance, assuming a type $\Naturals$ of natural numbers and $\mathbb{A}$
+of atoms, with $\Naturals\inj\mathbb{A}$ the type of embeddings of the naturals
+into the atoms, we do have
+$\NuprlInterpSeq{\Strings}{\cdot}{\Naturals\inj\mathbb{A}}$, because there are
+numerous ways to embed the naturals into the strings; however, because the
+embedding fails in every finite model $\AtomModel$, we do not have
+$\NuprlInterpSeq{*}{\cdot}{\Naturals\inj\mathbb{A}}$.  Furthermore, we also do
+not have $\NuprlInterpSeq{*}{\iota:\Naturals\inj\mathbb{A}}{\bot}$, because this
+would fail in any infinite model.
+
+\paragraph{Comparison with the present work}
+
+Our explanation of atoms is not equivalent to the supervaluation semantics.
+First, let us introduce a notion analogous to the Nuprl sequent: when
+$\RefinesCtx{\cdot}{\Phi}{\Gamma}$ and $\FRefines{\cdot}{\Phi}{\phi}{\tau}$, we
+can form the sequent $\NuprlSeq{\Phi}{\phi}$, meaning that there exists an
+$\Member{\Upsilon}{\SCtx}$ and
+$\Member{M}{\Exprs_\tau(\Upsilon\parallel\Gamma)}$ such that
+$\FRMemEq{\Upsilon}{\Phi}{M}{M}{\phi}$. We will write
+$\models\NuprlSeq{\Phi}{\phi}$ when $\NuprlSeq{\Phi}{\phi}$ obtains.
+
+\begin{thm}
+  \label{thm:no-nat-injection}
+  As in the supervaluation semantics, we have
+  $\not\models\NuprlSeq{\cdot}{\Naturals\inj\mathbb{A}}$.
+\end{thm}
+\begin{proof}
+  Suppose $\models\NuprlSeq{\cdot}{\Naturals\inj\mathbb{A}}$. Then, there
+  exists an $\Member{\Upsilon}{\SCtx}$ and $\Member{M}{\Exprs_\tau(\Upsilon)}$
+  such that $\RMemEq{\Upsilon}{M}{M}{\Naturals\inj\mathbb{A}}$; but this
+  is impossible because it would require a means of embedding the naturals into
+  finite $\Dom{\Upsilon}$.
+\end{proof}
+
+\begin{thm}
+  Contrary to the supervaluation semantics, we do have
+  $\models\NuprlSeq{\iota:\Naturals\inj\mathbb{A}}{\bot}$.
+\end{thm}
+\begin{proof}
+  It suffices to show that for all $\Upsilon$, there is no $\iota$ such that
+  $\RMemEq{\Upsilon}{\iota}{\iota}{\Naturals\inj\mathbb{A}}$. However, by the
+  definition of $\SCtx$, all $\Member{\Upsilon}{\SCtx}$ are finite, so there
+  can be no injection of the naturals into $\Dom{\Upsilon}$.
+\end{proof}
+
+The crux of the problem is that all worlds (symbol contexts) are finite in our
+semantics, so it is possible to observe that it will never be possible to
+inject the naturals into the atoms. One way to make our system more open-ended
+is to replace the current semantic universe $\Define{\SemU}{\Classes^\SCtx}$
+with an extended universe $\Define{\SemU_\infty}{\Classes^{\SCtx_\infty}}$ with
+$\SCtx_\infty$ the set of possibly-infinite symbol contexts and injective,
+sort-preserving renamings. Externally, by Theorem~\ref{thm:no-nat-injection} we
+could still never exhibit an embedding of the naturals, but internally, we
+would be unable to rule it out.
+
 \ifdraft{}{

### 2015-12-24 paper.tex
Add critical comparison w/ Allen's semantics (#9)
@freebroccolo This is kind of interesting---I was thinking about the
internal observations that can be made in our semantics but which cannot
under Allen's, and now that I think of it, I believe that the
open-endedness of the supervaluation semantics relies on a condition
which is actually impossible if you take a Brouwerian view concerning
construction and data.

@@
 \newcommand\OpMatch[1]{\mathtt{match}_{#1}}
+\newcommand\Atoms{\mathbb{A}}
 
@@
 
-Then, we can define a refinement $\GloballyRefines{\mathbb{A}_0}{\SortAtom}$ as
+Then, we can define a refinement $\GloballyRefines{\Atoms_0}{\SortAtom}$ as
 follows:
@@
       \App{\OpToken{a}}{}
-    }{\mathbb{A}_0}
+    }{\Atoms_0}
   }{
@@
 
-However, it is not the case that $\mathbb{A}_0$ is a type, because it doesn't
+However, it is not the case that $\Atoms_0$ is a type, because it doesn't
 respect computational equivalence. This is easily resolved by defining the type
-of atoms as the computational extension of $\mathbb{A}_0$,
+of atoms as the computational extension of $\Atoms_0$,
 \[
   \Define{
-    \mathbb{A}
-  }{
-    \VToCType(\mathbb{A}_0)
+    \Atoms
+  }{
+    \VToCType(\Atoms_0)
   }
@@
 \begin{thm}
-  For any $\Upsilon$, we have $\IsType{\Upsilon}{\mathbb{A}}{\SortAtom}$.
+  For any $\Upsilon$, we have $\IsType{\Upsilon}{\Atoms}{\SortAtom}$.
 \end{thm}
@@
 
-For instance, assuming a type $\Naturals$ of natural numbers and $\mathbb{A}$
-of atoms, with $\Naturals\inj\mathbb{A}$ the type of embeddings of the naturals
+For instance, assuming a type $\Naturals$ of natural numbers and $\Atoms$
+of atoms, with $\Naturals\inj\Atoms$ the type of embeddings of the naturals
 into the atoms, we do have
-$\NuprlInterpSeq{\Strings}{\cdot}{\Naturals\inj\mathbb{A}}$, because there are
+$\NuprlInterpSeq{\Strings}{\cdot}{\Naturals\inj\Atoms}$, because there are
 numerous ways to embed the naturals into the strings; however, because the
 embedding fails in every finite model $\AtomModel$, we do not have
-$\NuprlInterpSeq{*}{\cdot}{\Naturals\inj\mathbb{A}}$.  Furthermore, we also do
-not have $\NuprlInterpSeq{*}{\iota:\Naturals\inj\mathbb{A}}{\bot}$, because this
+$\NuprlInterpSeq{*}{\cdot}{\Naturals\inj\Atoms}$.  Furthermore, we also do
+not have $\NuprlInterpSeq{*}{\iota:\Naturals\inj\Atoms}{\bot}$, because this
 would fail in any infinite model.
@@
   As in the supervaluation semantics, we have
-  $\not\models\NuprlSeq{\cdot}{\Naturals\inj\mathbb{A}}$.
+  $\not\models\NuprlSeq{\cdot}{\Naturals\inj\Atoms}$.
 \end{thm}
 \begin{proof}
-  Suppose $\models\NuprlSeq{\cdot}{\Naturals\inj\mathbb{A}}$. Then, there
+  Suppose $\models\NuprlSeq{\cdot}{\Naturals\inj\Atoms}$. Then, there
   exists an $\Member{\Upsilon}{\SCtx}$ and $\Member{M}{\Exprs_\tau(\Upsilon)}$
-  such that $\RMemEq{\Upsilon}{M}{M}{\Naturals\inj\mathbb{A}}$; but this
+  such that $\RMemEq{\Upsilon}{M}{M}{\Naturals\inj\Atoms}$; but this
   is impossible because it would require a means of embedding the naturals into
@@
 \begin{thm}
+  \label{thm:never-nat-injection}
   Contrary to the supervaluation semantics, we do have
-  $\models\NuprlSeq{\iota:\Naturals\inj\mathbb{A}}{\bot}$.
+  $\models\NuprlSeq{\iota:\Naturals\inj\Atoms}{\bot}$.
 \end{thm}
@@
   It suffices to show that for all $\Upsilon$, there is no $\iota$ such that
-  $\RMemEq{\Upsilon}{\iota}{\iota}{\Naturals\inj\mathbb{A}}$. However, by the
+  $\RMemEq{\Upsilon}{\iota}{\iota}{\Naturals\inj\Atoms}$. However, by the
   definition of $\SCtx$, all $\Member{\Upsilon}{\SCtx}$ are finite, so there
@@
 
-The crux of the problem is that all worlds (symbol contexts) are finite in our
-semantics, so it is possible to observe that it will never be possible to
-inject the naturals into the atoms. One way to make our system more open-ended
-is to replace the current semantic universe $\Define{\SemU}{\Classes^\SCtx}$
-with an extended universe $\Define{\SemU_\infty}{\Classes^{\SCtx_\infty}}$ with
-$\SCtx_\infty$ the set of possibly-infinite symbol contexts and injective,
-sort-preserving renamings. Externally, by Theorem~\ref{thm:no-nat-injection} we
-could still never exhibit an embedding of the naturals, but internally, we
-would be unable to rule it out.
+Another seemingly paradoxical observation is also possible:
+
+\begin{thm}
+  \label{thm:never-atom-injection}
+  The atoms may not be embedded into the naturals, i.e.\
+  $\models\NuprlSeq{\iota:\Atoms\inj\Naturals}{\bot}$.
+\end{thm}
+\begin{proof}
+
+  Fix a symbol context $\Upsilon$; suppose we have an $\Member{\iota}{\Exprs(\Upsilon)}$ such that
+  $\RMemEq{\Upsilon}{\iota}{\iota}{\Atoms\inj\Naturals}$. Then, by inversion we
+  have
+  $\FRMemEq{\Upsilon}{x:\Atoms}{\App{\OpAp{}{}}{\iota;x}}{\App{\OpAp{}{}}{\iota;x}}{\Naturals}$;
+  in other words, for all
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}
+  and $\Member{a}{\Dom{\Upsilon}}$, we have
+  $\RMemEq{\Upsilon'}{\App{\OpAp{}{}}{\iota;a}}{\App{\OpAp{}{}}{\iota;a}}{\Naturals}$
+  such that $\iota$ is injective.
+  But the only way for $\iota$ to assign a different index to each atom is for
+  it to test each atom in $\Dom{\Upsilon'}$ separately---but $\iota$ may only
+  mention the atoms in $\Dom{\Upsilon}$.
+\end{proof}
+
+\paragraph{Should the above be worrisome?}
+
+At the very least, the fact that neither the naturals nor the atoms may be
+embedded into the other commits us to a truly intuitionistic ontology, as
+opposed to merely constructive one. In particular, our framework is
+anti-classical in at least a na\"ive sense, because the Axiom of Choice puts
+each set in bijection with a cardinal, whence it is impossible for neither of
+two sets to embed in the other. However, it is important to note that as far as
+the propositions-as-types principle is concerned, Computational Type Theory is
+already anti-classical, as the undecidability of the halting program can be
+observed internally.
+
+In fact, we view the simultaneous validity of
+Theorems~\ref{thm:never-nat-injection}~and~\ref{thm:never-atom-injection} as
+totally desirable, because we adhere to a strictly Brouwerian conception of
+data. The proper way to view the type of atoms is as an ongoing act of free
+generation---and it is a basic fact concerning the activity of an idealized
+mathematician that at each stage, only a finite number of constructions shall
+have been effected so far.
+
+On the other hand, (the analogues of) these theorems are \emph{not} valid under
+Allen's supervaluation semantics, because one would have to account for worlds
+at which an infinite number of atoms have already been generated; in this way,
+the open-endedness of Allen's semantics relies crucially on a condition that is
+simply impossible under a Brouwerian conception of mathematical activity.
+
+\begin{remark}
+  One way to recover something essentially equivalent to the supervaluation
+  semantics in our style would be to replace the current semantic universe
+  $\Define{\SemU}{\Classes^\SCtx}$ with an extended universe
+  $\Define{\SemU_\infty}{\Classes^{\SCtx_\infty}}$, $\SCtx_\infty$ the set of
+  possibly-infinite symbol contexts and injective, sort-preserving renamings.
+  Externally, by Theorem~\ref{thm:no-nat-injection} we could still never
+  exhibit an embedding of the naturals, but internally, we would be unable to
+  rule it out.
+\end{remark}
 

### 2015-12-24 paper.tex
wording

@@
 
-Another seemingly paradoxical observation is also possible:
+Another seemingly paradoxical observation is possible:
 
@@
 embedded into the other commits us to a truly intuitionistic ontology, as
-opposed to merely constructive one. In particular, our framework is
+opposed to a merely constructive one. In particular, our framework is
 anti-classical in at least a na\"ive sense, because the Axiom of Choice puts

### 2015-12-24 paper.tex
remove remark; improve wording

@@
 Allen's supervaluation semantics, because one would have to account for worlds
-at which an infinite number of atoms have already been generated; in this way,
+at which an infinite number of atoms had already been generated; in this way,
 the open-endedness of Allen's semantics relies crucially on a condition that is
 simply impossible under a Brouwerian conception of mathematical activity.
-
-\begin{remark}
-  One way to recover something essentially equivalent to the supervaluation
-  semantics in our style would be to replace the current semantic universe
-  $\Define{\SemU}{\Classes^\SCtx}$ with an extended universe
-  $\Define{\SemU_\infty}{\Classes^{\SCtx_\infty}}$, $\SCtx_\infty$ the set of
-  possibly-infinite symbol contexts and injective, sort-preserving renamings.
-  Externally, by Theorem~\ref{thm:no-nat-injection} we could still never
-  exhibit an embedding of the naturals, but internally, we would be unable to
-  rule it out.
-\end{remark}
 

### 2015-12-24 paper.tex
fix typo

@@
 sort-indexed (resp.\ valence-indexed) copresheaves on $\SCtx$ of closed values
-(resp\. bound terms) as follows:
+(resp.\ bound terms) as follows:
 \begin{align*}

### 2015-12-24 paper.tex
typo

@@
 the propositions-as-types principle is concerned, Computational Type Theory is
-already anti-classical, as the undecidability of the halting program can be
+already anti-classical, as the undecidability of the halting problem can be
 observed internally.

### 2015-12-29 paper.tex
remove aborted sentence
Update notation package

@@
 \newcommand\Values{\mathbf{V}}
-\newcommand\BTms{\mathbf{B}}
+\newcommand\BExprs{\mathbf{B}}
 \newcommand\Naturals{\mathbb{N}}
@@
   \ADefine{
-    \BTms_{v}(\Upsilon)
+    \BExprs_{v}(\Upsilon)
   }{
@@
 family of valenced-indexed relations
-$\Of{R_v}{\Hom{\SemU}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for
+$\Of{R_v}{\Hom{\SemU}{\mathbb{1}}{\Pow{\BExprs_v^2}}}$ for
 $\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
@@
     \PBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
-      \MkBTm{\vec{u}}{\vec{x}}{M}
-    }{
-      \MkBTm{\vec{v}}{\vec{y}}{N}
+      \MkAbs{\vec{u}}{\vec{x}}{M}
+    }{
+      \MkAbs{\vec{v}}{\vec{y}}{N}
     }
@@
   \end{tikzcd}%
-  , for all $\Member{\vec{E},\vec{F}}{\BTms^{[\vec{v}]}(\Upsilon')}$ and
+  , for all $\Member{\vec{E},\vec{F}}{\BExprs^{[\vec{v}]}(\Upsilon')}$ and
   $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$, if we can conclude
@@
   \end{tikzcd}%
-  and $\Member{\vec{E},\vec{F}}{\BTms^{[\vec{v}]}(\Upsilon')}$ and
+  and $\Member{\vec{E},\vec{F}}{\BExprs^{[\vec{v}]}(\Upsilon')}$ and
   $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$.
@@
 We will demonstrate how to define certain refinements, and prove that they are
-types. The refinements below will differ from the ones defined in Section~
+types.
 

### 2015-12-31 paper.tex
Define symbol contexts, valences, arities, operators

@@
 
-See~\cite{sterling-morrison:2015} for the development of abstract binding trees
-with symbols. \todo[inline]{give a brief description of the framework, and present its
-rules.}
+We will briefly reproduce the syntax of abts here;
+see~\cite{sterling-morrison:2015} for a full development of the syntax and
+semantics of multi-sorted nominal abstract binding trees, however. The point
+of nominal abts is that they extend the standard treatments of second order
+universal algebra~\cite{fiore-hur:2010} with a notion of \emph{symbolic
+parameter} and name abstraction; this is absolutely essential for correctly
+treating the syntax of assignables, open exceptions, fluid bindings, and
+processes, among other things. See Harper's \emph{Practical Foundations for
+Programming Languages}~$\cite{harper:2016}$ for numerous examples of the use of
+symbolic parameters.
+
+Fix a set of \emph{sorts} $\Sorts$; we'll say $\IsSort{\tau}$ when
+$\Member{\tau}{\Sorts}$. Let $\SymSets$ be the category of finite cardinals and
+injective maps between them; using the comma construction, we can define the
+category $\SCtx$ of \emph{contexts} of symbols, whose objects are finite sets
+of symbols $U$ and sort-assignments
+\begin{tikzcd}[cramped]
+  \IMode{U}\arrow[r, "\IMode{\mathfrak{s}}"] &\IMode{\Sorts}
+\end{tikzcd}, and whose morphisms are
+sort-preserving renamings; we will write $\Upsilon$ for a symbol context
+$(U,\mathfrak{s})$.
+
+We will say that an $\Sorts$-valence is a form
+$\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$, where
+$\IMode{\vec{\sigma}\cup\vec{\tau}}\subseteq\IMode{\Sorts}$ and
+$\IsSort{\tau}$; valences represent the \emph{binding structure} of an
+abstraction of sort $\tau$, where $\vec{\sigma}$ are the sorts of the bound
+symbols and $\vec{\tau}$ are the sorts of the bound variables. When $v$ is a
+valence, we will say $\IsValence{v}$. An $\Sorts$-arity is of the form
+$\MkArity{\vec{v}}{\tau}$, where $\vec{v}$ are $\Sorts$-valences and
+$\IsSort{\tau}$; arities classify \emph{operators} (also called ``function
+symbols'' in older treatments of algebra). When $a$ is an arity, we will say
+$\IsArity{a}$.
+
+An \emph{operator signature} is an arity-indexed family of copresheaves on
+$\SCtx$, namely $\HypJ{\Of{\Operators_a}{\Sets^\SCtx}}{\IsArity{a}}$; this is
+an intensional generalization of the state of affairs for standard universal
+algebra, where operators are arity-indexed families of \emph{sets}. The point
+of fibering over symbol contexts is to introduce an open-ended indexing
+mechanism for operators; the functorial liftings of $\SCtx$-morphisms
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}
+equip operators $\Member{\vartheta}{\Operators_a(\Upsilon)}$ with a suitable
+renaming action $\Member{\Rename{\rho}{\vartheta}}{\Operators_a(\Upsilon')}$.%
+%
+\footnote{
+  In~\cite{sterling-morrison:2015}, the authors defined operator signatures as
+  \emph{sheaves} on $\OpCat{(\SCtx)}$ under the atomic topology (alternatively,
+  pullback-preserving presheaves), where all non-empty sieves cover; the sheaf
+  condition induces the unique existence of \emph{strengthenings} for
+  operators. However, because we do not use strengthening in the present
+  development, we have omitted this part of the apparatus for the sake of
+  simplicity.
+}
+
+% \todo[inline]{give a brief description of the framework, and present its rules.}
 

### 2015-12-31 paper.tex
finish up definition of abts (close #14)

@@
 sort-preserving renamings; we will write $\Upsilon$ for a symbol context
-$(U,\mathfrak{s})$.
+$(U,\mathfrak{s})$.  We also write $\IsSymCtx{\Upsilon}$ when $\Member{\Upsilon}{\SCtx}$.
 
@@
 symbols and $\vec{\tau}$ are the sorts of the bound variables. When $v$ is a
-valence, we will say $\IsValence{v}$. An $\Sorts$-arity is of the form
-$\MkArity{\vec{v}}{\tau}$, where $\vec{v}$ are $\Sorts$-valences and
-$\IsSort{\tau}$; arities classify \emph{operators} (also called ``function
-symbols'' in older treatments of algebra). When $a$ is an arity, we will say
-$\IsArity{a}$.
-
-An \emph{operator signature} is an arity-indexed family of copresheaves on
+valence, we will say $\IsValence{v}$; let $\Valences$ be the set of $\Sorts$-valences.
+
+An $\Sorts$-arity is of the form $\MkArity{\vec{v}}{\tau}$, where $\vec{v}$ are
+$\Sorts$-valences and $\IsSort{\tau}$; arities classify \emph{operators} (also
+called ``function symbols'' in older treatments of algebra). When $a$ is an
+arity, we will say $\IsArity{a}$; let $\Arities$ be the set of $\Sorts$-arities.
+
+An $\Sorts$-\emph{operator signature} is an $\Sorts$-arity-indexed family of copresheaves on
 $\SCtx$, namely $\HypJ{\Of{\Operators_a}{\Sets^\SCtx}}{\IsArity{a}}$; this is
@@
 }
-
-% \todo[inline]{give a brief description of the framework, and present its rules.}
+%
+We call the pair $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ an \emph{abt signature}.
+
+Just as we have defined symbol contexts as $\SCtx$, we will use the category
+$\FinSets$ of finite cardinals and \emph{all} maps between them to define
+variable contexts $\Ctx$ and metavariable contexts $\MCtx$, writing
+$\IsVarCtx{\Gamma}$ when $\Member{\Gamma}{\Ctx}$ and $\IsMetaCtx{\Theta}$ when
+$\Member{\Theta}{\MCtx}$. The difference between $\SCtx$ and $\Ctx$ is that the
+latter allows renamings that may identify distinct names, whereas renamings in
+$\SCtx$ are always injective.
+
+For a given abt signature, the syntax of abts is given by the judgments
+$\IsAbt{\Theta}{\Upsilon}{\Gamma}{M}{\tau}$ (presupposing $\IsMetaCtx{\Theta}$,
+$\IsSymCtx{\Upsilon}$, $\IsVarCtx{\Gamma}$ and $\IsSort{\tau}$) and
+$\IsAbs{\Theta}{\Upsilon}{\Gamma}{E}{v}$ (presupposing $\IsMetaCtx{\Theta}$,
+$\IsSymCtx{\Upsilon}$, $\IsVarCtx{\Gamma}$ and $\IsValence{v}$), defined in
+mutual induction as follows:
+
+\[
+  \infer[\RuleIsAbtVar]{
+    \IsAbt{\Theta}{\Upsilon}{\Gamma}{x}{\tau}
+  }{
+    \Lookup{\Gamma}{x}{\tau}
+  }
+\]
+\[
+  \infer[\RuleIsAbtMVar]{
+    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
+      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
+    }{\tau}
+  }{
+    \begin{array}{l}
+      \Lookup{\Theta}{\MV{m}}{
+        \MkValence{
+          \sigma_0,\dots,\sigma_m
+        }{
+          \tau_0,\dots,\tau_n
+        }{\tau}
+      }
+\\
+      \HypJ{
+        \Lookup{\Upsilon}{u_i}{\sigma_i}
+      }{\Leq{i}{m}}
+\\
+      \HypJ{
+        \IsAbt{\Theta}{\Upsilon}{\Gamma}{M_i}{\tau_i}
+      }{\Leq{i}{n}}
+    \end{array}
+  }
+\]
+\[
+  \infer[\RuleIsAbtApp]{
+    \IsAbt{\Theta}{\Upsilon}{\Gamma}{
+      \App{\vartheta}{E_0,\dots,E_n}
+    }{\tau}
+  }{
+    \begin{array}{l}
+      \IsOperator{\Upsilon}{\vartheta}{
+        \MkArity{v_1,\dots,v_n}{\tau}
+      }
+\\
+      \HypJ{
+        \IsAbs{\Theta}{\Upsilon}{\Gamma}{E_i}{v_i}
+      }{\Leq{i}{n}}
+    \end{array}
+  }
+\]
+\[
+  \infer[\RuleIsAbtAbs]{
+    \IsAbs{\Theta}{\Upsilon}{\Gamma}{
+      \MkAbs{\vec{u}}{\vec{x}}{M}
+    }{
+      \MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}
+    }
+  }{
+    \IsAbt{
+      \Theta
+    }{
+      \Upsilon, \vec{u}:\vec{\sigma}
+    }{
+      \Gamma,\vec{x}:{\vec{\tau}}
+    }{
+      M
+    }{\tau}
+  }
+\]
+
+We have several meta-operations on abts, including the calculation of free
+symbols $\FS{M}$, the calculation of free variables $\FV{M}$, the transport
+$\Rename{\rho}{M}$ across a renaming $\rho$, the substitution $\Subst{N}{x}{M}$
+of a term $N$ for the variable $x$ in $M$, and the substitution
+$\Subst{E}{\MV{m}}{M}$ of an abstraction $E$ for the metavariable $\MV{m}$ in
+$M$. Please see~\cite{sterling-morrison:2015} for the explicit definition of
+these operations.
 

### 2016-01-01 paper.tex
introduce → induce
add nominal independence modality (#13)

@@
 algebra, where operators are arity-indexed families of \emph{sets}. The point
-of fibering over symbol contexts is to introduce an open-ended indexing
+of fibering over symbol contexts is to induce an open-ended indexing
 mechanism for operators; the functorial liftings of $\SCtx$-morphisms
@@
 \subsubsection{The type of atoms}
+\label{sec:atom-type}
 \newcommand\SortAtom{\mathtt{atom}}
@@
 
+\subsection{Refinement Modalities}
+
+In Section~\ref{sec:atom-type} we introduced our first \emph{intensional} type,
+that is, a type whose meaning varies over worlds $\Upsilon$. It is also
+possible to introduce \emph{refinement modalities} which change the
+meaning of a refinement in an intensional manner.
+
+\subsubsection{Nominal Independence}
+
+\newcommand\Indep[2]{#1\setminus #2}
+
+Mark Bickford introduced in~\cite{bickford:2008} a type in Nuprl that expresses
+the independence of a term $M$ of type $A$ from a symbol $a$, written
+$M:A\parallel a$, which is used in order to express the notion that some agent
+is not aware of a particular token/atom. In our setting, nominal independence
+is best expressed through a comonadic refinement modality, $\Indep{\phi}{a}$, whose
+inhabitants are the inhabitants of $\phi$ which do not make use of the symbol
+$a$. To be precise, we intend to explain the following refinement formation
+rule:
+\[
+  \infer={
+    \Refines{\Upsilon,a:\sigma}{\Indep{\phi}{a}}{\tau}
+  }{
+    \Refines{\Upsilon,a:\sigma}{\phi}{\tau}
+  }
+\]
+
+The definition of nominal independence is as follows, for any renaming
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon,a:\sigma}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+\end{tikzcd}%
+:
+\[
+  \infer={
+    \RMemEq{\Upsilon'}{M}{N}{\Indep{\phi}{\rho(a)}}
+  }{
+    \exists\Member{L}{\Exprs_\tau(\Upsilon'\setminus \rho(a))}.
+    \begin{array}{l}
+      \quad \RMemEq{\Upsilon'}{L}{M}{\phi}\\
+      \quad \RMemEq{\Upsilon'}{L}{N}{\phi}
+    \end{array}
+  }
+\]
+
+In other words, a construction is independent of a symbol when it is equal to
+one which was constructed without knowledge of that symbol.
+
+\begin{thm}
+  The refinement operator $\Indep{-}{a}$ is an idempotent comonad on refinements
+  in the sense that the following rules are valid:
+
+  \begin{gather*}
+    \infer[\textit{extract}]{
+      \RMemEq{\Upsilon}{M}{N}{\phi}
+    }{
+      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
+    }
+    \qquad
+    \infer=[\textit{duplicate}]{
+      \RMemEq{\Upsilon}{M}{N}{\Indep{\Indep{\phi}{a}}{a}}
+    }{
+      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
+    }
+    \\[6pt]
+    \infer[\textit{lift}]{
+      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
+      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
+      \implies
+      \RMemEq{\Upsilon}{M}{N}{\Indep{\psi}{a}}
+    }{
+      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
+      \RMemEq{\Upsilon}{M}{N}{\phi}
+      \implies
+      \RMemEq{\Upsilon}{M}{N}{\psi}
+    }
+  \end{gather*}
+\end{thm}
+\begin{proof}
+  The first two inferences are obviously evident. To validate \textit{lift},
+  suppose that $\RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}$; by inversion, we
+  have an $\Member{L}{\Exprs_\tau(\Upsilon\setminus a)}$ such that
+  $\RMemEq{\Upsilon}{L}{M}{\phi}$ and $\RMemEq{\Upsilon}{L}{N}{\phi}$.  Now, it
+  suffices to exhibit an $\Member{L'}{\Exprs_\tau(\Upsilon\setminus a)}$ such
+  that $\RMemEq{\Upsilon}{L'}{M}{\psi}$ and $\RMemEq{\Upsilon}{L'}{N}{\psi}$;
+  but we may simply choose $\Match{L'}{L}$ and discharge the remaining subgoals
+  by appealing to the premise of the rule.
+\end{proof}
+
+%% \subsubsection{Nominal Necessity}
+%%
+%% \subsubsection{Nominal Possibility}
+%%
 \section{Related Work}
@@
 
+\section*{Acknowledgements}
+
+The first author wishes to thank Robert Harper for the inspiration to pursue a
+generalization of the Nuprl type theory from the refinements perspective. We
+also thank Mark Bickford for clarifying the status of atoms and generative
+phenomena in Nuprl; finally, we owe a debt of gratitude to Vincent Rahli and
+Abhishek Anand for their tireless work on the \emph{Nuprl in Coq} project,
+which was helpful in our efforts to understand Howe's computational equivalence
+and its generalization to the nominal setting.
+
 \ifdraft{}{

### 2016-01-01 paper.tex
define necessity; add futher remark to independence (#13)

@@
 
-\begin{thm}
+\begin{remark}
+  Note that we could not have simply used
+  $\RMemEq{\Upsilon'\setminus\rho(a)}{M}{N}{\phi}$ as the premise in our
+  definition, since that would be outside the range of significance of this
+  form of judgment (i.e.\ the presuppositions fail to be satisfied); in
+  particular, it is not guaranteed that the refinement $\phi$ is defined at
+  $\Upsilon'\setminus\rho(a)$, and nor do we have
+  $\Member{M,N}{\Exprs_\tau(\Upsilon'\setminus\rho(a))}$.
+\end{remark}
+
+\begin{thm} \label{thm:independence-comonadic}
   The refinement operator $\Indep{-}{a}$ is an idempotent comonad on refinements
@@
 
-%% \subsubsection{Nominal Necessity}
-%%
+\subsubsection{Nominal Necessity}
+
+Another comonadic refinement modality is \emph{nominal necessity}
+$\HypJ{\Refines{\Upsilon}{\square\phi}{\tau}}{\Refines{\Upsilon}{\phi}{\tau}}$,
+which expresses the independence of an object from \emph{any} atoms at all. For
+any renaming
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+\end{tikzcd}%
+, we define nominal necessity as follows:
+\[
+  \infer={
+    \RMemEq{\Upsilon'}{M}{N}{\square\phi}
+  }{
+    \exists\Member{L}{\Exprs_\tau(\cdot)}.
+    \begin{array}{l}
+      \RMemEq{\Upsilon'}{L}{M}{\phi}\\
+      \RMemEq{\Upsilon'}{L}{N}{\phi}
+    \end{array}
+  }
+\]
+
+\begin{thm}
+  Nominal independence is an idempotent comonadic modality on refinements.
+\end{thm}
+\begin{proof}
+  The demonstration is essentially analogous to the one for independence in
+  Theorem~\ref{thm:independence-comonadic}.
+\end{proof}
+
 %% \subsubsection{Nominal Possibility}

### 2016-01-01 paper.tex
extend remark on behavioral character of the independence modality (#13)

@@
   $\Member{M,N}{\Exprs_\tau(\Upsilon'\setminus\rho(a))}$.
+
+  In fact, far from being a technical nuisance, the definition that is forced
+  by the nature of our presuppositions is completely desirable, since it
+  expresses a \emph{behavioral} notion of independence rather than a
+  \emph{structural} one. Moreover, under the structural definition, even if
+  $\phi$ is a type, $\phi\setminus a$ might fail to be a type in case there is
+  a member of $\phi$ that uses $a$ which is merely computationally equivalent
+  to another member that is independent of $a$.
 \end{remark}

### 2016-01-01 paper.tex
wordsmithing

@@
   by the nature of our presuppositions is completely desirable, since it
-  expresses a \emph{behavioral} notion of independence rather than a
+  expresses the \emph{behavioral} notion of independence rather than the
   \emph{structural} one. Moreover, under the structural definition, even if

### 2016-01-01 paper.tex
remove independed nominal possibility section (#13)

@@
 Another comonadic refinement modality is \emph{nominal necessity}
-$\HypJ{\Refines{\Upsilon}{\square\phi}{\tau}}{\Refines{\Upsilon}{\phi}{\tau}}$,
+$\HypJ{\Refines{\Upsilon}{\Box\phi}{\tau}}{\Refines{\Upsilon}{\phi}{\tau}}$,
 which expresses the independence of an object from \emph{any} atoms at all. For
@@
   \infer={
-    \RMemEq{\Upsilon'}{M}{N}{\square\phi}
+    \RMemEq{\Upsilon'}{M}{N}{\Box\phi}
   }{
@@
 
-%% \subsubsection{Nominal Possibility}
-%%
 \section{Related Work}

### 2016-01-01 paper.tex
Add possibility modality; fix rule for functorial lifting of modalities (#13)

@@
     \infer[\textit{lift}]{
-      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
-      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
+      \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}.\
+      \forall\Member{M,N}{\Exprs_\tau(\Upsilon')}.\
+      \RMemEq{\Upsilon'}{M}{N}{\Indep{\phi}{a}}
       \implies
-      \RMemEq{\Upsilon}{M}{N}{\Indep{\psi}{a}}
-    }{
-      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
-      \RMemEq{\Upsilon}{M}{N}{\phi}
+      \RMemEq{\Upsilon'}{M}{N}{\Indep{\psi}{a}}
+    }{
+      \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}.\
+      \forall\Member{M,N}{\Exprs_\tau(\Upsilon')}.\
+      \RMemEq{\Upsilon'}{M}{N}{\phi}
       \implies
-      \RMemEq{\Upsilon}{M}{N}{\psi}
+      \RMemEq{\Upsilon'}{M}{N}{\psi}
     }
@@
 
+\subsubsection{Nominal Possibility}
+
+\emph{Nominal possibility} expresses the idea that there \emph{exists} an
+extension of the current symbol context at which a refinement is inhabited.
+However, it is not possible to define a useful possibility modality in the same
+way as the others, without imposing requirements on the abt signature. In
+particular, we need to introduce a new sort and a new canonical operator for
+introducing possibility:
+\[
+  \infer{
+    \IsSort{\Diamond\tau}
+  }{
+    \IsSort{\tau}
+  }\qquad
+  \infer{
+    \IsCanOperator{\Upsilon}{\Diamond[\vec{\sigma}]}{
+      \MkArity{
+        \MkValence{\vec{\sigma}}{}{\tau}
+      }{\Diamond\tau}
+    }
+  }{
+  }
+\]
+
+Then, we can explain the following refinement formation rule:
+\[
+  \infer={
+    \Refines{\Upsilon}{\Diamond\phi}{\Diamond\tau}
+  }{
+    \Refines{\Upsilon}{\phi}{\tau}
+  }
+\]
+
+We will begin by defining a refinement $\Diamond_0\phi$ on values, for any renaming
+\begin{tikzcd}[cramped]
+  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+\end{tikzcd}%
+:
+\[
+  \infer={
+    \RMemEq{\Upsilon'}{
+      \App{
+        \Diamond[\vec{\sigma}]
+      }{
+        \MkAbs{\vec{u}}{}{M}
+      }
+    }{
+      \App{
+        \Diamond[\vec{\sigma}]
+      }{
+        \MkAbs{\vec{v}}{}{N}
+      }
+    }{\Diamond_0\phi}
+  }{
+    \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon'}\cup\vec{u}\cup\vec{v}}.\
+    \RMemEq{\Upsilon',\vec{w}:\vec{\sigma}}{\Rename{(\vec{u}\mapsto\vec{w}}{M})}{\Rename{(\vec{v}\mapsto\vec{w})}{N}}{\phi}
+  }
+\]
+
+Then, the possibility modality is defined as the computational extension of
+$\Diamond_0-$, $\Define{\Diamond\phi}{\VToCType\left(\Diamond_0\phi\right)}$.
+
+\begin{thm}
+
+  Nominal possibility is an idempotent monadic modality, in the sense that the
+  following rules are valid:
+%
+  \[
+    \infer[\textit{return}]{
+      \RMemEq{\Upsilon}{
+        \App{
+          \Diamond[\cdot]
+        }{
+          \MkAbs{}{}{M}
+        }
+      }{
+        \App{
+          \Diamond[\cdot]
+        }{
+          \MkAbs{}{}{N}
+        }
+      }{\Diamond\phi}
+    }{
+      \RMemEq{\Upsilon}{M}{N}{\phi}
+    }
+  \]
+  \[
+    \infer=[\textit{join}]{
+      \RMemEq{\Upsilon}{
+        \App{
+          \Diamond[\vec{\sigma}_0,\vec{\sigma}_1]
+        }{
+          \MkAbs{\vec{u}_0,\vec{u}_1}{}{M}
+        }
+      }{
+        \App{
+          \Diamond[\vec{\sigma}_0,\vec{\sigma}_1]
+        }{
+          \MkAbs{\vec{v}_0,\vec{v}_1}{}{M}
+        }
+      }{\Diamond\phi}
+    }{
+      \RMemEq{\Upsilon}{
+        \App{
+          \Diamond[\vec{\sigma}_0]
+        }{
+          \MkAbs{\vec{u}_0}{}{
+            \App{
+              \Diamond[\vec{\sigma}_1]
+            }{
+              \MkAbs{\vec{u}_1}{}{M}
+            }
+          }
+        }
+      }{
+        \App{
+          \Diamond[\vec{\sigma_0}]
+        }{
+          \MkAbs{\vec{v}_0}{}{
+            \App{
+              \Diamond[\vec{\sigma}_1]
+            }{
+              \MkAbs{\vec{v}_1}{}{N}
+            }
+          }
+        }
+      }{\Diamond\Diamond\phi}
+    }
+  \]
+  \[
+    \infer[\textit{lift}]{
+      \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}.\
+      \forall\Member{M,N}{\Exprs_{\Diamond\tau}(\Upsilon')}.\
+      \RMemEq{\Upsilon'}{M}{N}{\Diamond\phi}
+      \implies
+      \RMemEq{\Upsilon'}{M}{N}{\Diamond\psi}
+    }{
+      \forall
+        \begin{tikzcd}[cramped, ampersand replacement = \&]
+          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+        \end{tikzcd}.\
+      \forall\Member{M,N}{\Exprs_\tau(\Upsilon')}.\
+      \RMemEq{\Upsilon'}{M}{N}{\phi}
+      \implies
+      \RMemEq{\Upsilon'}{M}{N}{\psi}
+    }
+  \]
+\end{thm}
+\begin{proof}
+
+  As in Theorem~\ref{thm:independence-comonadic}, the validity first two rules
+  is obvious. To demonstrate the \textit{lift} rule, fix a renaming
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &
+    \IMode{\Upsilon'}\arrow[r,hook,"\IMode{\rho'}"] &
+    \IMode{\Upsilon''}\equiv\OMode{\Upsilon', \vec{w}:\vec{\sigma}}
+  \end{tikzcd}
+  and $\Member{M,N}{\Exprs_\tau(\Upsilon'')}$ such that
+  $\RMemEq{\Upsilon''}{M}{N}{\phi}$; then it suffices to show that
+  $\RMemEq{\Upsilon'}{\App{\Diamond[\vec{\sigma}]}{\MkAbs{\vec{w}}{}{M}}}{\App{\Diamond[\vec{\sigma}]}{\MkAbs{\vec{w}}{}{N}}}{\Diamond\psi}$, which is to say,
+  $\RMemEq{\Upsilon''}{M}{N}{\psi}$. But this follows directly from the premise, setting $\Match{\Upsilon'}{\Upsilon''}$.
+
+\end{proof}
+
 \section{Related Work}

### 2016-01-01 paper.tex
wordsmithing

@@
 is best expressed through a comonadic refinement modality, $\Indep{\phi}{a}$, whose
-inhabitants are the inhabitants of $\phi$ which do not make use of the symbol
+inhabitants are those of $\phi$ which do not make use of the symbol
 $a$. To be precise, we intend to explain the following refinement formation

### 2016-01-03 paper.tex
make modalities a section rather than a subsection
add "structural"
fix typos

@@
 operations, possibly subject to an arbitrary equational theory
-\cite{cartmell:1986, dybjer:1996}; we will call this \emph{formal type theory}.
-Typing, in such a setting, is a mere matter of grammar and is nearly always
-decidable. In hindsight, we may observe that this is the sort of type theory
-which Martin-L\"of first proposed in 1972~\cite{martin-lof:1972}, even if we
-will admit that this was not the intention at the time. A model for such a type
-theory is usually given by interpreting the types or sorts as presheaves or
-sheaves over contexts of hypotheses, and as such, a proof theoretic
-interpretation of the hypothetical judgment is possible.
+\cite{cartmell:1986, dybjer:1996}; we will call this \emph{formal} or
+\emph{structural type theory}.  Typing, in such a setting, is a mere matter of
+grammar and is nearly always decidable. In hindsight, we may observe that this
+is the sort of type theory which Martin-L\"of first proposed in
+1972~\cite{martin-lof:1972}, even if we will admit that this was not the
+intention at the time. A model for such a type theory is usually given by
+interpreting the types or sorts as presheaves or sheaves over contexts of
+hypotheses, and as such, a proof theoretic interpretation of the hypothetical
+judgment is possible.
 
@@
 ``implicit'' version of existential quantification, and can be used to
-simultaneously specify and implement program without the evidence for $\psi$
+simultaneously specify and implement programs without the evidence for $\psi$
 intruding into the extracted algorithm.
@@
       \end{tikzcd}%
-      , for all $\Member{\sigma}{\Sorts}$ and $\Member{X,X',Y}{\Exprs_\sigma(\Upsilon'')}$, if
+      , for all $\Member{\sigma}{\Sorts}$ and $\Member{X,X',Y}{\Exprs_\sigma(\Upsilon'')}$,
       from $\EvalN{\Upsilon''}{\sigma}{k'}{X}{X'}$ for $\IMode{k'}<\IMode{k}$ and
@@
 
-\subsection{Refinement Modalities}
+\section{Refinement Modalities}
 
@@
 
-\subsubsection{Nominal Independence}
+\subsection{Nominal Independence}
 
@@
 
-\subsubsection{Nominal Necessity}
+\subsection{Nominal Necessity}
 
@@
 
-\subsubsection{Nominal Possibility}
+\subsection{Nominal Possibility}
 

### 2016-01-03 paper.tex
Fix error in definition of approx

@@
 which respects a single ``layer'' of computation. We will say
-$\PBinRel{[R_\tau]}{\Upsilon}{M}{N}$ when, supposing
-$\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}}$ such that
-$\IsCanOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$,
-for any
+$\PBinRel{[R_\tau]}{\Upsilon}{M}{N}$ when, for any
 \begin{tikzcd}[cramped]
   \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
-\end{tikzcd}
-, we have $\Eval{\Upsilon'}{\tau}{\Rename{\rho}{N}}{\App{\Rename{\rho}{\vartheta}}{\Rename{\rho}{\Vec{F}}}}$ and
-$\PBinRel{R_{\vec{v}}}{\Upsilon'}{\Rename{\rho}{\vec{E}}}{\Rename{\rho}{\vec{F}}}$.
+\end{tikzcd}%
+, supposing
+$\Eval{\Upsilon'}{\tau}{\Rename{\rho}{M}}{\App{\vartheta}{\Vec{E}}}$ such that
+$\IsCanOperator{\Upsilon'}{\vartheta}{\MkArity{\vec{v}}{\tau}}$,
+we have $\Eval{\Upsilon'}{\tau}{\Rename{\rho}{N}}{\App{\vartheta}{\Vec{F}}}$ and
+$\PBinRel{R_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{F}}$.
 

### 2016-01-03 paper.tex
correct proof of approx-refl

@@
   of $[-]$, this means that we have to show that when
-  $\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\vec{E}}}$, then
+  $\Eval{\Upsilon'}{\tau}{\Rename{\rho}{M}}{\App{\vartheta}{\vec{E}}}$ for
+  \begin{tikzcd}[cramped]
+    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
+  \end{tikzcd}%
+  , then
   $\IsEq{\vec{E}}{\vec{E}}$; but this is always the case, because of the

### 2016-01-20 paper.tex
Spell out coinductive definition of approx
fix / clarify wording

@@
 an intensional generalization of the state of affairs for standard universal
-algebra, where operators are arity-indexed families of \emph{sets}. The point
+algebra, where the operators are an arity-indexed family of \emph{sets}. The point
 of fibering over symbol contexts is to induce an open-ended indexing
@@
   written pointwise as $\PBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}$.
+  In other words:
+  \[
+    \infer{
+      \PBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}
+    }{
+      \forall
+      \begin{tikzcd}[cramped, ampersand replacement = \&]
+        \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
+      \end{tikzcd}.\
+      \land\left\{
+        \begin{array}{l}
+          \Eval{\Upsilon'}{\tau}{\Rename{\rho}{M}}{\App{\vartheta}{\vec{E}}}\\
+          \IsCanOperator{\Upsilon'}{\vartheta}{\MkArity{\vec{v}}{\tau}}
+        \end{array}
+      \right\}
+      \implies
+      \land\left\{
+        \begin{array}{l}
+          \Eval{\Upsilon'}{\tau}{\Rename{\rho}{N}}{\App{\vartheta}{\Vec{F}}}\\
+          \PBinRel{\preccurlyeq_{\vec{v}}}{\Upsilon'}{\vec{E}}{\vec{F}}
+        \end{array}
+      \right\}
+    }
+  \]
 \end{definition}

### 2016-03-15 paper.tex
fix typos

@@
 $\FRefines{\Upsilon}{x:\phi}{\psi}{\tau}$ such that
-$\Refines{\Upsilon}{\phi}{\sigma}$, including the family intersection, the
-dependent intersection, the family union, and the set comprehension.
+$\Refines{\Upsilon}{\phi}{\sigma}$, we may construct a number of new refinement
+connectives, including the family intersection, the dependent intersection, the
+family union, and the set comprehension.
 
@@
 
-In this section we generalize Howe's notion of lazy computation
+In this section we adapt Howe's notion of lazy computation
 system~\cite{howe:1989} to the multi-sorted, symbol-parameterized setting. An
@@
 \begin{thm}
-  Nominal independence is an idempotent comonadic modality on refinements.
+  Nominal necessity is an idempotent comonadic modality on refinements.
 \end{thm}

### 2016-07-05 paper.tex
corrections, typos, update notation package

@@
 
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 \usepackage{leftidx}
@@
 intention at the time. A model for such a type theory is usually given by
-interpreting the types or sorts as presheaves or sheaves over contexts of
-hypotheses, and as such, a proof theoretic interpretation of the hypothetical
-judgment is possible.
-
-Secondly, there is the view of type theory as semi-formal theory of
+interpreting the types or sorts as (families of) presheaves or sheaves over
+contexts of hypotheses, and as such, a proof theoretic interpretation of the
+hypothetical judgment is possible.
+
+Secondly, there is the view of type theory as a semi-formal theory of
 constructions for the Brouwer-Heyting-Kolmogorov interpretation of
